---
title: "PBC attained-age landmark grid: overlap and event-count preflight"
output:
  html_document:
    self_contained: true
    toc: true
    code_folding: show
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center")
library(here)
devtools::load_all(here::here(), quiet = TRUE)

prediction_window <- 5
dense_age_landmarks <- seq(42, 56, by = 2)
cause_labels <- c("Transplant", "Death")
```

# Purpose

I kept the original time-scale comparison unchanged. This separate analysis
tests whether a more closely spaced attained-age grid identifies landmark main
effects and whether its competing-event counts support a bootstrap analysis.
The candidate grid is a sensitivity analysis, not a replacement for the
entry-age-quartile analysis.

```{r analysis-data}
pbc_long <- get_pbc_long()
entry_ages <- unique(pbc_long[c("id", "age")])$age
current_age_landmarks <- round(
  unname(quantile(entry_ages, c(0.25, 0.50, 0.75), type = 8)), 1
)

pbc_age <- transform(
  pbc_long,
  log_bili_z = (log(bili) - mean(log(pbc_long$bili))) /
    sd(log(pbc_long$bili)),
  albumin_z = (albumin - mean(pbc_long$albumin)) / sd(pbc_long$albumin),
  entry_age = age,
  exit_age = age + time,
  visit_age = age + tstart
)

stopifnot(
  !anyDuplicated(unique(pbc_age[c("id", "entry_age")])$id),
  all(is.finite(pbc_age$log_bili_z)),
  all(is.finite(pbc_age$albumin_z)),
  all(diff(current_age_landmarks) > prediction_window),
  all(diff(dense_age_landmarks) < prediction_window)
)
```

# Why the current age grid cannot estimate landmark main effects

I plotted the original landmark windows to make the lack of overlap visible.
Each event risk set lies within one window. An LM main effect is therefore
constant within that risk set and cancels from the Cox partial likelihood.
The model still contains bilirubin-by-LM and albumin-by-LM interactions; it
does not contain an estimable LM main effect on this grid.

```{r current-window-plot, fig.width=7, fig.height=3.3}
draw_current_windows <- function() {
  window_ends <- current_age_landmarks + prediction_window
  plot(NA, xlim = range(c(current_age_landmarks, window_ends)) + c(-0.5, 0.5),
       ylim = c(0.5, length(current_age_landmarks) + 0.5),
       yaxt = "n", xlab = "Attained age (years)", ylab = "",
       main = "Current attained-age prediction windows", bty = "n")
  axis(2, at = seq_along(current_age_landmarks),
       labels = paste0("LM ", format(current_age_landmarks, nsmall = 1)),
       las = 1, tick = FALSE)
  abline(v = c(current_age_landmarks, window_ends), col = "grey90", lty = 3)
  segments(current_age_landmarks, seq_along(current_age_landmarks),
           window_ends, seq_along(current_age_landmarks),
           lwd = 7, col = "#32658A", lend = "butt")
  points(current_age_landmarks, seq_along(current_age_landmarks), pch = 16)
  points(window_ends, seq_along(current_age_landmarks), pch = 1)
  text((current_age_landmarks + window_ends) / 2,
       seq_along(current_age_landmarks) + 0.16,
       labels = sprintf("%.1f–%.1f", current_age_landmarks, window_ends),
       cex = 0.85)
}

draw_current_windows()
pdf(here::here("analysis", "current_age_windows.pdf"),
    width = 7, height = 3.3)
draw_current_windows()
dev.off()
```

```{r current-window-check}
current_stack <- stack_data(
  pbc_age, list(time = "exit_age", status = "status"),
  lms = current_age_landmarks, w = prediction_window,
  covs = list(fixed = c("male", "stage", "trt"),
              varying = c("log_bili_z", "albumin_z")),
  format = "long", id = "id", rtime = "visit_age",
  entry = "entry_age", entry_mode = "strict"
)

risk_set_lm_counts <- function(stack) {
  event_times <- sort(unique(stack$data$exit_age[stack$data$status > 0]))
  vapply(event_times, function(event_age) {
    at_risk <- stack$data$LM < event_age & stack$data$exit_age >= event_age
    length(unique(stack$data$LM[at_risk]))
  }, integer(1))
}

current_lm_counts <- risk_set_lm_counts(current_stack)
stopifnot(length(current_lm_counts) > 0L,
          all(current_lm_counts == 1L))
knitr::kable(data.frame(
  grid = "Current entry-age quartiles",
  event_times = length(current_lm_counts),
  event_times_with_multiple_LMs = sum(current_lm_counts > 1L)
))
```

# Dense-grid risk sets

I used the same strict-entry rule, event coding, prediction window, and
laboratory transformations as the primary attained-age analysis. I checked
the number of people and both event types in every window before fitting or
resampling. These counts are stacked-window counts: an event can contribute
to more than one overlapping window.

```{r dense-risk-sets}
dense_stack_raw <- stack_data(
  pbc_age, list(time = "exit_age", status = "status"),
  lms = dense_age_landmarks, w = prediction_window,
  covs = list(fixed = c("male", "stage", "trt"),
              varying = c("log_bili_z", "albumin_z")),
  format = "long", id = "id", rtime = "visit_age",
  entry = "entry_age", entry_mode = "strict"
)

dense_risk_sets <- do.call(rbind, lapply(dense_age_landmarks, function(age) {
  rows <- dense_stack_raw$data[dense_stack_raw$data$LM == age, , drop = FALSE]
  data.frame(
    landmark = age,
    window_end = age + prediction_window,
    subjects = length(unique(rows$id)),
    deaths = sum(rows$status == 2L),
    transplants = sum(rows$status == 1L),
    censored = sum(rows$status == 0L)
  )
}))

dense_lm_counts <- risk_set_lm_counts(dense_stack_raw)
stopifnot(
  identical(dense_stack_raw$entry_mode, "strict"),
  all(dense_stack_raw$data$entry_age <= dense_stack_raw$data$LM),
  all(dense_stack_raw$data$LM < dense_stack_raw$data$exit_age),
  !anyDuplicated(dense_stack_raw$data[c("id", "LM")]),
  all(dense_risk_sets$subjects > 0L),
  all(dense_risk_sets$subjects ==
        dense_risk_sets$deaths + dense_risk_sets$transplants +
        dense_risk_sets$censored),
  any(dense_lm_counts > 1L)
)

knitr::kable(dense_risk_sets,
             caption = "Strict-entry risk sets on the candidate age grid")
knitr::kable(data.frame(
  grid = "Candidate two-year grid",
  event_times = length(dense_lm_counts),
  event_times_with_multiple_LMs = sum(dense_lm_counts > 1L),
  maximum_LMs_in_one_risk_set = max(dense_lm_counts)
))
```

The dense windows overlap, and the observed event risk sets contain multiple
landmark values. This removes the exact non-identifiability of LM main effects
seen on the original grid. It does not by itself establish that the
competing-risk fit or its bootstrap is stable: the event-count table includes
windows with very few transplants.

# Candidate supermodel fit

I used a centered quadratic landmark basis for both main effects and the two
laboratory interactions. The other predictors and the cause-specific CSC fit
match the primary attained-age specification.

```{r dense-supermodel}
dense_center <- mean(dense_age_landmarks)
dense_half_range <- diff(range(dense_age_landmarks)) / 2
dense_basis <- list(
  function(age) (age - dense_center) / dense_half_range,
  function(age) ((age - dense_center) / dense_half_range)^2
)
dense_stack <- add_interactions(
  dense_stack_raw,
  lm_covs = c("log_bili_z", "albumin_z"),
  func_covars = dense_basis,
  func_lms = dense_basis
)
stopifnot(all(c("LM1", "LM2", "log_bili_z_LM1", "log_bili_z_LM2",
                "albumin_z_LM1", "albumin_z_LM2") %in%
              names(dense_stack$data)))

dense_formula <- Hist(exit_age, status, LM) ~
  male + stage + trt +
  log_bili_z + log_bili_z_LM1 + log_bili_z_LM2 +
  albumin_z + albumin_z_LM1 + albumin_z_LM2 +
  LM1 + LM2 + cluster(id)

capture_conditions <- function(expression) {
  warnings <- character()
  value <- withCallingHandlers(
    tryCatch(eval.parent(substitute(expression)), error = function(error) error),
    warning = function(warning) {
      warnings <<- c(warnings, conditionMessage(warning))
      invokeRestart("muffleWarning")
    }
  )
  list(value = value, warnings = unique(warnings))
}

dense_fit_record <- capture_conditions(dynamic_lm(
  dense_stack, dense_formula, type = "CSC", method = "breslow", x = TRUE
))

fit_status <- data.frame(
  error = if (inherits(dense_fit_record$value, "error")) {
    conditionMessage(dense_fit_record$value)
  } else "",
  warnings = paste(dense_fit_record$warnings, collapse = " | ")
)
knitr::kable(fit_status)

if (!inherits(dense_fit_record$value, "error")) {
  dense_fit <- dense_fit_record$value
  dense_lm_inference <- do.call(rbind, lapply(seq_along(
    dense_fit$model$causes
  ), function(index) {
    fit <- dense_fit$model$models[[index]]
    terms <- c("LM1", "LM2")
    data.frame(
      cause = cause_labels[as.integer(dense_fit$model$causes[index])],
      term = terms,
      coefficient = unname(coef(fit)[terms]),
      robust_se = unname(sqrt(diag(vcov(fit)))[terms])
    )
  }))
  stopifnot(all(is.finite(dense_lm_inference$coefficient)),
            all(is.finite(dense_lm_inference$robust_se)))
  knitr::kable(dense_lm_inference, digits = 3)
}
```

# Separate landmark fits as a sparsity check

I also tried the unchanged, one-model-per-landmark specification. I recorded
all fit warnings and errors rather than modifying a sparse fit to make it
converge. The pooled supermodel can be estimable even when a separate
competing-risk fit at one age is not.

```{r dense-disjoint-fit-check}
dense_disjoint_formula <- Hist(exit_age, status, LM) ~
  male + stage + trt + log_bili_z + albumin_z + cluster(id)

dense_disjoint_conditions <- do.call(rbind, lapply(
  dense_age_landmarks, function(age) {
    one_landmark <- dense_stack_raw
    one_landmark$data <- dense_stack_raw$data[
      dense_stack_raw$data$LM == age, , drop = FALSE
    ]
    one_landmark$end_time <- age
    fit_record <- capture_conditions(dynamic_lm(
      one_landmark, dense_disjoint_formula,
      type = "CSC", method = "breslow", x = TRUE
    ))
    data.frame(
      landmark = age,
      fit_error = if (inherits(fit_record$value, "error")) {
        conditionMessage(fit_record$value)
      } else "",
      warnings = paste(fit_record$warnings, collapse = " | ")
    )
  }
))

stopifnot(nrow(dense_disjoint_conditions) == length(dense_age_landmarks))
knitr::kable(dense_disjoint_conditions)
```

# Bootstrap decision

I did not rerun the full subject bootstrap at this preflight stage. The
separate-model fit table shows an original-sample failure in a window with no
transplants. Resampling cannot repair that specification. The pooled fit above
can still be finite because the overlapping windows and quadratic basis borrow
information, but that observation is not evidence of complete train/test
predictions across bootstrap replicates.
The next decision is whether to retain this grid, trim its sparse end, or
prespecify a different transplant model. That choice should precede an
optimism-correction run rather than be made in response to its result.

```{r bootstrap-preflight}
bootstrap_preflight <- data.frame(
  minimum_subjects = min(dense_risk_sets$subjects),
  minimum_deaths = min(dense_risk_sets$deaths),
  minimum_transplants = min(dense_risk_sets$transplants),
  windows_with_zero_transplants = sum(dense_risk_sets$transplants == 0L),
  full_bootstrap_run = FALSE
)
knitr::kable(bootstrap_preflight)
```

# Session

```{r session-info}
sessionInfo()
```
