---
title: "PBC attained-age landmark sensitivity analysis"
output:
  html_document:
    self_contained: true
    toc: true
    code_folding: show
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center")
library(here)
devtools::load_all(here::here(), quiet = TRUE)

prediction_window <- 5
dense_age_landmarks <- seq(42, 56, by = 2)
cause_labels <- c("Transplant", "Death")
n_boot <- getOption("dense_age.n_boot", 200L)
bootstrap_seed <- 20260922L
stopifnot(length(n_boot) == 1L, is.finite(n_boot),
          n_boot >= 2L, n_boot == as.integer(n_boot))
n_boot <- as.integer(n_boot)
```

# Purpose

I kept the original time-scale comparison unchanged. This separate analysis
tests whether a more closely spaced attained-age grid identifies landmark main
effects, how its hazard ratios compare with separate landmark fits, and how
often its predictions survive subject-level resampling.
The candidate grid is a sensitivity analysis, not a replacement for the
entry-age-quartile analysis.

```{r analysis-data}
pbc_long <- get_pbc_long()
entry_ages <- unique(pbc_long[c("id", "age")])$age
current_age_landmarks <- round(
  unname(quantile(entry_ages, c(0.25, 0.50, 0.75), type = 8)), 1
)

pbc_age <- transform(
  pbc_long,
  log_bili_z = (log(bili) - mean(log(pbc_long$bili))) /
    sd(log(pbc_long$bili)),
  albumin_z = (albumin - mean(pbc_long$albumin)) / sd(pbc_long$albumin),
  entry_age = age,
  exit_age = age + time,
  visit_age = age + tstart
)

stopifnot(
  !anyDuplicated(unique(pbc_age[c("id", "entry_age")])$id),
  all(is.finite(pbc_age$log_bili_z)),
  all(is.finite(pbc_age$albumin_z)),
  all(diff(current_age_landmarks) > prediction_window),
  all(diff(dense_age_landmarks) < prediction_window)
)
```

# Why the current age grid cannot estimate landmark main effects

I plotted the original landmark windows to make the lack of overlap visible.
Each event risk set lies within one window. An LM main effect is therefore
constant within that risk set and cancels from the Cox partial likelihood.
The model still contains bilirubin-by-LM and albumin-by-LM interactions; it
does not contain an estimable LM main effect on this grid.

```{r current-window-plot, fig.width=7, fig.height=3.3}
draw_current_windows <- function() {
  window_ends <- current_age_landmarks + prediction_window
  plot(NA, xlim = range(c(current_age_landmarks, window_ends)) + c(-0.5, 0.5),
       ylim = c(0.5, length(current_age_landmarks) + 0.5),
       yaxt = "n", xlab = "Attained age (years)", ylab = "",
       main = "Current attained-age prediction windows", bty = "n")
  axis(2, at = seq_along(current_age_landmarks),
       labels = paste0("LM ", format(current_age_landmarks, nsmall = 1)),
       las = 1, tick = FALSE)
  abline(v = c(current_age_landmarks, window_ends), col = "grey90", lty = 3)
  segments(current_age_landmarks, seq_along(current_age_landmarks),
           window_ends, seq_along(current_age_landmarks),
           lwd = 7, col = "#32658A", lend = "butt")
  points(current_age_landmarks, seq_along(current_age_landmarks), pch = 16)
  points(window_ends, seq_along(current_age_landmarks), pch = 1)
  text((current_age_landmarks + window_ends) / 2,
       seq_along(current_age_landmarks) + 0.16,
       labels = sprintf("%.1f–%.1f", current_age_landmarks, window_ends),
       cex = 0.85)
}

draw_current_windows()
pdf(here::here("analysis", "current_age_windows.pdf"),
    width = 7, height = 3.3)
draw_current_windows()
dev.off()
```

```{r current-window-check}
current_stack <- stack_data(
  pbc_age, list(time = "exit_age", status = "status"),
  lms = current_age_landmarks, w = prediction_window,
  covs = list(fixed = c("male", "stage", "trt"),
              varying = c("log_bili_z", "albumin_z")),
  format = "long", id = "id", rtime = "visit_age",
  entry = "entry_age", entry_mode = "strict"
)

risk_set_lm_counts <- function(stack) {
  event_times <- sort(unique(stack$data$exit_age[stack$data$status > 0]))
  vapply(event_times, function(event_age) {
    at_risk <- stack$data$LM < event_age & stack$data$exit_age >= event_age
    length(unique(stack$data$LM[at_risk]))
  }, integer(1))
}

current_lm_counts <- risk_set_lm_counts(current_stack)
stopifnot(length(current_lm_counts) > 0L,
          all(current_lm_counts == 1L))
knitr::kable(data.frame(
  grid = "Current entry-age quartiles",
  event_times = length(current_lm_counts),
  event_times_with_multiple_LMs = sum(current_lm_counts > 1L)
))
```

# Dense-grid risk sets

I used the same strict-entry rule, event coding, prediction window, and
laboratory transformations as the primary attained-age analysis. I checked
the number of people and both event types in every window before fitting or
resampling. These counts are stacked-window counts: an event can contribute
to more than one overlapping window.

```{r dense-risk-sets}
dense_stack_raw <- stack_data(
  pbc_age, list(time = "exit_age", status = "status"),
  lms = dense_age_landmarks, w = prediction_window,
  covs = list(fixed = c("male", "stage", "trt"),
              varying = c("log_bili_z", "albumin_z")),
  format = "long", id = "id", rtime = "visit_age",
  entry = "entry_age", entry_mode = "strict"
)

dense_risk_sets <- do.call(rbind, lapply(dense_age_landmarks, function(age) {
  rows <- dense_stack_raw$data[dense_stack_raw$data$LM == age, , drop = FALSE]
  data.frame(
    landmark = age,
    window_end = age + prediction_window,
    subjects = length(unique(rows$id)),
    deaths = sum(rows$status == 2L),
    transplants = sum(rows$status == 1L),
    censored = sum(rows$status == 0L)
  )
}))

dense_lm_counts <- risk_set_lm_counts(dense_stack_raw)
stopifnot(
  identical(dense_stack_raw$entry_mode, "strict"),
  all(dense_stack_raw$data$entry_age <= dense_stack_raw$data$LM),
  all(dense_stack_raw$data$LM < dense_stack_raw$data$exit_age),
  !anyDuplicated(dense_stack_raw$data[c("id", "LM")]),
  all(dense_risk_sets$subjects > 0L),
  all(dense_risk_sets$subjects ==
        dense_risk_sets$deaths + dense_risk_sets$transplants +
        dense_risk_sets$censored),
  any(dense_lm_counts > 1L)
)

knitr::kable(dense_risk_sets,
             caption = "Strict-entry risk sets on the candidate age grid")
knitr::kable(data.frame(
  grid = "Candidate two-year grid",
  event_times = length(dense_lm_counts),
  event_times_with_multiple_LMs = sum(dense_lm_counts > 1L),
  maximum_LMs_in_one_risk_set = max(dense_lm_counts)
))
```

The dense windows overlap, and the observed event risk sets contain multiple
landmark values. This removes the exact non-identifiability of LM main effects
seen on the original grid. It does not by itself establish that the
competing-risk fit or its bootstrap is stable: the event-count table includes
windows with very few transplants.

# Candidate supermodel fit

I used a centered quadratic landmark basis for both main effects and the two
laboratory interactions. The other predictors and the cause-specific CSC fit
match the primary attained-age specification.

```{r dense-supermodel}
dense_center <- mean(dense_age_landmarks)
dense_half_range <- diff(range(dense_age_landmarks)) / 2
dense_basis <- list(
  function(age) (age - dense_center) / dense_half_range,
  function(age) ((age - dense_center) / dense_half_range)^2
)
dense_stack <- add_interactions(
  dense_stack_raw,
  lm_covs = c("log_bili_z", "albumin_z"),
  func_covars = dense_basis,
  func_lms = dense_basis
)
stopifnot(all(c("LM1", "LM2", "log_bili_z_LM1", "log_bili_z_LM2",
                "albumin_z_LM1", "albumin_z_LM2") %in%
              names(dense_stack$data)))

dense_formula <- Hist(exit_age, status, LM) ~
  male + stage + trt +
  log_bili_z + log_bili_z_LM1 + log_bili_z_LM2 +
  albumin_z + albumin_z_LM1 + albumin_z_LM2 +
  LM1 + LM2 + cluster(id)

capture_conditions <- function(expression) {
  warnings <- character()
  value <- withCallingHandlers(
    tryCatch(eval.parent(substitute(expression)), error = function(error) error),
    warning = function(warning) {
      warnings <<- c(warnings, conditionMessage(warning))
      invokeRestart("muffleWarning")
    }
  )
  list(value = value, warnings = unique(warnings))
}

dense_fit_record <- capture_conditions(dynamic_lm(
  dense_stack, dense_formula, type = "CSC", method = "breslow", x = TRUE
))

fit_status <- data.frame(
  error = if (inherits(dense_fit_record$value, "error")) {
    conditionMessage(dense_fit_record$value)
  } else "",
  warnings = paste(dense_fit_record$warnings, collapse = " | ")
)
knitr::kable(fit_status)

if (!inherits(dense_fit_record$value, "error")) {
  dense_fit <- dense_fit_record$value
  dense_lm_inference <- do.call(rbind, lapply(seq_along(
    dense_fit$model$causes
  ), function(index) {
    fit <- dense_fit$model$models[[index]]
    terms <- c("LM1", "LM2")
    data.frame(
      cause = cause_labels[as.integer(dense_fit$model$causes[index])],
      term = terms,
      coefficient = unname(coef(fit)[terms]),
      robust_se = unname(sqrt(diag(vcov(fit)))[terms])
    )
  }))
  stopifnot(all(is.finite(dense_lm_inference$coefficient)),
            all(is.finite(dense_lm_inference$robust_se)))
  knitr::kable(dense_lm_inference, digits = 3)
}
```

## Linear landmark sensitivity

I fitted a second pooled model on the same stack and with the same predictors,
keeping only the centered linear landmark main effect and linear laboratory
interactions. The quadratic model above remains the reference analysis.

```{r linear-supermodel}
linear_basis <- dense_basis[1L]
linear_stack <- add_interactions(
  dense_stack_raw,
  lm_covs = c("log_bili_z", "albumin_z"),
  func_covars = linear_basis,
  func_lms = linear_basis
)
stopifnot(
  all(c("LM1", "log_bili_z_LM1", "albumin_z_LM1") %in%
        names(linear_stack$data)),
  !any(c("LM2", "log_bili_z_LM2", "albumin_z_LM2") %in%
         names(linear_stack$data)),
  identical(dense_stack_raw$data[c("id", "LM")],
            linear_stack$data[c("id", "LM")])
)

linear_formula <- Hist(exit_age, status, LM) ~
  male + stage + trt +
  log_bili_z + log_bili_z_LM1 +
  albumin_z + albumin_z_LM1 + LM1 + cluster(id)
linear_fit_record <- capture_conditions(dynamic_lm(
  linear_stack, linear_formula, type = "CSC", method = "breslow", x = TRUE
))
linear_fit_status <- data.frame(
  error = if (inherits(linear_fit_record$value, "error")) {
    conditionMessage(linear_fit_record$value)
  } else "",
  warnings = paste(linear_fit_record$warnings, collapse = " | ")
)
knitr::kable(linear_fit_status)
stopifnot(!inherits(linear_fit_record$value, "error"))
linear_fit <- linear_fit_record$value
linear_death_fit <- linear_fit$model$models[[
  match(2L, linear_fit$model$causes)
]]
linear_robust_se <- sqrt(diag(vcov(linear_death_fit)))
stopifnot(
  all(is.finite(coef(linear_death_fit))),
  all(is.finite(linear_robust_se)),
  max(abs(linear_robust_se -
            summary(linear_death_fit)$coefficients[, "robust se"])) < 1e-12
)
```

# Separate landmark fits as a sparsity check

I also tried the unchanged, one-model-per-landmark specification. I recorded
all fit warnings and errors rather than modifying a sparse fit to make it
converge. The pooled supermodel can be estimable even when a separate
competing-risk fit at one age is not.

```{r dense-disjoint-fit-check}
dense_disjoint_formula <- Hist(exit_age, status, LM) ~
  male + stage + trt + log_bili_z + albumin_z + cluster(id)

dense_disjoint_records <- setNames(lapply(
  dense_age_landmarks, function(age) {
    one_landmark <- dense_stack_raw
    one_landmark$data <- dense_stack_raw$data[
      dense_stack_raw$data$LM == age, , drop = FALSE
    ]
    one_landmark$end_time <- age
    capture_conditions(dynamic_lm(
      one_landmark, dense_disjoint_formula,
      type = "CSC", method = "breslow", x = TRUE
    ))
  }), as.character(dense_age_landmarks)
)
dense_disjoint_conditions <- do.call(rbind, lapply(
  dense_age_landmarks, function(age) {
    fit_record <- dense_disjoint_records[[as.character(age)]]
    data.frame(landmark = age,
               fit_error = if (inherits(fit_record$value, "error")) {
                 conditionMessage(fit_record$value)
               } else "",
               warnings = paste(fit_record$warnings, collapse = " | "))
  }))

stopifnot(nrow(dense_disjoint_conditions) == length(dense_age_landmarks))
knitr::kable(dense_disjoint_conditions)
```

I traced each flagged fit back to its two cause-specific Cox models. The
direct fits use the same counting-process response, predictors, tie method,
and subject clustering as the separate CSC fit. I kept the CSC wrapper's
message alongside the cause-specific messages, since a zero-event cause can
make the wrapper fail without making `coxph()` itself throw an error.

```{r disjoint-warning-attribution}
flagged_ages <- dense_disjoint_conditions$landmark[
  nzchar(dense_disjoint_conditions$fit_error) |
    nzchar(dense_disjoint_conditions$warnings)
]
disjoint_cause_conditions <- do.call(rbind, lapply(flagged_ages, function(age) {
  rows <- dense_stack_raw$data[dense_stack_raw$data$LM == age, , drop = FALSE]
  outer <- dense_disjoint_records[[as.character(age)]]
  direct_records <- lapply(c(1L, 2L), function(cause) {
    formula <- as.formula(paste0(
      "survival::Surv(LM, exit_age, status == ", cause, ") ~ ",
      "male + stage + trt + log_bili_z + albumin_z + cluster(id)"
    ))
    capture_conditions(survival::coxph(
      formula, data = rows, method = "breslow", x = TRUE
    ))
  })
  if (!inherits(outer$value, "error")) {
    direct_warnings <- unique(unlist(lapply(direct_records, `[[`, "warnings")))
    stopifnot(setequal(direct_warnings, outer$warnings))
    for (index in seq_along(direct_records)) {
      direct <- direct_records[[index]]$value
      wrapped <- outer$value$model$models[[
        match(index, outer$value$model$causes)
      ]]
      if (!inherits(direct, "error")) {
        terms <- intersect(names(coef(direct)), names(coef(wrapped)))
        finite <- terms[is.finite(coef(direct)[terms]) &
                          is.finite(coef(wrapped)[terms])]
        if (length(finite)) {
          stopifnot(max(abs(coef(direct)[finite] -
                            coef(wrapped)[finite])) < 1e-8)
        }
      }
    }
  }
  cause_rows <- do.call(rbind, lapply(seq_along(direct_records), function(index) {
    record <- direct_records[[index]]
    data.frame(
      landmark = age,
      source = cause_labels[index],
      events = sum(rows$status == index),
      finite_coefficients = if (inherits(record$value, "error")) NA_integer_ else
        sum(is.finite(coef(record$value))),
      warning = paste(record$warnings, collapse = " | "),
      error = if (inherits(record$value, "error")) {
        conditionMessage(record$value)
      } else "",
      row.names = NULL
    )
  }))
  if (inherits(outer$value, "error")) {
    cause_rows <- rbind(cause_rows, data.frame(
      landmark = age, source = "CSC wrapper", events = NA_integer_,
      finite_coefficients = NA_integer_,
      warning = paste(outer$warnings, collapse = " | "),
      error = conditionMessage(outer$value), row.names = NULL
    ))
  }
  cause_rows
}))
knitr::kable(disjoint_cause_conditions,
             caption = "Exact conditions by cause; the CSC wrapper row is shown when it fails")
```

# Death-model laboratory hazard ratios

I combined each laboratory main effect with its landmark interactions and
used the fitted robust covariance for the supermodel interval. For each
separate landmark fit, I used the death-model coefficient and robust standard
error. CI-width ratios are on the log-HR scale, as in the primary comparison.
An unavailable separate fit remains unavailable in the table and figure.

```{r dense-hazard-ratio-table}
death_index <- match(2L, dense_fit$model$causes)
stopifnot(length(death_index) == 1L, !is.na(death_index))
dense_death_fit <- dense_fit$model$models[[death_index]]
robust_se <- sqrt(diag(vcov(dense_death_fit)))
summary_se <- summary(dense_death_fit)$coefficients[, "robust se"]
stopifnot(identical(names(robust_se), names(coef(dense_death_fit))),
          max(abs(robust_se - summary_se), na.rm = TRUE) < 1e-12)

critical_value <- qnorm(0.975)
laboratory_terms <- c(log_bili_z = "Log bilirubin (per SD)",
                      albumin_z = "Albumin (per SD)")

supermodel_log_hr <- function(marker, ages) {
  terms <- c(marker, paste0(marker, c("_LM1", "_LM2")))
  design <- cbind(1, dense_basis[[1]](ages), dense_basis[[2]](ages))
  beta <- coef(dense_death_fit)[terms]
  covariance <- vcov(dense_death_fit)[terms, terms, drop = FALSE]
  estimate <- drop(design %*% beta)
  se <- sqrt(rowSums((design %*% covariance) * design))
  stopifnot(all(is.finite(estimate)), all(is.finite(se)))
  data.frame(landmark = ages, estimate = estimate,
             lower = estimate - critical_value * se,
             upper = estimate + critical_value * se)
}

linear_log_hr <- function(marker, ages) {
  terms <- c(marker, paste0(marker, "_LM1"))
  design <- cbind(1, linear_basis[[1L]](ages))
  beta <- coef(linear_death_fit)[terms]
  covariance <- vcov(linear_death_fit)[terms, terms, drop = FALSE]
  estimate <- drop(design %*% beta)
  se <- sqrt(rowSums((design %*% covariance) * design))
  stopifnot(all(is.finite(estimate)), all(is.finite(se)))
  data.frame(landmark = ages, estimate = estimate,
             lower = estimate - critical_value * se,
             upper = estimate + critical_value * se)
}

disjoint_log_hr <- function(age, marker) {
  record <- dense_disjoint_records[[as.character(age)]]
  if (inherits(record$value, "error")) {
    return(c(estimate = NA_real_, lower = NA_real_, upper = NA_real_))
  }
  cause_index <- match(2L, record$value$model$causes)
  stopifnot(!is.na(cause_index))
  fit <- record$value$model$models[[cause_index]]
  estimate <- coef(fit)[marker]
  se <- sqrt(diag(vcov(fit)))[marker]
  if (!is.finite(estimate) || !is.finite(se)) {
    return(c(estimate = NA_real_, lower = NA_real_, upper = NA_real_))
  }
  c(estimate = estimate,
    lower = estimate - critical_value * se,
    upper = estimate + critical_value * se)
}

laboratory_comparison <- do.call(rbind, lapply(names(laboratory_terms),
  function(marker) {
    super <- supermodel_log_hr(marker, dense_age_landmarks)
    linear <- linear_log_hr(marker, dense_age_landmarks)
    separate <- t(vapply(dense_age_landmarks,
      function(age) disjoint_log_hr(age, marker),
      c(estimate = 0, lower = 0, upper = 0)))
    conditions <- dense_disjoint_conditions[
      match(dense_age_landmarks, dense_disjoint_conditions$landmark), ]
    data.frame(
      covariate = unname(laboratory_terms[marker]),
      landmark = dense_age_landmarks,
      super_estimate = super$estimate,
      super_lower = super$lower,
      super_upper = super$upper,
      linear_estimate = linear$estimate,
      linear_lower = linear$lower,
      linear_upper = linear$upper,
      disjoint_estimate = separate[, "estimate"],
      disjoint_lower = separate[, "lower"],
      disjoint_upper = separate[, "upper"],
      CI_width_ratio_log_scale =
        (separate[, "upper"] - separate[, "lower"]) /
        (super$upper - super$lower),
      fit_flag = ifelse(nzchar(conditions$fit_error), "failed",
                        ifelse(nzchar(conditions$warnings), "warning", "ok")),
      row.names = NULL
    )
  }
))

format_hr <- function(estimate, lower, upper) {
  ifelse(is.finite(estimate) & is.finite(lower) & is.finite(upper),
         sprintf("%.3f (%.3f, %.3f)",
                 exp(estimate), exp(lower), exp(upper)), "—")
}
laboratory_display <- data.frame(
  covariate = laboratory_comparison$covariate,
  landmark = laboratory_comparison$landmark,
  supermodel_HR_95_CI = with(laboratory_comparison,
    format_hr(super_estimate, super_lower, super_upper)),
  linear_HR_95_CI = with(laboratory_comparison,
    format_hr(linear_estimate, linear_lower, linear_upper)),
  disjoint_HR_95_CI = with(laboratory_comparison,
    format_hr(disjoint_estimate, disjoint_lower, disjoint_upper)),
  disjoint_to_supermodel_CI_width =
    laboratory_comparison$CI_width_ratio_log_scale,
  fit_flag = laboratory_comparison$fit_flag
)
stopifnot(nrow(laboratory_display) ==
            length(dense_age_landmarks) * length(laboratory_terms))
knitr::kable(laboratory_display, digits = 2)

finite_comparisons <- laboratory_comparison[
  is.finite(laboratory_comparison$disjoint_estimate), ]
different_at_each_landmark <- all(abs(
  finite_comparisons$super_estimate - finite_comparisons$disjoint_estimate
) > 1e-8)
cat(if (different_at_each_landmark) {
  "The quadratic supermodel now smooths: its laboratory HRs do not reproduce the separate-model HRs at any estimable landmark."
} else {
  "The quadratic supermodel differs from the separate models at some, but not all, estimable landmarks."
})

condition_rows <- dense_disjoint_conditions[
  nzchar(dense_disjoint_conditions$fit_error) |
    nzchar(dense_disjoint_conditions$warnings), ]
condition_text <- vapply(seq_len(nrow(condition_rows)), function(index) {
  row <- condition_rows[index, ]
  paste0("age ", row$landmark, ": ",
         paste(c(row$fit_error, row$warnings)[
           nzchar(c(row$fit_error, row$warnings))], collapse = " | "))
}, character(1))
hr_figure_caption <- paste(
  "Death-model HR per SD of the laboratory measure; solid line and grey band: quadratic supermodel and robust 95% CI; dashed line: linear supermodel; point bars: disjoint robust 95% CIs.",
  "Filled points: unflagged separate fits; open points: warned fits; crosses at the lower axis: failed fits.",
  "Recorded conditions:", paste(condition_text, collapse = "; ")
)
```

```{r dense-hazard-ratio-plot, fig.width=10, fig.height=5, fig.cap=hr_figure_caption}
plot_laboratory_hr <- function(marker, label) {
  ages <- seq(min(dense_age_landmarks), max(dense_age_landmarks),
              length.out = 200L)
  curve <- supermodel_log_hr(marker, ages)
  linear_curve <- linear_log_hr(marker, ages)
  rows <- laboratory_comparison[
    laboratory_comparison$covariate == label, ]
  y_values <- exp(c(curve$lower, curve$upper, linear_curve$estimate,
                    rows$disjoint_lower, rows$disjoint_upper, 0))
  y_limits <- range(y_values[is.finite(y_values) & y_values > 0])
  plot(ages, exp(curve$estimate), type = "n", log = "y",
       ylim = y_limits, xlab = "Attained age (years)",
       ylab = "Hazard ratio (log scale)", main = label, bty = "l")
  polygon(c(ages, rev(ages)),
          c(exp(curve$lower), rev(exp(curve$upper))),
          col = "grey85", border = NA)
  abline(h = 1, lty = 2, col = "grey45")
  lines(ages, exp(curve$estimate), lwd = 2)
  lines(ages, exp(linear_curve$estimate), lwd = 2, lty = 2)
  finite <- is.finite(rows$disjoint_estimate) &
    is.finite(rows$disjoint_lower) & is.finite(rows$disjoint_upper)
  arrows(rows$landmark[finite], exp(rows$disjoint_lower[finite]),
         rows$landmark[finite], exp(rows$disjoint_upper[finite]),
         angle = 90, code = 3, length = 0.04)
  points(rows$landmark[finite], exp(rows$disjoint_estimate[finite]),
         pch = ifelse(rows$fit_flag[finite] == "warning", 21, 19),
         bg = "white")
  failed <- rows$fit_flag == "failed" | !finite
  if (any(failed)) {
    points(rows$landmark[failed], y_limits[1] * 1.05, pch = 4, cex = 1.2)
  }
}

old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
for (marker in names(laboratory_terms)) {
  plot_laboratory_hr(marker, unname(laboratory_terms[marker]))
}
par(old_par)
```

# Supermodel prediction and bootstrap

I kept the disjoint fit failures above as findings. For prediction I evaluated
the pooled supermodel alone. The bootstrap resampled people rather than
stacked rows, assigned a new ID to each resampled copy, rebuilt the strict
age stack, refitted both cause-specific hazards, and scored the training and
original cohorts at each landmark. The grid, window, transformations, and
seed were fixed before resampling.

```{r dense-bootstrap-functions}
subset_prediction <- function(prediction, age) {
  rows <- prediction$data$LM == age
  if (!any(rows)) stop("No predicted rows at age ", age)
  out <- prediction
  out$data <- prediction$data[rows, , drop = FALSE]
  out$preds <- prediction$preds[rows, , drop = FALSE]
  stopifnot(nrow(out$data) == nrow(out$preds))
  out
}

score_prediction <- function(prediction, age) {
  if (anyNA(prediction$preds$risk)) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = "NA predicted risk"))
  }
  record <- capture_conditions(score(
    list(Model = prediction), times = age, metrics = c("auc", "brier"),
    id_col = "id", cause = 2L, summary = FALSE, contrasts = FALSE
  ))
  if (inherits(record$value, "error")) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = conditionMessage(record$value)))
  }
  auc <- as.data.frame(record$value$AUC$score)
  brier <- as.data.frame(record$value$Brier$score)
  auc <- auc[as.character(auc$model) == "Model", , drop = FALSE]
  brier <- brier[as.character(brier$model) == "Model", , drop = FALSE]
  if (nrow(auc) != 1L || nrow(brier) != 1L ||
      !is.finite(auc$AUC) || !is.finite(brier$Brier)) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = "non-finite or missing score"))
  }
  list(AUC = auc$AUC, Brier = brier$Brier, error = "")
}

resample_people <- function(data, sampled_ids) {
  copies <- lapply(seq_along(sampled_ids), function(index) {
    rows <- data[data$id == sampled_ids[index], , drop = FALSE]
    stopifnot(nrow(rows) > 0L)
    rows$id <- index
    rows
  })
  sample_data <- do.call(rbind, copies)
  rownames(sample_data) <- NULL
  stopifnot(length(unique(sample_data$id)) == length(sampled_ids))
  sample_data
}

build_dense_stack <- function(data) {
  raw <- stack_data(
    data, list(time = "exit_age", status = "status"),
    lms = dense_age_landmarks, w = prediction_window,
    covs = list(fixed = c("male", "stage", "trt"),
                varying = c("log_bili_z", "albumin_z")),
    format = "long", id = "id", rtime = "visit_age",
    entry = "entry_age", entry_mode = "strict"
  )
  stopifnot(!anyDuplicated(raw$data[c("id", "LM")]),
            all(raw$data$entry_age <= raw$data$LM))
  model <- add_interactions(
    raw, lm_covs = c("log_bili_z", "albumin_z"),
    func_covars = dense_basis, func_lms = dense_basis
  )
  stopifnot(!anyDuplicated(model$data[c("id", "LM")]),
            identical(raw$data[c("id", "LM")], model$data[c("id", "LM")]))
  model
}

empty_bootstrap_rows <- function(replicate) {
  data.frame(
    replicate = replicate, landmark = dense_age_landmarks,
    zero_transplants = NA, zero_deaths = NA,
    transplant_fit_error = "", transplant_fit_warning = "",
    death_fit_error = "", death_fit_warning = "",
    nonfinite_cause = "", pooled_fit_warning = "",
    train_AUC = NA_real_, test_AUC = NA_real_,
    train_Brier = NA_real_, test_Brier = NA_real_,
    failed_step = "", failure_detail = "", complete = FALSE
  )
}

diagnose_cause <- function(stack, cause) {
  formula <- as.formula(paste0(
    "survival::Surv(LM, exit_age, status == ", cause, ") ~ ",
    "male + stage + trt + log_bili_z + log_bili_z_LM1 + ",
    "log_bili_z_LM2 + albumin_z + albumin_z_LM1 + ",
    "albumin_z_LM2 + LM1 + LM2 + cluster(id)"
  ))
  record <- capture_conditions(survival::coxph(
    formula, data = stack$data, method = "breslow", x = TRUE
  ))
  list(
    error = if (inherits(record$value, "error")) {
      conditionMessage(record$value)
    } else "",
    warning = paste(record$warnings, collapse = " | ")
  )
}

original_prediction_record <- capture_conditions(predict(
  dense_fit, newdata = dense_stack, cause = 2L
))
stopifnot(!inherits(original_prediction_record$value, "error"))
original_prediction <- original_prediction_record$value
apparent_scores <- do.call(rbind, lapply(dense_age_landmarks, function(age) {
  metric <- score_prediction(subset_prediction(original_prediction, age), age)
  data.frame(landmark = age, AUC = metric$AUC, Brier = metric$Brier,
             score_error = metric$error)
}))
stopifnot(all(is.finite(apparent_scores$AUC)),
          all(is.finite(apparent_scores$Brier)),
          !any(nzchar(apparent_scores$score_error)))
knitr::kable(apparent_scores, digits = 4)
```

```{r dense-bootstrap, cache=FALSE}
set.seed(bootstrap_seed)
original_ids <- unique(pbc_age$id)
bootstrap_samples <- lapply(seq_len(n_boot), function(index) {
  sample(original_ids, length(original_ids), replace = TRUE)
})

run_dense_replicate <- function(index) {
  row <- empty_bootstrap_rows(index)
  sampled <- capture_conditions(resample_people(
    pbc_age, bootstrap_samples[[index]]
  ))
  if (inherits(sampled$value, "error")) {
    row$failed_step <- "resample"
    row$failure_detail <- conditionMessage(sampled$value)
    return(row)
  }
  stack_record <- capture_conditions(build_dense_stack(sampled$value))
  if (inherits(stack_record$value, "error")) {
    row$failed_step <- "stack"
    row$failure_detail <- conditionMessage(stack_record$value)
    return(row)
  }
  training_stack <- stack_record$value
  for (age in dense_age_landmarks) {
    at_age <- training_stack$data$LM == age
    row$zero_transplants[row$landmark == age] <-
      sum(training_stack$data$status[at_age] == 1L) == 0L
    row$zero_deaths[row$landmark == age] <-
      sum(training_stack$data$status[at_age] == 2L) == 0L
  }

  fit_record <- capture_conditions(dynamic_lm(
    training_stack, dense_formula, type = "CSC",
    method = "breslow", x = TRUE
  ))
  row$pooled_fit_warning <- paste(fit_record$warnings, collapse = " | ")
  if (inherits(fit_record$value, "error") || length(fit_record$warnings)) {
    transplant <- diagnose_cause(training_stack, 1L)
    death <- diagnose_cause(training_stack, 2L)
    row$transplant_fit_error <- transplant$error
    row$transplant_fit_warning <- transplant$warning
    row$death_fit_error <- death$error
    row$death_fit_warning <- death$warning
  }
  if (inherits(fit_record$value, "error")) {
    row$failed_step <- "pooled fit"
    row$failure_detail <- conditionMessage(fit_record$value)
    return(row)
  }

  nonfinite <- vapply(seq_along(fit_record$value$model$causes),
    function(cause_index) {
      fit <- fit_record$value$model$models[[cause_index]]
      any(!is.finite(coef(fit)))
    }, logical(1))
  row$nonfinite_cause <- paste(
    fit_record$value$model$causes[nonfinite], collapse = ", "
  )

  train_prediction <- capture_conditions(predict(
    fit_record$value, newdata = training_stack, cause = 2L
  ))
  test_prediction <- capture_conditions(predict(
    fit_record$value, newdata = dense_stack, cause = 2L
  ))
  for (age in dense_age_landmarks) {
    at_age <- which(row$landmark == age)
    if (inherits(train_prediction$value, "error")) {
      row$failed_step[at_age] <- "train prediction"
      row$failure_detail[at_age] <- conditionMessage(train_prediction$value)
      next
    }
    if (inherits(test_prediction$value, "error")) {
      row$failed_step[at_age] <- "test prediction"
      row$failure_detail[at_age] <- conditionMessage(test_prediction$value)
      next
    }
    train_subset <- capture_conditions(subset_prediction(
      train_prediction$value, age
    ))
    test_subset <- capture_conditions(subset_prediction(
      test_prediction$value, age
    ))
    if (inherits(train_subset$value, "error")) {
      row$failed_step[at_age] <- "train prediction subset"
      row$failure_detail[at_age] <- conditionMessage(train_subset$value)
      next
    }
    if (inherits(test_subset$value, "error")) {
      row$failed_step[at_age] <- "test prediction subset"
      row$failure_detail[at_age] <- conditionMessage(test_subset$value)
      next
    }
    train <- score_prediction(train_subset$value, age)
    test <- score_prediction(test_subset$value, age)
    row$train_AUC[at_age] <- train$AUC
    row$test_AUC[at_age] <- test$AUC
    row$train_Brier[at_age] <- train$Brier
    row$test_Brier[at_age] <- test$Brier
    if (nzchar(train$error)) {
      row$failed_step[at_age] <- if (train$error == "NA predicted risk") {
        "train prediction"
      } else "train score"
      row$failure_detail[at_age] <- train$error
    } else if (nzchar(test$error)) {
      row$failed_step[at_age] <- if (test$error == "NA predicted risk") {
        "test prediction"
      } else "test score"
      row$failure_detail[at_age] <- test$error
    } else {
      row$complete[at_age] <- TRUE
    }
  }
  row
}

safe_dense_replicate <- function(index) {
  tryCatch(run_dense_replicate(index), error = function(error) {
    row <- empty_bootstrap_rows(index)
    row$failed_step <- "unexpected error"
    row$failure_detail <- conditionMessage(error)
    row
  })
}

bootstrap_cores <- getOption(
  "timescale.bootstrap.cores",
  if (.Platform$OS.type == "windows") 1L else
    min(4L, parallel::detectCores(logical = FALSE))
)
bootstrap_rows <- if (.Platform$OS.type == "windows" ||
                      bootstrap_cores <= 1L) {
  lapply(seq_len(n_boot), safe_dense_replicate)
} else {
  parallel::mclapply(seq_len(n_boot), safe_dense_replicate,
                     mc.cores = bootstrap_cores, mc.preschedule = FALSE)
}
bootstrap_diagnostics <- do.call(rbind, bootstrap_rows)
rownames(bootstrap_diagnostics) <- NULL
finite_scores <- with(bootstrap_diagnostics,
  is.finite(train_AUC) & is.finite(test_AUC) &
    is.finite(train_Brier) & is.finite(test_Brier))
if (!identical(bootstrap_diagnostics$complete, finite_scores)) {
  print(bootstrap_diagnostics[
    bootstrap_diagnostics$complete != finite_scores,
    c("replicate", "landmark", "complete", "failed_step",
      "train_AUC", "test_AUC", "train_Brier", "test_Brier")
  ])
}
stopifnot(
  nrow(bootstrap_diagnostics) == n_boot * length(dense_age_landmarks),
  !anyDuplicated(bootstrap_diagnostics[c("replicate", "landmark")]),
  identical(bootstrap_diagnostics$complete, finite_scores),
  all(nzchar(bootstrap_diagnostics$failed_step[!bootstrap_diagnostics$complete]))
)
```

I separated zero-transplant windows from actual failures: a zero count in one
window does not necessarily invalidate a pooled fit. The reason table gives
the first step that prevented a complete train/test score, retaining the
error text. Cause-specific fit conditions are shown separately.

```{r dense-bootstrap-diagnostics}
bootstrap_completion <- do.call(rbind, lapply(dense_age_landmarks,
  function(age) {
    rows <- bootstrap_diagnostics[
      bootstrap_diagnostics$landmark == age, ]
    data.frame(
      landmark = age,
      complete_score_pairs = sum(rows$complete),
      incomplete_score_pairs = sum(!rows$complete),
      complete_with_fit_warning = sum(
        rows$complete & nzchar(rows$pooled_fit_warning)),
      zero_transplant_resamples = sum(rows$zero_transplants, na.rm = TRUE),
      zero_death_resamples = sum(rows$zero_deaths, na.rm = TRUE)
    )
  }
))
complete_everywhere <- tapply(bootstrap_diagnostics$complete,
                              bootstrap_diagnostics$replicate, all)
stopifnot(length(complete_everywhere) == n_boot,
          all(bootstrap_completion$complete_score_pairs +
                bootstrap_completion$incomplete_score_pairs == n_boot))

failure_rows <- bootstrap_diagnostics[!bootstrap_diagnostics$complete, ]
failure_reasons <- if (nrow(failure_rows)) {
  aggregate(replicate ~ landmark + failed_step + failure_detail,
            data = failure_rows, FUN = length)
} else data.frame()
if (nrow(failure_reasons)) {
  names(failure_reasons)[names(failure_reasons) == "replicate"] <- "resamples"
  failure_reasons <- failure_reasons[order(
    failure_reasons$landmark, failure_reasons$failed_step,
    failure_reasons$failure_detail), ]
  rownames(failure_reasons) <- NULL
}

cause_conditions <- unique(bootstrap_diagnostics[c(
  "replicate", "transplant_fit_error", "transplant_fit_warning",
  "death_fit_error", "death_fit_warning", "nonfinite_cause",
  "pooled_fit_warning"
)])
cause_condition_counts <- data.frame(
  condition = c("Transplant fit error", "Transplant fit warning",
                "Death fit error", "Death fit warning",
                "Non-finite cause coefficient", "Pooled fit warning"),
  resamples = c(
    sum(nzchar(cause_conditions$transplant_fit_error)),
    sum(nzchar(cause_conditions$transplant_fit_warning)),
    sum(nzchar(cause_conditions$death_fit_error)),
    sum(nzchar(cause_conditions$death_fit_warning)),
    sum(nzchar(cause_conditions$nonfinite_cause)),
    sum(nzchar(cause_conditions$pooled_fit_warning))
  )
)

knitr::kable(bootstrap_completion)
knitr::kable(data.frame(
  bootstrap_resamples = n_boot,
  complete_at_every_landmark = sum(complete_everywhere)
))
if (nrow(failure_reasons)) knitr::kable(failure_reasons)
knitr::kable(cause_condition_counts)
```

At age `r dense_age_landmarks[which.max(bootstrap_completion$zero_transplant_resamples)]`,
`r max(bootstrap_completion$zero_transplant_resamples)` resamples had no
transplants in that window, yet `r bootstrap_completion$complete_score_pairs[which.max(bootstrap_completion$zero_transplant_resamples)]`
still yielded complete pooled-model scores. Across all landmarks,
`r sum(complete_everywhere)` resamples were complete throughout. Finite scores
from a warned cause-specific fit remain flagged; I did not count a warning
alone as a scoring failure.

## Apparent and optimism-corrected performance

I subtracted the mean training-minus-original optimism from each apparent
landmark score. When not every resample had finite training and test scores,
the corrected value is conditional on the complete resamples at that
landmark; I do not treat it as an unconditional estimate.

```{r dense-performance-table}
dense_performance <- do.call(rbind, lapply(dense_age_landmarks,
  function(age) {
    apparent <- apparent_scores[apparent_scores$landmark == age, ]
    counts <- dense_risk_sets[dense_risk_sets$landmark == age, ]
    paired <- bootstrap_diagnostics[
      bootstrap_diagnostics$landmark == age &
        bootstrap_diagnostics$complete, ]
    stopifnot(nrow(apparent) == 1L, nrow(counts) == 1L)
    data.frame(
      landmark = age,
      deaths = counts$deaths,
      transplants = counts$transplants,
      apparent_AUC = apparent$AUC,
      optimism_corrected_AUC = if (nrow(paired)) {
        apparent$AUC - mean(paired$train_AUC - paired$test_AUC)
      } else NA_real_,
      apparent_Brier = apparent$Brier,
      optimism_corrected_Brier = if (nrow(paired)) {
        apparent$Brier - mean(paired$train_Brier - paired$test_Brier)
      } else NA_real_,
      complete_score_pairs = nrow(paired),
      interpretation = if (nrow(paired) == n_boot) {
        "all resamples"
      } else "conditional on complete resamples"
    )
  }
))
stopifnot(identical(dense_performance$complete_score_pairs,
                    bootstrap_completion$complete_score_pairs))
knitr::kable(dense_performance, digits = 4)
```

## Linear and separate-model prediction

I used the same `score()` call for all three models. The separate model at a
landmark was evaluated on that landmark's original risk set. I kept a row for
every landmark even when the separate CSC fit or its prediction failed.

The first separate fit had no warning, but its `male` coefficient was `NA` in
both cause models: every subject in that risk set had `male` equal to zero.
The other fitted coefficients were unchanged when I omitted `male` from a
direct cause-specific fit. Prediction nevertheless failed because an aliased
coefficient was passed through as missing, even though its contribution was
zero for every training and target row. I resolved only that zero-valued alias
at the first landmark; I left all other fit and prediction paths unchanged.

```{r additional-apparent-performance}
one_landmark_stack <- function(stack, age) {
  out <- stack
  out$data <- stack$data[stack$data$LM == age, , drop = FALSE]
  out$end_time <- age
  stopifnot(nrow(out$data) > 0L)
  out
}

first_dense_age <- min(dense_age_landmarks)
first_age_stack <- one_landmark_stack(dense_stack_raw, first_dense_age)
first_age_record <- dense_disjoint_records[[as.character(first_dense_age)]]
first_age_aliases <- do.call(rbind, lapply(seq_along(
  first_age_record$value$model$models
), function(index) {
  fit <- first_age_record$value$model$models[[index]]
  data.frame(
    cause = cause_labels[index],
    aliased_terms = paste(names(coef(fit))[is.na(coef(fit))], collapse = ", "),
    unique_male_values = paste(unique(first_age_stack$data$male),
                               collapse = ", ")
  )
}))
stopifnot(
  !inherits(first_age_record$value, "error"),
  !anyNA(first_age_stack$data$male),
  all(first_age_stack$data$male == 0),
  all(first_age_aliases$aliased_terms == "male")
)
knitr::kable(first_age_aliases,
             caption = "Unwarned zero-valued alias in the first separate fit")

first_age_omit_male <- do.call(rbind, lapply(seq_along(
  first_age_record$value$model$models
), function(index) {
  reduced <- survival::coxph(as.formula(paste0(
    "survival::Surv(LM, exit_age, status == ", index, ") ~ ",
    "stage + trt + log_bili_z + albumin_z + cluster(id)"
  )), data = first_age_stack$data, method = "breslow", x = TRUE)
  full <- first_age_record$value$model$models[[index]]
  shared_terms <- names(coef(reduced))
  data.frame(cause = cause_labels[index],
             largest_other_coefficient_difference = max(abs(
               coef(full)[shared_terms] - coef(reduced)
             )))
}))
stopifnot(all(first_age_omit_male$largest_other_coefficient_difference == 0))
knitr::kable(first_age_omit_male)

unrepaired_first_age_prediction <- capture_conditions(predict(
  first_age_record$value, newdata = first_age_stack, cause = 2L
))
stopifnot(inherits(unrepaired_first_age_prediction$value, "error"))
knitr::kable(data.frame(
  prediction_error = conditionMessage(unrepaired_first_age_prediction$value)
))

repair_zero_aliases <- function(record, training, original) {
  if (inherits(record$value, "error")) return(record)
  repaired <- character()
  for (index in seq_along(record$value$model$models)) {
    fit <- record$value$model$models[[index]]
    aliased <- names(coef(fit))[is.na(coef(fit))]
    zero_in_both <- vapply(aliased, function(term) {
      term %in% names(training$data) && term %in% names(original$data) &&
        !anyNA(training$data[[term]]) && !anyNA(original$data[[term]]) &&
        all(training$data[[term]] == 0) &&
        all(original$data[[term]] == 0)
    }, logical(1))
    if (any(zero_in_both)) {
      terms <- aliased[zero_in_both]
      record$value$model$models[[index]]$coefficients[terms] <- 0
      repaired <- c(repaired, paste(record$value$model$causes[index],
                                   terms, sep = ":"))
    }
  }
  record$zero_alias_terms <- paste(repaired, collapse = "; ")
  record
}

first_age_scoring_record <- repair_zero_aliases(
  first_age_record, first_age_stack, first_age_stack
)
stopifnot(setequal(strsplit(first_age_scoring_record$zero_alias_terms,
                            "; ", fixed = TRUE)[[1L]],
                   c("1:male", "2:male")))

score_original_fit <- function(record, target, age) {
  if (inherits(record$value, "error")) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = conditionMessage(record$value)))
  }
  prediction <- capture_conditions(predict(
    record$value, newdata = target, cause = 2L
  ))
  if (inherits(prediction$value, "error")) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = conditionMessage(prediction$value)))
  }
  one_age <- capture_conditions(subset_prediction(prediction$value, age))
  if (inherits(one_age$value, "error")) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = conditionMessage(one_age$value)))
  }
  target_rows <- target$data[target$data$LM == age, c("id", "LM")]
  predicted_rows <- one_age$value$data[c("id", "LM")]
  if (!identical(target_rows, predicted_rows)) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = "prediction did not retain the full landmark risk set"))
  }
  score_prediction(one_age$value, age)
}

linear_apparent <- do.call(rbind, lapply(dense_age_landmarks, function(age) {
  metric <- score_original_fit(linear_fit_record, linear_stack, age)
  data.frame(model = "Linear supermodel", landmark = age,
             AUC = metric$AUC, Brier = metric$Brier,
             error = metric$error)
}))
disjoint_apparent <- do.call(rbind, lapply(dense_age_landmarks, function(age) {
  record <- if (age == first_dense_age) {
    first_age_scoring_record
  } else dense_disjoint_records[[as.character(age)]]
  metric <- score_original_fit(
    record,
    one_landmark_stack(dense_stack_raw, age), age
  )
  data.frame(model = "Disjoint", landmark = age,
             AUC = metric$AUC, Brier = metric$Brier,
             error = metric$error)
}))
additional_apparent <- rbind(linear_apparent, disjoint_apparent)
knitr::kable(additional_apparent, digits = 4)
```

For optimism correction I reused the sampled subject IDs generated above.
No new seed or resampling rule was introduced. The two new fitting paths were
run separately from the quadratic path, so its stored scores and diagnostics
were not rewritten.

```{r additional-bootstrap-functions}
empty_additional_rows <- function(index, model, ages) {
  data.frame(
    replicate = index, model = model, landmark = ages,
    fit_warning = "", zero_alias_terms = "",
    train_AUC = NA_real_, test_AUC = NA_real_,
    train_Brier = NA_real_, test_Brier = NA_real_,
    failed_step = "", failure_detail = "", complete = FALSE
  )
}

score_fit_in_resample <- function(record, training, original, ages,
                                  index, model) {
  rows <- empty_additional_rows(index, model, ages)
  rows$fit_warning <- paste(record$warnings, collapse = " | ")
  if (!is.null(record$zero_alias_terms)) {
    rows$zero_alias_terms <- record$zero_alias_terms
  }
  if (inherits(record$value, "error")) {
    rows$failed_step <- "fit"
    rows$failure_detail <- conditionMessage(record$value)
    return(rows)
  }
  train_prediction <- capture_conditions(predict(
    record$value, newdata = training, cause = 2L
  ))
  test_prediction <- capture_conditions(predict(
    record$value, newdata = original, cause = 2L
  ))
  for (age in ages) {
    at_age <- which(rows$landmark == age)
    if (inherits(train_prediction$value, "error")) {
      rows$failed_step[at_age] <- "train prediction"
      rows$failure_detail[at_age] <- conditionMessage(train_prediction$value)
      next
    }
    if (inherits(test_prediction$value, "error")) {
      rows$failed_step[at_age] <- "test prediction"
      rows$failure_detail[at_age] <- conditionMessage(test_prediction$value)
      next
    }
    train_subset <- capture_conditions(subset_prediction(
      train_prediction$value, age
    ))
    test_subset <- capture_conditions(subset_prediction(
      test_prediction$value, age
    ))
    if (inherits(train_subset$value, "error")) {
      rows$failed_step[at_age] <- "train prediction subset"
      rows$failure_detail[at_age] <- conditionMessage(train_subset$value)
      next
    }
    if (inherits(test_subset$value, "error")) {
      rows$failed_step[at_age] <- "test prediction subset"
      rows$failure_detail[at_age] <- conditionMessage(test_subset$value)
      next
    }
    train <- score_prediction(train_subset$value, age)
    test <- score_prediction(test_subset$value, age)
    rows$train_AUC[at_age] <- train$AUC
    rows$test_AUC[at_age] <- test$AUC
    rows$train_Brier[at_age] <- train$Brier
    rows$test_Brier[at_age] <- test$Brier
    if (nzchar(train$error)) {
      rows$failed_step[at_age] <- if (train$error == "NA predicted risk") {
        "train prediction"
      } else "train score"
      rows$failure_detail[at_age] <- train$error
    } else if (nzchar(test$error)) {
      rows$failed_step[at_age] <- if (test$error == "NA predicted risk") {
        "test prediction"
      } else "test score"
      rows$failure_detail[at_age] <- test$error
    } else {
      rows$complete[at_age] <- TRUE
    }
  }
  rows
}

additional_stacks <- function(data) {
  raw <- stack_data(
    data, list(time = "exit_age", status = "status"),
    lms = dense_age_landmarks, w = prediction_window,
    covs = list(fixed = c("male", "stage", "trt"),
                varying = c("log_bili_z", "albumin_z")),
    format = "long", id = "id", rtime = "visit_age",
    entry = "entry_age", entry_mode = "strict"
  )
  stopifnot(!anyDuplicated(raw$data[c("id", "LM")]),
            all(raw$data$entry_age <= raw$data$LM))
  linear <- add_interactions(
    raw, lm_covs = c("log_bili_z", "albumin_z"),
    func_covars = linear_basis, func_lms = linear_basis
  )
  stopifnot(identical(raw$data[c("id", "LM")],
                      linear$data[c("id", "LM")]))
  list(raw = raw, linear = linear)
}

run_additional_replicate <- function(index) {
  missing_rows <- rbind(
    empty_additional_rows(index, "Linear supermodel", dense_age_landmarks),
    empty_additional_rows(index, "Disjoint", dense_age_landmarks)
  )
  sampled <- capture_conditions(resample_people(
    pbc_age, bootstrap_samples[[index]]
  ))
  if (inherits(sampled$value, "error")) {
    missing_rows$failed_step <- "resample"
    missing_rows$failure_detail <- conditionMessage(sampled$value)
    return(missing_rows)
  }
  stacks <- capture_conditions(additional_stacks(sampled$value))
  if (inherits(stacks$value, "error")) {
    missing_rows$failed_step <- "stack"
    missing_rows$failure_detail <- conditionMessage(stacks$value)
    return(missing_rows)
  }
  linear_record <- capture_conditions(dynamic_lm(
    stacks$value$linear, linear_formula,
    type = "CSC", method = "breslow", x = TRUE
  ))
  linear_rows <- score_fit_in_resample(
    linear_record, stacks$value$linear, linear_stack,
    dense_age_landmarks, index, "Linear supermodel"
  )
  separate_rows <- lapply(dense_age_landmarks, function(age) {
    training <- one_landmark_stack(stacks$value$raw, age)
    original <- one_landmark_stack(dense_stack_raw, age)
    record <- capture_conditions(dynamic_lm(
      training, dense_disjoint_formula,
      type = "CSC", method = "breslow", x = TRUE
    ))
    if (age == first_dense_age) {
      record <- repair_zero_aliases(record, training, original)
    }
    score_fit_in_resample(record, training, original,
                          age, index, "Disjoint")
  })
  rbind(linear_rows, do.call(rbind, separate_rows))
}

safe_additional_replicate <- function(index) {
  tryCatch(run_additional_replicate(index), error = function(error) {
    rows <- rbind(
      empty_additional_rows(index, "Linear supermodel", dense_age_landmarks),
      empty_additional_rows(index, "Disjoint", dense_age_landmarks)
    )
    rows$failed_step <- "unexpected error"
    rows$failure_detail <- conditionMessage(error)
    rows
  })
}
```

```{r additional-bootstrap, cache=FALSE}
additional_results <- if (.Platform$OS.type == "windows" ||
                          bootstrap_cores <= 1L) {
  lapply(seq_len(n_boot), safe_additional_replicate)
} else {
  parallel::mclapply(seq_len(n_boot), safe_additional_replicate,
                     mc.cores = bootstrap_cores, mc.preschedule = FALSE)
}
additional_bootstrap <- do.call(rbind, additional_results)
rownames(additional_bootstrap) <- NULL
additional_finite <- with(additional_bootstrap,
  is.finite(train_AUC) & is.finite(test_AUC) &
    is.finite(train_Brier) & is.finite(test_Brier))
stopifnot(
  nrow(additional_bootstrap) ==
    n_boot * length(dense_age_landmarks) * 2L,
  !anyDuplicated(additional_bootstrap[c("replicate", "model", "landmark")]),
  identical(additional_bootstrap$complete, additional_finite),
  all(nzchar(additional_bootstrap$failed_step[!additional_bootstrap$complete]))
)
```

```{r additional-bootstrap-diagnostics}
additional_failures <- additional_bootstrap[!additional_bootstrap$complete, ]
additional_failure_counts <- if (nrow(additional_failures)) {
  aggregate(replicate ~ model + landmark + failed_step + failure_detail,
            data = additional_failures, FUN = length)
} else data.frame()
if (nrow(additional_failure_counts)) {
  names(additional_failure_counts)[
    names(additional_failure_counts) == "replicate"
  ] <- "resamples"
}
rownames(additional_failure_counts) <- NULL
if (nrow(additional_failure_counts)) {
  knitr::kable(additional_failure_counts,
               caption = "Reasons for incomplete training/test score pairs")
}
first_age_bootstrap <- additional_bootstrap[
  additional_bootstrap$model == "Disjoint" &
    additional_bootstrap$landmark == first_dense_age, ]
stopifnot(nrow(first_age_bootstrap) == n_boot)
knitr::kable(data.frame(
  landmark = first_dense_age,
  zero_alias_repairs = sum(nzchar(first_age_bootstrap$zero_alias_terms)),
  complete_score_pairs = sum(first_age_bootstrap$complete),
  bootstrap_resamples = n_boot
), caption = "First-landmark separate-model prediction check")
```

## Performance across the three specifications

I applied the same training-minus-original correction to each model. The
counts show how many subject-resampling replicates yielded complete scores
at each landmark. A corrected score based on fewer than `r n_boot` replicates
is conditional on the finite fits; the failed separate fit remains visible.

```{r all-model-performance}
model_performance <- do.call(rbind, lapply(dense_age_landmarks, function(age) {
  counts <- dense_risk_sets[dense_risk_sets$landmark == age, ]
  stopifnot(nrow(counts) == 1L)
  do.call(rbind, lapply(c("Quadratic supermodel", "Linear supermodel", "Disjoint"),
    function(model) {
      if (model == "Quadratic supermodel") {
        apparent <- apparent_scores[apparent_scores$landmark == age, ]
        scored <- bootstrap_diagnostics[
          bootstrap_diagnostics$landmark == age &
            bootstrap_diagnostics$complete, ]
        apparent_error <- apparent$score_error
      } else {
        apparent <- additional_apparent[
          additional_apparent$model == model &
            additional_apparent$landmark == age, ]
        scored <- additional_bootstrap[
          additional_bootstrap$model == model &
            additional_bootstrap$landmark == age &
            additional_bootstrap$complete, ]
        apparent_error <- apparent$error
      }
      stopifnot(nrow(apparent) == 1L, nrow(scored) <= n_boot)
      data.frame(
        landmark = age, deaths = counts$deaths,
        transplants = counts$transplants, model = model,
        apparent_AUC = apparent$AUC,
        corrected_AUC = if (is.finite(apparent$AUC) && nrow(scored)) {
          apparent$AUC - mean(scored$train_AUC - scored$test_AUC)
        } else NA_real_,
        apparent_Brier = apparent$Brier,
        corrected_Brier = if (is.finite(apparent$Brier) && nrow(scored)) {
          apparent$Brier - mean(scored$train_Brier - scored$test_Brier)
        } else NA_real_,
        complete_resamples = nrow(scored),
        status = if (nzchar(apparent_error)) {
          paste("unavailable:", apparent_error)
        } else if (nrow(scored) < n_boot) {
          "conditional on complete resamples"
        } else "all resamples",
        row.names = NULL
      )
    }))
}))
stopifnot(
  nrow(model_performance) == length(dense_age_landmarks) * 3L,
  !anyDuplicated(model_performance[c("landmark", "model")]),
  isTRUE(all.equal(
    model_performance$corrected_AUC[
      model_performance$model == "Quadratic supermodel"],
    dense_performance$optimism_corrected_AUC, tolerance = 0
  )),
  isTRUE(all.equal(
    model_performance$corrected_Brier[
      model_performance$model == "Quadratic supermodel"],
    dense_performance$optimism_corrected_Brier, tolerance = 0
  ))
)
knitr::kable(model_performance, digits = 4,
             caption = "Death prediction at each attained-age landmark")
```

## Paired supermodel-minus-disjoint differences

I paired resamples by replicate and landmark, using only those with complete
training and original-cohort scores for both models. Positive AUC differences
and negative Brier differences favor the supermodel. Each percentile interval
spans the 2.5th and 97.5th percentiles of the replicate-specific
optimism-corrected difference; it is conditional on the paired finite fits.

```{r dense-paired-performance-differences}
paired_difference <- function(age, model) {
  super <- if (model == "Quadratic supermodel") {
    bootstrap_diagnostics[
      bootstrap_diagnostics$landmark == age &
        bootstrap_diagnostics$complete,
      c("replicate", "train_AUC", "test_AUC", "train_Brier", "test_Brier")
    ]
  } else {
    additional_bootstrap[
      additional_bootstrap$landmark == age &
        additional_bootstrap$model == model &
        additional_bootstrap$complete,
      c("replicate", "train_AUC", "test_AUC", "train_Brier", "test_Brier")
    ]
  }
  separate <- additional_bootstrap[
    additional_bootstrap$landmark == age &
      additional_bootstrap$model == "Disjoint" &
      additional_bootstrap$complete,
    c("replicate", "train_AUC", "test_AUC", "train_Brier", "test_Brier")
  ]
  paired <- merge(super, separate, by = "replicate",
                  suffixes = c("_super", "_disjoint"), sort = FALSE)
  stopifnot(!anyDuplicated(paired$replicate), nrow(paired) <= n_boot)
  apparent_super <- model_performance[
    model_performance$landmark == age & model_performance$model == model, ]
  apparent_separate <- model_performance[
    model_performance$landmark == age &
      model_performance$model == "Disjoint", ]
  stopifnot(nrow(apparent_super) == 1L, nrow(apparent_separate) == 1L)
  apparent_auc <- apparent_super$apparent_AUC - apparent_separate$apparent_AUC
  apparent_brier <- apparent_super$apparent_Brier -
    apparent_separate$apparent_Brier
  corrected_auc <- apparent_auc - with(paired,
    (train_AUC_super - test_AUC_super) -
      (train_AUC_disjoint - test_AUC_disjoint))
  corrected_brier <- apparent_brier - with(paired,
    (train_Brier_super - test_Brier_super) -
      (train_Brier_disjoint - test_Brier_disjoint))
  summarize_difference <- function(values) {
    if (length(values) < 2L || any(!is.finite(values))) {
      return(c(mean = NA_real_, lower = NA_real_, upper = NA_real_))
    }
    c(mean = mean(values),
      lower = unname(quantile(values, 0.025, names = FALSE)),
      upper = unname(quantile(values, 0.975, names = FALSE)))
  }
  auc <- summarize_difference(corrected_auc)
  brier <- summarize_difference(corrected_brier)
  data.frame(
    landmark = age, supermodel = model,
    apparent_AUC_difference = apparent_auc,
    corrected_AUC_difference = auc["mean"],
    AUC_lower = auc["lower"], AUC_upper = auc["upper"],
    apparent_Brier_difference = apparent_brier,
    corrected_Brier_difference = brier["mean"],
    Brier_lower = brier["lower"], Brier_upper = brier["upper"],
    paired_resamples = nrow(paired),
    status = if (!is.finite(apparent_auc) || !is.finite(apparent_brier)) {
      "unavailable original separate fit"
    } else if (nrow(paired) < 2L) {
      "insufficient paired resamples"
    } else if (nrow(paired) < n_boot) {
      "conditional on paired finite fits"
    } else "all resamples",
    row.names = NULL
  )
}
paired_differences <- do.call(rbind, lapply(dense_age_landmarks, function(age) {
  do.call(rbind, lapply(c("Quadratic supermodel", "Linear supermodel"),
                         function(model) paired_difference(age, model)))
}))
stopifnot(nrow(paired_differences) == length(dense_age_landmarks) * 2L)
knitr::kable(paired_differences, digits = 4,
             caption = "Supermodel minus separate-model death prediction")
```

These comparisons use only resamples for which the separate model could be
scored on both the resampled and original risk sets.

# Session

```{r session-info}
sessionInfo()
```
