---
title: "PBC attained-age landmark sensitivity analysis"
output:
  html_document:
    self_contained: true
    toc: true
    code_folding: show
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center")
library(here)
devtools::load_all(here::here(), quiet = TRUE)

prediction_window <- 5
dense_age_landmarks <- seq(42, 56, by = 2)
cause_labels <- c("Transplant", "Death")
n_boot <- getOption("dense_age.n_boot", 200L)
bootstrap_seed <- 20260922L
stopifnot(length(n_boot) == 1L, is.finite(n_boot),
          n_boot >= 2L, n_boot == as.integer(n_boot))
n_boot <- as.integer(n_boot)
```

# Purpose

I kept the original time-scale comparison unchanged. This separate analysis
tests whether a more closely spaced attained-age grid identifies landmark main
effects, how its hazard ratios compare with separate landmark fits, and how
often its predictions survive subject-level resampling.
The candidate grid is a sensitivity analysis, not a replacement for the
entry-age-quartile analysis.

```{r analysis-data}
pbc_long <- get_pbc_long()
entry_ages <- unique(pbc_long[c("id", "age")])$age
current_age_landmarks <- round(
  unname(quantile(entry_ages, c(0.25, 0.50, 0.75), type = 8)), 1
)

pbc_age <- transform(
  pbc_long,
  log_bili_z = (log(bili) - mean(log(pbc_long$bili))) /
    sd(log(pbc_long$bili)),
  albumin_z = (albumin - mean(pbc_long$albumin)) / sd(pbc_long$albumin),
  entry_age = age,
  exit_age = age + time,
  visit_age = age + tstart
)

stopifnot(
  !anyDuplicated(unique(pbc_age[c("id", "entry_age")])$id),
  all(is.finite(pbc_age$log_bili_z)),
  all(is.finite(pbc_age$albumin_z)),
  all(diff(current_age_landmarks) > prediction_window),
  all(diff(dense_age_landmarks) < prediction_window)
)
```

# Why the current age grid cannot estimate landmark main effects

I plotted the original landmark windows to make the lack of overlap visible.
Each event risk set lies within one window. An LM main effect is therefore
constant within that risk set and cancels from the Cox partial likelihood.
The model still contains bilirubin-by-LM and albumin-by-LM interactions; it
does not contain an estimable LM main effect on this grid.

```{r current-window-plot, fig.width=7, fig.height=3.3}
draw_current_windows <- function() {
  window_ends <- current_age_landmarks + prediction_window
  plot(NA, xlim = range(c(current_age_landmarks, window_ends)) + c(-0.5, 0.5),
       ylim = c(0.5, length(current_age_landmarks) + 0.5),
       yaxt = "n", xlab = "Attained age (years)", ylab = "",
       main = "Current attained-age prediction windows", bty = "n")
  axis(2, at = seq_along(current_age_landmarks),
       labels = paste0("LM ", format(current_age_landmarks, nsmall = 1)),
       las = 1, tick = FALSE)
  abline(v = c(current_age_landmarks, window_ends), col = "grey90", lty = 3)
  segments(current_age_landmarks, seq_along(current_age_landmarks),
           window_ends, seq_along(current_age_landmarks),
           lwd = 7, col = "#32658A", lend = "butt")
  points(current_age_landmarks, seq_along(current_age_landmarks), pch = 16)
  points(window_ends, seq_along(current_age_landmarks), pch = 1)
  text((current_age_landmarks + window_ends) / 2,
       seq_along(current_age_landmarks) + 0.16,
       labels = sprintf("%.1f–%.1f", current_age_landmarks, window_ends),
       cex = 0.85)
}

draw_current_windows()
pdf(here::here("analysis", "current_age_windows.pdf"),
    width = 7, height = 3.3)
draw_current_windows()
dev.off()
```

```{r current-window-check}
current_stack <- stack_data(
  pbc_age, list(time = "exit_age", status = "status"),
  lms = current_age_landmarks, w = prediction_window,
  covs = list(fixed = c("male", "stage", "trt"),
              varying = c("log_bili_z", "albumin_z")),
  format = "long", id = "id", rtime = "visit_age",
  entry = "entry_age", entry_mode = "strict"
)

risk_set_lm_counts <- function(stack) {
  event_times <- sort(unique(stack$data$exit_age[stack$data$status > 0]))
  vapply(event_times, function(event_age) {
    at_risk <- stack$data$LM < event_age & stack$data$exit_age >= event_age
    length(unique(stack$data$LM[at_risk]))
  }, integer(1))
}

current_lm_counts <- risk_set_lm_counts(current_stack)
stopifnot(length(current_lm_counts) > 0L,
          all(current_lm_counts == 1L))
knitr::kable(data.frame(
  grid = "Current entry-age quartiles",
  event_times = length(current_lm_counts),
  event_times_with_multiple_LMs = sum(current_lm_counts > 1L)
))
```

# Dense-grid risk sets

I used the same strict-entry rule, event coding, prediction window, and
laboratory transformations as the primary attained-age analysis. I checked
the number of people and both event types in every window before fitting or
resampling. These counts are stacked-window counts: an event can contribute
to more than one overlapping window.

```{r dense-risk-sets}
dense_stack_raw <- stack_data(
  pbc_age, list(time = "exit_age", status = "status"),
  lms = dense_age_landmarks, w = prediction_window,
  covs = list(fixed = c("male", "stage", "trt"),
              varying = c("log_bili_z", "albumin_z")),
  format = "long", id = "id", rtime = "visit_age",
  entry = "entry_age", entry_mode = "strict"
)

dense_risk_sets <- do.call(rbind, lapply(dense_age_landmarks, function(age) {
  rows <- dense_stack_raw$data[dense_stack_raw$data$LM == age, , drop = FALSE]
  data.frame(
    landmark = age,
    window_end = age + prediction_window,
    subjects = length(unique(rows$id)),
    deaths = sum(rows$status == 2L),
    transplants = sum(rows$status == 1L),
    censored = sum(rows$status == 0L)
  )
}))

dense_lm_counts <- risk_set_lm_counts(dense_stack_raw)
stopifnot(
  identical(dense_stack_raw$entry_mode, "strict"),
  all(dense_stack_raw$data$entry_age <= dense_stack_raw$data$LM),
  all(dense_stack_raw$data$LM < dense_stack_raw$data$exit_age),
  !anyDuplicated(dense_stack_raw$data[c("id", "LM")]),
  all(dense_risk_sets$subjects > 0L),
  all(dense_risk_sets$subjects ==
        dense_risk_sets$deaths + dense_risk_sets$transplants +
        dense_risk_sets$censored),
  any(dense_lm_counts > 1L)
)

knitr::kable(dense_risk_sets,
             caption = "Strict-entry risk sets on the candidate age grid")
knitr::kable(data.frame(
  grid = "Candidate two-year grid",
  event_times = length(dense_lm_counts),
  event_times_with_multiple_LMs = sum(dense_lm_counts > 1L),
  maximum_LMs_in_one_risk_set = max(dense_lm_counts)
))
```

The dense windows overlap, and the observed event risk sets contain multiple
landmark values. This removes the exact non-identifiability of LM main effects
seen on the original grid. It does not by itself establish that the
competing-risk fit or its bootstrap is stable: the event-count table includes
windows with very few transplants.

# Candidate supermodel fit

I used a centered quadratic landmark basis for both main effects and the two
laboratory interactions. The other predictors and the cause-specific CSC fit
match the primary attained-age specification.

```{r dense-supermodel}
dense_center <- mean(dense_age_landmarks)
dense_half_range <- diff(range(dense_age_landmarks)) / 2
dense_basis <- list(
  function(age) (age - dense_center) / dense_half_range,
  function(age) ((age - dense_center) / dense_half_range)^2
)
dense_stack <- add_interactions(
  dense_stack_raw,
  lm_covs = c("log_bili_z", "albumin_z"),
  func_covars = dense_basis,
  func_lms = dense_basis
)
stopifnot(all(c("LM1", "LM2", "log_bili_z_LM1", "log_bili_z_LM2",
                "albumin_z_LM1", "albumin_z_LM2") %in%
              names(dense_stack$data)))

dense_formula <- Hist(exit_age, status, LM) ~
  male + stage + trt +
  log_bili_z + log_bili_z_LM1 + log_bili_z_LM2 +
  albumin_z + albumin_z_LM1 + albumin_z_LM2 +
  LM1 + LM2 + cluster(id)

capture_conditions <- function(expression) {
  warnings <- character()
  value <- withCallingHandlers(
    tryCatch(eval.parent(substitute(expression)), error = function(error) error),
    warning = function(warning) {
      warnings <<- c(warnings, conditionMessage(warning))
      invokeRestart("muffleWarning")
    }
  )
  list(value = value, warnings = unique(warnings))
}

dense_fit_record <- capture_conditions(dynamic_lm(
  dense_stack, dense_formula, type = "CSC", method = "breslow", x = TRUE
))

fit_status <- data.frame(
  error = if (inherits(dense_fit_record$value, "error")) {
    conditionMessage(dense_fit_record$value)
  } else "",
  warnings = paste(dense_fit_record$warnings, collapse = " | ")
)
knitr::kable(fit_status)

if (!inherits(dense_fit_record$value, "error")) {
  dense_fit <- dense_fit_record$value
  dense_lm_inference <- do.call(rbind, lapply(seq_along(
    dense_fit$model$causes
  ), function(index) {
    fit <- dense_fit$model$models[[index]]
    terms <- c("LM1", "LM2")
    data.frame(
      cause = cause_labels[as.integer(dense_fit$model$causes[index])],
      term = terms,
      coefficient = unname(coef(fit)[terms]),
      robust_se = unname(sqrt(diag(vcov(fit)))[terms])
    )
  }))
  stopifnot(all(is.finite(dense_lm_inference$coefficient)),
            all(is.finite(dense_lm_inference$robust_se)))
  knitr::kable(dense_lm_inference, digits = 3)
}
```

# Separate landmark fits as a sparsity check

I also tried the unchanged, one-model-per-landmark specification. I recorded
all fit warnings and errors rather than modifying a sparse fit to make it
converge. The pooled supermodel can be estimable even when a separate
competing-risk fit at one age is not.

```{r dense-disjoint-fit-check}
dense_disjoint_formula <- Hist(exit_age, status, LM) ~
  male + stage + trt + log_bili_z + albumin_z + cluster(id)

dense_disjoint_records <- setNames(lapply(
  dense_age_landmarks, function(age) {
    one_landmark <- dense_stack_raw
    one_landmark$data <- dense_stack_raw$data[
      dense_stack_raw$data$LM == age, , drop = FALSE
    ]
    one_landmark$end_time <- age
    capture_conditions(dynamic_lm(
      one_landmark, dense_disjoint_formula,
      type = "CSC", method = "breslow", x = TRUE
    ))
  }), as.character(dense_age_landmarks)
)
dense_disjoint_conditions <- do.call(rbind, lapply(
  dense_age_landmarks, function(age) {
    fit_record <- dense_disjoint_records[[as.character(age)]]
    data.frame(landmark = age,
               fit_error = if (inherits(fit_record$value, "error")) {
                 conditionMessage(fit_record$value)
               } else "",
               warnings = paste(fit_record$warnings, collapse = " | "))
  }))

stopifnot(nrow(dense_disjoint_conditions) == length(dense_age_landmarks))
knitr::kable(dense_disjoint_conditions)
```

# Death-model laboratory hazard ratios

I combined each laboratory main effect with its landmark interactions and
used the fitted robust covariance for the supermodel interval. For each
separate landmark fit, I used the death-model coefficient and robust standard
error. CI-width ratios are on the log-HR scale, as in the primary comparison.
An unavailable separate fit remains unavailable in the table and figure.

```{r dense-hazard-ratio-table}
death_index <- match(2L, dense_fit$model$causes)
stopifnot(length(death_index) == 1L, !is.na(death_index))
dense_death_fit <- dense_fit$model$models[[death_index]]
robust_se <- sqrt(diag(vcov(dense_death_fit)))
summary_se <- summary(dense_death_fit)$coefficients[, "robust se"]
stopifnot(identical(names(robust_se), names(coef(dense_death_fit))),
          max(abs(robust_se - summary_se), na.rm = TRUE) < 1e-12)

critical_value <- qnorm(0.975)
laboratory_terms <- c(log_bili_z = "Log bilirubin (per SD)",
                      albumin_z = "Albumin (per SD)")

supermodel_log_hr <- function(marker, ages) {
  terms <- c(marker, paste0(marker, c("_LM1", "_LM2")))
  design <- cbind(1, dense_basis[[1]](ages), dense_basis[[2]](ages))
  beta <- coef(dense_death_fit)[terms]
  covariance <- vcov(dense_death_fit)[terms, terms, drop = FALSE]
  estimate <- drop(design %*% beta)
  se <- sqrt(rowSums((design %*% covariance) * design))
  stopifnot(all(is.finite(estimate)), all(is.finite(se)))
  data.frame(landmark = ages, estimate = estimate,
             lower = estimate - critical_value * se,
             upper = estimate + critical_value * se)
}

disjoint_log_hr <- function(age, marker) {
  record <- dense_disjoint_records[[as.character(age)]]
  if (inherits(record$value, "error")) {
    return(c(estimate = NA_real_, lower = NA_real_, upper = NA_real_))
  }
  cause_index <- match(2L, record$value$model$causes)
  stopifnot(!is.na(cause_index))
  fit <- record$value$model$models[[cause_index]]
  estimate <- coef(fit)[marker]
  se <- sqrt(diag(vcov(fit)))[marker]
  if (!is.finite(estimate) || !is.finite(se)) {
    return(c(estimate = NA_real_, lower = NA_real_, upper = NA_real_))
  }
  c(estimate = estimate,
    lower = estimate - critical_value * se,
    upper = estimate + critical_value * se)
}

laboratory_comparison <- do.call(rbind, lapply(names(laboratory_terms),
  function(marker) {
    super <- supermodel_log_hr(marker, dense_age_landmarks)
    separate <- t(vapply(dense_age_landmarks,
      function(age) disjoint_log_hr(age, marker),
      c(estimate = 0, lower = 0, upper = 0)))
    conditions <- dense_disjoint_conditions[
      match(dense_age_landmarks, dense_disjoint_conditions$landmark), ]
    data.frame(
      covariate = unname(laboratory_terms[marker]),
      landmark = dense_age_landmarks,
      super_estimate = super$estimate,
      super_lower = super$lower,
      super_upper = super$upper,
      disjoint_estimate = separate[, "estimate"],
      disjoint_lower = separate[, "lower"],
      disjoint_upper = separate[, "upper"],
      CI_width_ratio_log_scale =
        (separate[, "upper"] - separate[, "lower"]) /
        (super$upper - super$lower),
      fit_flag = ifelse(nzchar(conditions$fit_error), "failed",
                        ifelse(nzchar(conditions$warnings), "warning", "ok")),
      row.names = NULL
    )
  }
))

format_hr <- function(estimate, lower, upper) {
  ifelse(is.finite(estimate) & is.finite(lower) & is.finite(upper),
         sprintf("%.3f (%.3f, %.3f)",
                 exp(estimate), exp(lower), exp(upper)), "—")
}
laboratory_display <- data.frame(
  covariate = laboratory_comparison$covariate,
  landmark = laboratory_comparison$landmark,
  supermodel_HR_95_CI = with(laboratory_comparison,
    format_hr(super_estimate, super_lower, super_upper)),
  disjoint_HR_95_CI = with(laboratory_comparison,
    format_hr(disjoint_estimate, disjoint_lower, disjoint_upper)),
  disjoint_to_supermodel_CI_width =
    laboratory_comparison$CI_width_ratio_log_scale,
  fit_flag = laboratory_comparison$fit_flag
)
stopifnot(nrow(laboratory_display) ==
            length(dense_age_landmarks) * length(laboratory_terms))
knitr::kable(laboratory_display, digits = 2)

finite_comparisons <- laboratory_comparison[
  is.finite(laboratory_comparison$disjoint_estimate), ]
different_at_each_landmark <- all(abs(
  finite_comparisons$super_estimate - finite_comparisons$disjoint_estimate
) > 1e-8)
cat(if (different_at_each_landmark) {
  "The quadratic supermodel now smooths: its laboratory HRs do not reproduce the separate-model HRs at any estimable landmark."
} else {
  "The quadratic supermodel differs from the separate models at some, but not all, estimable landmarks."
})

condition_rows <- dense_disjoint_conditions[
  nzchar(dense_disjoint_conditions$fit_error) |
    nzchar(dense_disjoint_conditions$warnings), ]
condition_text <- vapply(seq_len(nrow(condition_rows)), function(index) {
  row <- condition_rows[index, ]
  paste0("age ", row$landmark, ": ",
         paste(c(row$fit_error, row$warnings)[
           nzchar(c(row$fit_error, row$warnings))], collapse = " | "))
}, character(1))
hr_figure_caption <- paste(
  "Death-model HR per SD of the laboratory measure; shaded band and bars are robust 95% CIs.",
  "Filled points: unflagged separate fits; open points: warned fits; crosses at the lower axis: failed fits.",
  "Recorded conditions:", paste(condition_text, collapse = "; ")
)
```

```{r dense-hazard-ratio-plot, fig.width=10, fig.height=5, fig.cap=hr_figure_caption}
plot_laboratory_hr <- function(marker, label) {
  ages <- seq(min(dense_age_landmarks), max(dense_age_landmarks),
              length.out = 200L)
  curve <- supermodel_log_hr(marker, ages)
  rows <- laboratory_comparison[
    laboratory_comparison$covariate == label, ]
  y_values <- exp(c(curve$lower, curve$upper,
                    rows$disjoint_lower, rows$disjoint_upper, 0))
  y_limits <- range(y_values[is.finite(y_values) & y_values > 0])
  plot(ages, exp(curve$estimate), type = "n", log = "y",
       ylim = y_limits, xlab = "Attained age (years)",
       ylab = "Hazard ratio (log scale)", main = label, bty = "l")
  polygon(c(ages, rev(ages)),
          c(exp(curve$lower), rev(exp(curve$upper))),
          col = "grey85", border = NA)
  abline(h = 1, lty = 2, col = "grey45")
  lines(ages, exp(curve$estimate), lwd = 2)
  finite <- is.finite(rows$disjoint_estimate) &
    is.finite(rows$disjoint_lower) & is.finite(rows$disjoint_upper)
  arrows(rows$landmark[finite], exp(rows$disjoint_lower[finite]),
         rows$landmark[finite], exp(rows$disjoint_upper[finite]),
         angle = 90, code = 3, length = 0.04)
  points(rows$landmark[finite], exp(rows$disjoint_estimate[finite]),
         pch = ifelse(rows$fit_flag[finite] == "warning", 21, 19),
         bg = "white")
  failed <- rows$fit_flag == "failed" | !finite
  if (any(failed)) {
    points(rows$landmark[failed], y_limits[1] * 1.05, pch = 4, cex = 1.2)
  }
}

old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
for (marker in names(laboratory_terms)) {
  plot_laboratory_hr(marker, unname(laboratory_terms[marker]))
}
par(old_par)
```

# Supermodel prediction and bootstrap

I kept the disjoint fit failures above as findings. For prediction I evaluated
the pooled supermodel alone. The bootstrap resampled people rather than
stacked rows, assigned a new ID to each resampled copy, rebuilt the strict
age stack, refitted both cause-specific hazards, and scored the training and
original cohorts at each landmark. The grid, window, transformations, and
seed were fixed before resampling.

```{r dense-bootstrap-functions}
subset_prediction <- function(prediction, age) {
  rows <- prediction$data$LM == age
  if (!any(rows)) stop("No predicted rows at age ", age)
  out <- prediction
  out$data <- prediction$data[rows, , drop = FALSE]
  out$preds <- prediction$preds[rows, , drop = FALSE]
  stopifnot(nrow(out$data) == nrow(out$preds))
  out
}

score_prediction <- function(prediction, age) {
  if (anyNA(prediction$preds$risk)) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = "NA predicted risk"))
  }
  record <- capture_conditions(score(
    list(Model = prediction), times = age, metrics = c("auc", "brier"),
    id_col = "id", cause = 2L, summary = FALSE, contrasts = FALSE
  ))
  if (inherits(record$value, "error")) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = conditionMessage(record$value)))
  }
  auc <- as.data.frame(record$value$AUC$score)
  brier <- as.data.frame(record$value$Brier$score)
  auc <- auc[as.character(auc$model) == "Model", , drop = FALSE]
  brier <- brier[as.character(brier$model) == "Model", , drop = FALSE]
  if (nrow(auc) != 1L || nrow(brier) != 1L ||
      !is.finite(auc$AUC) || !is.finite(brier$Brier)) {
    return(list(AUC = NA_real_, Brier = NA_real_,
                error = "non-finite or missing score"))
  }
  list(AUC = auc$AUC, Brier = brier$Brier, error = "")
}

resample_people <- function(data, sampled_ids) {
  copies <- lapply(seq_along(sampled_ids), function(index) {
    rows <- data[data$id == sampled_ids[index], , drop = FALSE]
    stopifnot(nrow(rows) > 0L)
    rows$id <- index
    rows
  })
  sample_data <- do.call(rbind, copies)
  rownames(sample_data) <- NULL
  stopifnot(length(unique(sample_data$id)) == length(sampled_ids))
  sample_data
}

build_dense_stack <- function(data) {
  raw <- stack_data(
    data, list(time = "exit_age", status = "status"),
    lms = dense_age_landmarks, w = prediction_window,
    covs = list(fixed = c("male", "stage", "trt"),
                varying = c("log_bili_z", "albumin_z")),
    format = "long", id = "id", rtime = "visit_age",
    entry = "entry_age", entry_mode = "strict"
  )
  stopifnot(!anyDuplicated(raw$data[c("id", "LM")]),
            all(raw$data$entry_age <= raw$data$LM))
  model <- add_interactions(
    raw, lm_covs = c("log_bili_z", "albumin_z"),
    func_covars = dense_basis, func_lms = dense_basis
  )
  stopifnot(!anyDuplicated(model$data[c("id", "LM")]),
            identical(raw$data[c("id", "LM")], model$data[c("id", "LM")]))
  model
}

empty_bootstrap_rows <- function(replicate) {
  data.frame(
    replicate = replicate, landmark = dense_age_landmarks,
    zero_transplants = NA, zero_deaths = NA,
    transplant_fit_error = "", transplant_fit_warning = "",
    death_fit_error = "", death_fit_warning = "",
    nonfinite_cause = "", pooled_fit_warning = "",
    train_AUC = NA_real_, test_AUC = NA_real_,
    train_Brier = NA_real_, test_Brier = NA_real_,
    failed_step = "", failure_detail = "", complete = FALSE
  )
}

diagnose_cause <- function(stack, cause) {
  formula <- as.formula(paste0(
    "survival::Surv(LM, exit_age, status == ", cause, ") ~ ",
    "male + stage + trt + log_bili_z + log_bili_z_LM1 + ",
    "log_bili_z_LM2 + albumin_z + albumin_z_LM1 + ",
    "albumin_z_LM2 + LM1 + LM2 + cluster(id)"
  ))
  record <- capture_conditions(survival::coxph(
    formula, data = stack$data, method = "breslow", x = TRUE
  ))
  list(
    error = if (inherits(record$value, "error")) {
      conditionMessage(record$value)
    } else "",
    warning = paste(record$warnings, collapse = " | ")
  )
}

original_prediction_record <- capture_conditions(predict(
  dense_fit, newdata = dense_stack, cause = 2L
))
stopifnot(!inherits(original_prediction_record$value, "error"))
original_prediction <- original_prediction_record$value
apparent_scores <- do.call(rbind, lapply(dense_age_landmarks, function(age) {
  metric <- score_prediction(subset_prediction(original_prediction, age), age)
  data.frame(landmark = age, AUC = metric$AUC, Brier = metric$Brier,
             score_error = metric$error)
}))
stopifnot(all(is.finite(apparent_scores$AUC)),
          all(is.finite(apparent_scores$Brier)),
          !any(nzchar(apparent_scores$score_error)))
knitr::kable(apparent_scores, digits = 4)
```

```{r dense-bootstrap, cache=FALSE}
set.seed(bootstrap_seed)
original_ids <- unique(pbc_age$id)
bootstrap_samples <- lapply(seq_len(n_boot), function(index) {
  sample(original_ids, length(original_ids), replace = TRUE)
})

run_dense_replicate <- function(index) {
  row <- empty_bootstrap_rows(index)
  sampled <- capture_conditions(resample_people(
    pbc_age, bootstrap_samples[[index]]
  ))
  if (inherits(sampled$value, "error")) {
    row$failed_step <- "resample"
    row$failure_detail <- conditionMessage(sampled$value)
    return(row)
  }
  stack_record <- capture_conditions(build_dense_stack(sampled$value))
  if (inherits(stack_record$value, "error")) {
    row$failed_step <- "stack"
    row$failure_detail <- conditionMessage(stack_record$value)
    return(row)
  }
  training_stack <- stack_record$value
  for (age in dense_age_landmarks) {
    at_age <- training_stack$data$LM == age
    row$zero_transplants[row$landmark == age] <-
      sum(training_stack$data$status[at_age] == 1L) == 0L
    row$zero_deaths[row$landmark == age] <-
      sum(training_stack$data$status[at_age] == 2L) == 0L
  }

  fit_record <- capture_conditions(dynamic_lm(
    training_stack, dense_formula, type = "CSC",
    method = "breslow", x = TRUE
  ))
  row$pooled_fit_warning <- paste(fit_record$warnings, collapse = " | ")
  if (inherits(fit_record$value, "error") || length(fit_record$warnings)) {
    transplant <- diagnose_cause(training_stack, 1L)
    death <- diagnose_cause(training_stack, 2L)
    row$transplant_fit_error <- transplant$error
    row$transplant_fit_warning <- transplant$warning
    row$death_fit_error <- death$error
    row$death_fit_warning <- death$warning
  }
  if (inherits(fit_record$value, "error")) {
    row$failed_step <- "pooled fit"
    row$failure_detail <- conditionMessage(fit_record$value)
    return(row)
  }

  nonfinite <- vapply(seq_along(fit_record$value$model$causes),
    function(cause_index) {
      fit <- fit_record$value$model$models[[cause_index]]
      any(!is.finite(coef(fit)))
    }, logical(1))
  row$nonfinite_cause <- paste(
    fit_record$value$model$causes[nonfinite], collapse = ", "
  )

  train_prediction <- capture_conditions(predict(
    fit_record$value, newdata = training_stack, cause = 2L
  ))
  test_prediction <- capture_conditions(predict(
    fit_record$value, newdata = dense_stack, cause = 2L
  ))
  for (age in dense_age_landmarks) {
    at_age <- which(row$landmark == age)
    if (inherits(train_prediction$value, "error")) {
      row$failed_step[at_age] <- "train prediction"
      row$failure_detail[at_age] <- conditionMessage(train_prediction$value)
      next
    }
    if (inherits(test_prediction$value, "error")) {
      row$failed_step[at_age] <- "test prediction"
      row$failure_detail[at_age] <- conditionMessage(test_prediction$value)
      next
    }
    train_subset <- capture_conditions(subset_prediction(
      train_prediction$value, age
    ))
    test_subset <- capture_conditions(subset_prediction(
      test_prediction$value, age
    ))
    if (inherits(train_subset$value, "error")) {
      row$failed_step[at_age] <- "train prediction subset"
      row$failure_detail[at_age] <- conditionMessage(train_subset$value)
      next
    }
    if (inherits(test_subset$value, "error")) {
      row$failed_step[at_age] <- "test prediction subset"
      row$failure_detail[at_age] <- conditionMessage(test_subset$value)
      next
    }
    train <- score_prediction(train_subset$value, age)
    test <- score_prediction(test_subset$value, age)
    row$train_AUC[at_age] <- train$AUC
    row$test_AUC[at_age] <- test$AUC
    row$train_Brier[at_age] <- train$Brier
    row$test_Brier[at_age] <- test$Brier
    if (nzchar(train$error)) {
      row$failed_step[at_age] <- if (train$error == "NA predicted risk") {
        "train prediction"
      } else "train score"
      row$failure_detail[at_age] <- train$error
    } else if (nzchar(test$error)) {
      row$failed_step[at_age] <- if (test$error == "NA predicted risk") {
        "test prediction"
      } else "test score"
      row$failure_detail[at_age] <- test$error
    } else {
      row$complete[at_age] <- TRUE
    }
  }
  row
}

safe_dense_replicate <- function(index) {
  tryCatch(run_dense_replicate(index), error = function(error) {
    row <- empty_bootstrap_rows(index)
    row$failed_step <- "unexpected error"
    row$failure_detail <- conditionMessage(error)
    row
  })
}

bootstrap_cores <- getOption(
  "timescale.bootstrap.cores",
  if (.Platform$OS.type == "windows") 1L else
    min(4L, parallel::detectCores(logical = FALSE))
)
bootstrap_rows <- if (.Platform$OS.type == "windows" ||
                      bootstrap_cores <= 1L) {
  lapply(seq_len(n_boot), safe_dense_replicate)
} else {
  parallel::mclapply(seq_len(n_boot), safe_dense_replicate,
                     mc.cores = bootstrap_cores, mc.preschedule = FALSE)
}
bootstrap_diagnostics <- do.call(rbind, bootstrap_rows)
rownames(bootstrap_diagnostics) <- NULL
finite_scores <- with(bootstrap_diagnostics,
  is.finite(train_AUC) & is.finite(test_AUC) &
    is.finite(train_Brier) & is.finite(test_Brier))
if (!identical(bootstrap_diagnostics$complete, finite_scores)) {
  print(bootstrap_diagnostics[
    bootstrap_diagnostics$complete != finite_scores,
    c("replicate", "landmark", "complete", "failed_step",
      "train_AUC", "test_AUC", "train_Brier", "test_Brier")
  ])
}
stopifnot(
  nrow(bootstrap_diagnostics) == n_boot * length(dense_age_landmarks),
  !anyDuplicated(bootstrap_diagnostics[c("replicate", "landmark")]),
  identical(bootstrap_diagnostics$complete, finite_scores),
  all(nzchar(bootstrap_diagnostics$failed_step[!bootstrap_diagnostics$complete]))
)
```

I separated zero-transplant windows from actual failures: a zero count in one
window does not necessarily invalidate a pooled fit. The reason table gives
the first step that prevented a complete train/test score, retaining the
error text. Cause-specific fit conditions are shown separately.

```{r dense-bootstrap-diagnostics}
bootstrap_completion <- do.call(rbind, lapply(dense_age_landmarks,
  function(age) {
    rows <- bootstrap_diagnostics[
      bootstrap_diagnostics$landmark == age, ]
    data.frame(
      landmark = age,
      complete_score_pairs = sum(rows$complete),
      incomplete_score_pairs = sum(!rows$complete),
      complete_with_fit_warning = sum(
        rows$complete & nzchar(rows$pooled_fit_warning)),
      zero_transplant_resamples = sum(rows$zero_transplants, na.rm = TRUE),
      zero_death_resamples = sum(rows$zero_deaths, na.rm = TRUE)
    )
  }
))
complete_everywhere <- tapply(bootstrap_diagnostics$complete,
                              bootstrap_diagnostics$replicate, all)
stopifnot(length(complete_everywhere) == n_boot,
          all(bootstrap_completion$complete_score_pairs +
                bootstrap_completion$incomplete_score_pairs == n_boot))

failure_rows <- bootstrap_diagnostics[!bootstrap_diagnostics$complete, ]
failure_reasons <- if (nrow(failure_rows)) {
  aggregate(replicate ~ landmark + failed_step + failure_detail,
            data = failure_rows, FUN = length)
} else data.frame()
if (nrow(failure_reasons)) {
  names(failure_reasons)[names(failure_reasons) == "replicate"] <- "resamples"
  failure_reasons <- failure_reasons[order(
    failure_reasons$landmark, failure_reasons$failed_step,
    failure_reasons$failure_detail), ]
  rownames(failure_reasons) <- NULL
}

cause_conditions <- unique(bootstrap_diagnostics[c(
  "replicate", "transplant_fit_error", "transplant_fit_warning",
  "death_fit_error", "death_fit_warning", "nonfinite_cause",
  "pooled_fit_warning"
)])
cause_condition_counts <- data.frame(
  condition = c("Transplant fit error", "Transplant fit warning",
                "Death fit error", "Death fit warning",
                "Non-finite cause coefficient", "Pooled fit warning"),
  resamples = c(
    sum(nzchar(cause_conditions$transplant_fit_error)),
    sum(nzchar(cause_conditions$transplant_fit_warning)),
    sum(nzchar(cause_conditions$death_fit_error)),
    sum(nzchar(cause_conditions$death_fit_warning)),
    sum(nzchar(cause_conditions$nonfinite_cause)),
    sum(nzchar(cause_conditions$pooled_fit_warning))
  )
)

knitr::kable(bootstrap_completion)
knitr::kable(data.frame(
  bootstrap_resamples = n_boot,
  complete_at_every_landmark = sum(complete_everywhere)
))
if (nrow(failure_reasons)) knitr::kable(failure_reasons)
knitr::kable(cause_condition_counts)
```

At age `r dense_age_landmarks[which.max(bootstrap_completion$zero_transplant_resamples)]`,
`r max(bootstrap_completion$zero_transplant_resamples)` resamples had no
transplants in that window, yet `r bootstrap_completion$complete_score_pairs[which.max(bootstrap_completion$zero_transplant_resamples)]`
still yielded complete pooled-model scores. Across all landmarks,
`r sum(complete_everywhere)` resamples were complete throughout. Finite scores
from a warned cause-specific fit remain flagged; I did not count a warning
alone as a scoring failure.

## Apparent and optimism-corrected performance

I subtracted the mean training-minus-original optimism from each apparent
landmark score. When not every resample had finite training and test scores,
the corrected value is conditional on the complete resamples at that
landmark; I do not treat it as an unconditional estimate.

```{r dense-performance-table}
dense_performance <- do.call(rbind, lapply(dense_age_landmarks,
  function(age) {
    apparent <- apparent_scores[apparent_scores$landmark == age, ]
    paired <- bootstrap_diagnostics[
      bootstrap_diagnostics$landmark == age &
        bootstrap_diagnostics$complete, ]
    stopifnot(nrow(apparent) == 1L)
    data.frame(
      landmark = age,
      apparent_AUC = apparent$AUC,
      optimism_corrected_AUC = if (nrow(paired)) {
        apparent$AUC - mean(paired$train_AUC - paired$test_AUC)
      } else NA_real_,
      apparent_Brier = apparent$Brier,
      optimism_corrected_Brier = if (nrow(paired)) {
        apparent$Brier - mean(paired$train_Brier - paired$test_Brier)
      } else NA_real_,
      complete_score_pairs = nrow(paired),
      interpretation = if (nrow(paired) == n_boot) {
        "all resamples"
      } else "conditional on complete resamples"
    )
  }
))
stopifnot(identical(dense_performance$complete_score_pairs,
                    bootstrap_completion$complete_score_pairs))
knitr::kable(dense_performance, digits = 4)
```

# Session

```{r session-info}
sessionInfo()
```
