---
title: "PBC landmark supermodels on enrollment and attained-age time scales"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  out.width = "95%"
)
options(width = 110)

library(here)
devtools::load_all(here::here(), quiet = TRUE)
```

# Purpose

I compared the PBC landmark supermodel on two clocks. The enrollment-time
analysis asked about risk a given number of years after enrollment. The
attained-age analysis asked about risk at a given age. I used the same PBC
records, prediction horizon, outcome, laboratory transformations, and non-age
predictors on both sides.

These are not competing estimates of the same quantity. An enrollment landmark
conditions on survival for a specified time after enrollment; an age landmark
conditions on being observed and event-free at a specified age. The risk sets
and scientific questions therefore differ. A difference between the fitted
models is not bias in either direction.

The age scale is useful when a common time zero has no defensible clinical
meaning, as in a cohort assembled from electronic health records. It also
requires an additional eligibility rule: a person must have entered observation
by the landmark age. I used strict eligibility, `entry <= landmark < exit`, for
that reason.

# Source and API checks

I read the function definitions in `R/` and their matching help files in `man/`
before setting up the analysis. I then loaded this checkout directly and
inspected the formal arguments from the active functions.

```{r api-signatures}
api_functions <- list(
  get_pbc_long = get_pbc_long,
  stack_data = stack_data,
  add_interactions = add_interactions,
  dynamic_lm = dynamic_lm,
  dynamic_lm_LMdataframe = getS3method("dynamic_lm", "LMdataframe"),
  predict_dynamicLM = getS3method("predict", "dynamicLM"),
  score = score,
  calplot = calplot
)

api_signatures <- data.frame(
  function_name = names(api_functions),
  arguments = vapply(api_functions, function(fun) {
    argument_names <- names(formals(fun))
    if (length(argument_names)) paste(argument_names, collapse = ", ") else "none"
  }, character(1L)),
  row.names = NULL
)

knitr::kable(api_signatures, col.names = c("Function", "Formal arguments"))
```

The source checks changed several details from the reference snippet.

```{r source-discrepancies}
source_discrepancies <- data.frame(
  item = c(
    "Longitudinal fields",
    "Interaction call",
    "Response formula",
    "Age-scale summary score",
    "Covariate parity",
    "Calibration identifier",
    "Calibration evaluation"
  ),
  source_finding = c(
    "get_pbc_long() also constructs chol; the reference list omitted it.",
    "lm_covs is the second add_interactions() argument; omitting it interacts every tracked covariate.",
    "Hist/CSC and Surv/coxph are supported, but the response parser rejects namespace-qualified Hist or Surv calls.",
    "score(summary = TRUE) assumes the first landmark contains every eventual subject and stops for staggered age risk sets.",
    "Updated attained age belongs only in the enrollment-time model because age is the time axis in the attained-age model.",
    "calplot() defaults to ID, whereas the PBC identifier returned here is id.",
    "calplot() does not forward its caller environment when it evaluates captured arguments."
  ),
  analysis_decision = c(
    "I retained only bili and albumin as the requested updated laboratories.",
    "I named the two laboratory covariates explicitly.",
    "I used the unqualified Hist response with type = CSC for both models.",
    "I used score() for landmark values and score_left_truncated() for the age-scale arithmetic summary.",
    "I asserted equality of every non-age model term and reported age as the sole structural exception.",
    "I passed id_col = id explicitly.",
    "I supplied an already evaluated argument list through do.call()."
  ),
  row.names = NULL
)

knitr::kable(source_discrepancies)
```

## PBC data returned by the fork

I ran `get_pbc_long()` without modifying its complete-case restriction. The
object returned `time` and `status` for the outcome and `tstart` for the visit
time. Age was already measured in years.

```{r pbc-structure}
pbc_long <- get_pbc_long()
str(pbc_long)
```

```{r pbc-head}
knitr::kable(head(pbc_long), digits = 3)
```

```{r pbc-dimensions}
pbc_dimensions <- data.frame(
  rows = nrow(pbc_long),
  subjects = length(unique(pbc_long$id)),
  minimum_age = min(pbc_long$age),
  maximum_age = max(pbc_long$age),
  row.names = NULL
)

outcome_counts <- as.data.frame(table(pbc_long$status), stringsAsFactors = FALSE)
names(outcome_counts) <- c("status", "rows")

stopifnot(
  all(c("id", "time", "status", "tstart", "bili", "albumin") %in% names(pbc_long)),
  min(pbc_long$age) > 18,
  max(pbc_long$age) < 100
)

knitr::kable(pbc_dimensions, digits = 3)
knitr::kable(outcome_counts)
```

# Evidence that bilirubin and albumin were updated

I selected bilirubin and albumin only after checking their within-person
variation. The short extract below shows repeated measurements for subjects
with several visits.

```{r repeated-lab-examples}
visits_per_subject <- table(pbc_long$id)
example_ids <- as.integer(names(head(sort(visits_per_subject, decreasing = TRUE), 3L)))
example_visits <- do.call(rbind, lapply(example_ids, function(subject_id) {
  head(
    pbc_long[pbc_long$id == subject_id, c("id", "tstart", "bili", "albumin")],
    5L
  )
}))
rownames(example_visits) <- NULL

knitr::kable(example_visits, digits = 3)
```

I summarized both the number of visits and the within-person standard
deviations. A subject counted as changing when at least two distinct recorded
values were present.

```{r longitudinal-summaries}
visit_summary <- data.frame(
  statistic = c("Minimum", "First quartile", "Median", "Mean", "Third quartile", "Maximum"),
  visits = as.numeric(c(
    min(visits_per_subject),
    quantile(visits_per_subject, 0.25, names = FALSE, type = 8),
    median(visits_per_subject),
    mean(visits_per_subject),
    quantile(visits_per_subject, 0.75, names = FALSE, type = 8),
    max(visits_per_subject)
  )),
  row.names = NULL
)

summarize_within_person <- function(variable, label) {
  values_by_subject <- split(pbc_long[[variable]], pbc_long$id)
  within_sd <- vapply(values_by_subject, stats::sd, numeric(1L))
  data.frame(
    laboratory = label,
    subjects = length(values_by_subject),
    subjects_with_multiple_visits = sum(lengths(values_by_subject) > 1L),
    subjects_with_change = sum(vapply(
      values_by_subject,
      function(values) length(unique(values)) > 1L,
      logical(1L)
    )),
    median_within_person_sd = median(within_sd, na.rm = TRUE),
    first_quartile_within_person_sd = quantile(within_sd, 0.25, na.rm = TRUE, names = FALSE, type = 8),
    third_quartile_within_person_sd = quantile(within_sd, 0.75, na.rm = TRUE, names = FALSE, type = 8),
    row.names = NULL
  )
}

laboratory_variation <- rbind(
  summarize_within_person("bili", "Bilirubin"),
  summarize_within_person("albumin", "Albumin")
)

stopifnot(all(laboratory_variation$subjects_with_change > 0L))

knitr::kable(visit_summary, digits = 2)
knitr::kable(laboratory_variation, digits = 3)
```

# Common analysis definitions

I retained the original three-category endpoint: status zero was censoring,
status one was transplant, and status two was death. I fitted a cause-specific
Cox supermodel through `Hist(..., entry)` and `type = "CSC"`, with death as the
event of interest and transplant retained as the competing event. This avoided
changing the outcome when I changed the time axis.

I log-transformed bilirubin because of its right-skewed scale and standardized
both laboratory measures once, before constructing either landmark stack. The
same transformed values therefore entered both models.

```{r analysis-definitions}
prediction_window <- 5
cause_of_interest <- 2L
enrollment_landmarks <- seq(0, 4, by = 1)

log_bili_center <- mean(log(pbc_long$bili))
log_bili_scale <- sd(log(pbc_long$bili))
albumin_center <- mean(pbc_long$albumin)
albumin_scale <- sd(pbc_long$albumin)

subject_entry_ages <- unique(pbc_long[c("id", "age")])
age_center <- mean(subject_entry_ages$age)
age_scale <- sd(subject_entry_ages$age)

pbc_analysis <- transform(
  pbc_long,
  age_at_enrollment = age,
  log_bili_z = (log(bili) - log_bili_center) / log_bili_scale,
  albumin_z = (albumin - albumin_center) / albumin_scale
)

shared_fixed_covariates <- c("male", "stage", "trt")
updated_laboratory_covariates <- c("log_bili_z", "albumin_z")
shared_model_covariates <- c(shared_fixed_covariates, updated_laboratory_covariates)
enrollment_model_covariates <- c("attained_age_z", shared_model_covariates)
age_model_covariates <- shared_model_covariates

stopifnot(
  identical(
    setdiff(enrollment_model_covariates, "attained_age_z"),
    age_model_covariates
  ),
  all(is.finite(pbc_analysis$log_bili_z)),
  all(is.finite(pbc_analysis$albumin_z))
)

analysis_constants <- data.frame(
  quantity = c("Prediction window", "Cause of interest", "Enrollment landmarks"),
  value = c(
    paste(prediction_window, "years"),
    paste(cause_of_interest, "(death)"),
    paste(enrollment_landmarks, collapse = ", ")
  ),
  row.names = NULL
)

covariate_contract <- data.frame(
  scale = c("Enrollment time", "Attained age"),
  non_age_covariates = c(
    paste(shared_model_covariates, collapse = ", "),
    paste(shared_model_covariates, collapse = ", ")
  ),
  age_treatment = c(
    "Updated attained age entered as a predictor",
    "Attained age was the time axis and was excluded from the right-hand side"
  ),
  row.names = NULL
)

knitr::kable(analysis_constants)
knitr::kable(covariate_contract)
```

# Landmark selection and risk sets

I chose the age landmarks from the empirical first, second, and third quartiles
of entry age and rounded them to one decimal place. This covered the central
span of entry ages without selecting round ages first and, as the risk-set table
shows, retained deaths and horizon controls at each landmark.

```{r age-landmark-selection}
age_landmark_probabilities <- c(0.25, 0.50, 0.75)
age_landmarks <- round(
  unname(quantile(
    subject_entry_ages$age,
    age_landmark_probabilities,
    type = 8
  )),
  1
)

age_landmark_selection <- data.frame(
  entry_age_percentile = paste0(age_landmark_probabilities * 100, "%"),
  landmark_age = age_landmarks,
  row.names = NULL
)

stopifnot(length(unique(age_landmarks)) == length(age_landmarks))
knitr::kable(age_landmark_selection, digits = 1)
```

For enrollment time, I carried each person's baseline age forward by the
landmark time. For attained age, I instead shifted the outcome and visit times
onto the age scale and passed entry age to the strict stacker. I did not add age
to the attained-age model's right-hand side.

```{r build-stacks}
enrollment_stack <- stack_data(
  data = pbc_analysis,
  outcome = list(time = "time", status = "status"),
  lms = enrollment_landmarks,
  w = prediction_window,
  covs = list(
    fixed = c(shared_fixed_covariates, "age_at_enrollment"),
    varying = updated_laboratory_covariates
  ),
  format = "long",
  id = "id",
  rtime = "tstart"
)

enrollment_stack$data$attained_age_z <- (
  enrollment_stack$data$age_at_enrollment + enrollment_stack$data$LM - age_center
) / age_scale

pbc_age <- transform(
  pbc_analysis,
  entry_age = age_at_enrollment,
  exit_age = age_at_enrollment + time,
  visit_age = age_at_enrollment + tstart
)

age_stack <- stack_data(
  data = pbc_age,
  outcome = list(time = "exit_age", status = "status"),
  lms = age_landmarks,
  w = prediction_window,
  covs = list(
    fixed = shared_fixed_covariates,
    varying = updated_laboratory_covariates
  ),
  format = "long",
  id = "id",
  rtime = "visit_age",
  entry = "entry_age",
  entry_mode = "strict"
)

stopifnot(
  identical(age_stack$entry_mode, "strict"),
  identical(age_stack$entry_col, "LM"),
  all(age_stack$data$entry_age <= age_stack$data$LM),
  all(age_stack$data$LM < age_stack$data$exit_age),
  !anyDuplicated(age_stack$data[c("id", "LM")]),
  !anyDuplicated(enrollment_stack$data[c("id", "LM")])
)
```

```{r risk-set-table}
summarize_risk_sets <- function(stack, scale_label, time_column) {
  do.call(rbind, lapply(split(stack$data, stack$data$LM), function(landmark_data) {
    landmark <- unique(landmark_data$LM)
    stopifnot(length(landmark) == 1L)
    data.frame(
      scale = scale_label,
      landmark = landmark,
      subjects = length(unique(landmark_data$id)),
      deaths = sum(landmark_data$status == cause_of_interest),
      transplants = sum(landmark_data$status == 1L),
      horizon_controls = sum(
        landmark_data$status == 0L &
          landmark_data[[time_column]] >= landmark + prediction_window
      ),
      censorings_before_horizon = sum(
        landmark_data$status == 0L &
          landmark_data[[time_column]] < landmark + prediction_window
      ),
      row.names = NULL
    )
  }))
}

risk_sets <- rbind(
  summarize_risk_sets(enrollment_stack, "Enrollment time", "time"),
  summarize_risk_sets(age_stack, "Attained age", "exit_age")
)

enrollment_risk_counts <- risk_sets$subjects[risk_sets$scale == "Enrollment time"]
age_risk_counts <- risk_sets$subjects[risk_sets$scale == "Attained age"]

stopifnot(
  all(diff(enrollment_risk_counts) <= 0L),
  all(risk_sets$deaths <= risk_sets$subjects),
  all(risk_sets$transplants <= risk_sets$subjects)
)

knitr::kable(risk_sets, digits = 1)
```

The enrollment risk sets decreased monotonically. The attained-age risk sets
did not have to decrease: later landmarks lost people to events and censoring
but could also gain people who entered observation at older ages. This is why a
monotonicity check that is sensible after a shared origin is not an age-scale
eligibility check.

# Model fitting

I allowed the effects of the two updated laboratory covariates to vary
quadratically over each scale. I centered and scaled the landmark basis within
each analysis rather than putting raw attained-age polynomials in the model.
I did not include a standalone landmark main effect because the unspecified
baseline hazard already varies over analysis time.

```{r add-landmark-interactions}
make_landmark_basis <- function(center, half_range) {
  force(center)
  force(half_range)
  list(
    function(time) (time - center) / half_range,
    function(time) ((time - center) / half_range)^2
  )
}

enrollment_basis <- make_landmark_basis(
  mean(enrollment_landmarks),
  diff(range(enrollment_landmarks)) / 2
)
age_basis <- make_landmark_basis(
  mean(age_landmarks),
  diff(range(age_landmarks)) / 2
)

enrollment_stack <- add_interactions(
  enrollment_stack,
  lm_covs = updated_laboratory_covariates,
  func_covars = enrollment_basis,
  func_lms = list()
)
age_stack <- add_interactions(
  age_stack,
  lm_covs = updated_laboratory_covariates,
  func_covars = age_basis,
  func_lms = list()
)

interaction_names <- c(
  "log_bili_z_LM1", "log_bili_z_LM2",
  "albumin_z_LM1", "albumin_z_LM2"
)

stopifnot(
  all(interaction_names %in% names(enrollment_stack$data)),
  all(interaction_names %in% names(age_stack$data))
)

interaction_columns <- data.frame(
  updated_covariate = updated_laboratory_covariates,
  linear_interaction = paste0(updated_laboratory_covariates, "_LM1"),
  quadratic_interaction = paste0(updated_laboratory_covariates, "_LM2"),
  row.names = NULL
)

knitr::kable(interaction_columns)
```

I used the same cause-specific outcome specification on both scales. Repeated
landmark records were clustered by subject, so the standard errors below are
the robust values returned by the cause-specific Cox fits.

```{r fit-models}
enrollment_formula <- as.formula(
  paste(
    "Hist(time, status, LM) ~",
    "attained_age_z + male + stage + trt +",
    "log_bili_z + log_bili_z_LM1 + log_bili_z_LM2 +",
    "albumin_z + albumin_z_LM1 + albumin_z_LM2 + cluster(id)"
  )
)

age_formula <- as.formula(
  paste(
    "Hist(exit_age, status, LM) ~",
    "male + stage + trt +",
    "log_bili_z + log_bili_z_LM1 + log_bili_z_LM2 +",
    "albumin_z + albumin_z_LM1 + albumin_z_LM2 + cluster(id)"
  )
)

capture_conditions <- function(expression) {
  warnings <- character()
  messages <- character()
  value <- withCallingHandlers(
    tryCatch(eval.parent(substitute(expression)), error = function(error) error),
    warning = function(warning) {
      warnings <<- c(warnings, conditionMessage(warning))
      invokeRestart("muffleWarning")
    },
    message = function(message) {
      messages <<- c(messages, conditionMessage(message))
      invokeRestart("muffleMessage")
    }
  )
  list(value = value, warnings = unique(warnings), messages = unique(messages))
}

enrollment_fit_record <- capture_conditions(dynamic_lm(
  enrollment_stack,
  enrollment_formula,
  type = "CSC",
  method = "breslow",
  x = TRUE
))
age_fit_record <- capture_conditions(dynamic_lm(
  age_stack,
  age_formula,
  type = "CSC",
  method = "breslow",
  x = TRUE
))

fit_condition_table <- data.frame(
  scale = c("Enrollment time", "Attained age"),
  warning_count = c(
    length(enrollment_fit_record$warnings),
    length(age_fit_record$warnings)
  ),
  error = c(
    if (inherits(enrollment_fit_record$value, "error")) conditionMessage(enrollment_fit_record$value) else "",
    if (inherits(age_fit_record$value, "error")) conditionMessage(age_fit_record$value) else ""
  ),
  row.names = NULL
)

knitr::kable(fit_condition_table)

stopifnot(
  !inherits(enrollment_fit_record$value, "error"),
  !inherits(age_fit_record$value, "error"),
  length(enrollment_fit_record$warnings) == 0L,
  length(age_fit_record$warnings) == 0L
)

enrollment_fit <- enrollment_fit_record$value
age_fit <- age_fit_record$value
```

## Shared coefficient table

The interaction bases are centered on different clocks, so the interaction
coefficients are not estimates of a common parameter. I placed them together to
show the fitted models, not to interpret their differences as contrasts.

```{r coefficient-table}
extract_cause_fit <- function(supermodel, cause) {
  cause_index <- match(cause, supermodel$model$causes)
  stopifnot(!is.na(cause_index))
  supermodel$model$models[[cause_index]]
}

extract_coefficients <- function(supermodel, scale_label) {
  cause_fit <- extract_cause_fit(supermodel, cause_of_interest)
  robust_standard_errors <- sqrt(diag(vcov(cause_fit)))
  data.frame(
    scale = scale_label,
    term = names(coef(cause_fit)),
    estimate = unname(coef(cause_fit)),
    robust_standard_error = unname(robust_standard_errors),
    row.names = NULL
  )
}

enrollment_coefficients <- extract_coefficients(enrollment_fit, "Enrollment time")
age_coefficients <- extract_coefficients(age_fit, "Attained age")

shared_terms <- c(
  "male", "stage", "trt",
  "log_bili_z", "log_bili_z_LM1", "log_bili_z_LM2",
  "albumin_z", "albumin_z_LM1", "albumin_z_LM2"
)

stopifnot(
  all(shared_terms %in% enrollment_coefficients$term),
  all(shared_terms %in% age_coefficients$term),
  identical(
    enrollment_coefficients$term[enrollment_coefficients$term %in% shared_terms],
    age_coefficients$term[age_coefficients$term %in% shared_terms]
  )
)

coefficient_comparison <- merge(
  enrollment_coefficients[enrollment_coefficients$term %in% shared_terms, ],
  age_coefficients[age_coefficients$term %in% shared_terms, ],
  by = "term",
  suffixes = c("_enrollment", "_age"),
  sort = FALSE
)
coefficient_comparison <- coefficient_comparison[match(shared_terms, coefficient_comparison$term), ]
coefficient_comparison <- coefficient_comparison[c(
  "term",
  "estimate_enrollment", "robust_standard_error_enrollment",
  "estimate_age", "robust_standard_error_age"
)]

knitr::kable(coefficient_comparison, digits = 4)
```

## Dynamic laboratory effects

The curves below show the fitted log hazard ratio for a one-standard-deviation
increase in each laboratory value as the landmark changes. The bands use the
robust covariance matrix from the corresponding cause-specific fit.

```{r dynamic-log-hazard-ratios, fig.width=10, fig.height=8}
dynamic_log_hazard_ratio <- function(supermodel, marker, landmarks, basis) {
  cause_fit <- extract_cause_fit(supermodel, cause_of_interest)
  marker_terms <- c(marker, paste0(marker, "_LM1"), paste0(marker, "_LM2"))
  stopifnot(all(marker_terms %in% names(coef(cause_fit))))
  design <- cbind(
    1,
    basis[[1L]](landmarks),
    basis[[2L]](landmarks)
  )
  coefficients <- coef(cause_fit)[marker_terms]
  covariance <- vcov(cause_fit)[marker_terms, marker_terms, drop = FALSE]
  estimate <- as.numeric(design %*% coefficients)
  standard_error <- sqrt(rowSums((design %*% covariance) * design))
  data.frame(
    landmark = landmarks,
    estimate = estimate,
    lower = estimate - qnorm(0.975) * standard_error,
    upper = estimate + qnorm(0.975) * standard_error,
    row.names = NULL
  )
}

plot_dynamic_effect <- function(curve, title, x_label) {
  plot(
    curve$landmark,
    curve$estimate,
    type = "n",
    xlab = x_label,
    ylab = "Log hazard ratio per SD",
    main = title,
    ylim = range(curve$lower, curve$upper)
  )
  polygon(
    c(curve$landmark, rev(curve$landmark)),
    c(curve$lower, rev(curve$upper)),
    border = NA,
    col = "grey85"
  )
  abline(h = 0, lty = 2, col = "grey45")
  lines(curve$landmark, curve$estimate, lwd = 2)
}

enrollment_grid <- seq(
  min(enrollment_landmarks),
  max(enrollment_landmarks),
  length.out = 200L
)
age_grid <- seq(min(age_landmarks), max(age_landmarks), length.out = 200L)

dynamic_curves <- list(
  enrollment_bili = dynamic_log_hazard_ratio(
    enrollment_fit, "log_bili_z", enrollment_grid, enrollment_basis
  ),
  age_bili = dynamic_log_hazard_ratio(
    age_fit, "log_bili_z", age_grid, age_basis
  ),
  enrollment_albumin = dynamic_log_hazard_ratio(
    enrollment_fit, "albumin_z", enrollment_grid, enrollment_basis
  ),
  age_albumin = dynamic_log_hazard_ratio(
    age_fit, "albumin_z", age_grid, age_basis
  )
)

old_parameters <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))
plot_dynamic_effect(
  dynamic_curves$enrollment_bili,
  "Bilirubin: enrollment time",
  "Years after enrollment"
)
plot_dynamic_effect(
  dynamic_curves$age_bili,
  "Bilirubin: attained age",
  "Attained age"
)
plot_dynamic_effect(
  dynamic_curves$enrollment_albumin,
  "Albumin: enrollment time",
  "Years after enrollment"
)
plot_dynamic_effect(
  dynamic_curves$age_albumin,
  "Albumin: attained age",
  "Attained age"
)
par(old_parameters)
```

# Apparent predictive performance

I evaluated the same fitted records used for model estimation, so these are
apparent rather than externally validated performance estimates. I used
`score()` directly for the enrollment summary and for the landmark-specific
values on both scales.

The current summary implementation assumes that the first landmark contains
every subject who appears later. That is true after a shared enrollment origin
but false under staggered age entry. I retained the exact age-scale failure
below and used `score_left_truncated()` for the arithmetic age summary. The
analytic staggered-entry influence function remains unavailable.

```{r predictions-and-scores}
enrollment_prediction_record <- capture_conditions(predict(
  enrollment_fit,
  cause = cause_of_interest
))
age_prediction_record <- capture_conditions(predict(
  age_fit,
  cause = cause_of_interest
))

stopifnot(
  !inherits(enrollment_prediction_record$value, "error"),
  !inherits(age_prediction_record$value, "error"),
  length(enrollment_prediction_record$warnings) == 0L,
  length(age_prediction_record$warnings) == 0L
)

enrollment_predictions <- enrollment_prediction_record$value
age_predictions <- age_prediction_record$value

enrollment_score_record <- capture_conditions(score(
  list("Enrollment scale" = enrollment_predictions),
  times = enrollment_landmarks,
  metrics = c("auc", "brier"),
  id_col = "id",
  cause = cause_of_interest,
  summary = TRUE,
  contrasts = FALSE
))

age_landmark_score_record <- capture_conditions(score(
  list("Age scale" = age_predictions),
  times = age_landmarks,
  metrics = c("auc", "brier"),
  id_col = "id",
  cause = cause_of_interest,
  summary = FALSE,
  contrasts = FALSE
))

age_summary_failure <- capture_conditions(score(
  list("Age scale" = age_predictions),
  times = age_landmarks,
  metrics = c("auc", "brier"),
  id_col = "id",
  cause = cause_of_interest,
  summary = TRUE,
  contrasts = FALSE
))

age_left_truncated_score_record <- capture_conditions(score_left_truncated(
  risk = setNames(data.frame(age_predictions$preds$risk), "Age scale"),
  data = age_predictions$data,
  time_col = "exit_age",
  status_col = "status",
  entry_col = "LM",
  landmark_col = "LM",
  id_col = "id",
  cause = cause_of_interest,
  w = prediction_window
))

score_condition_table <- data.frame(
  operation = c(
    "Enrollment score with summary",
    "Age score without summary",
    "Age score with summary",
    "Age left-truncated point summary"
  ),
  warning_count = c(
    length(enrollment_score_record$warnings),
    length(age_landmark_score_record$warnings),
    length(age_summary_failure$warnings),
    length(age_left_truncated_score_record$warnings)
  ),
  error = c(
    if (inherits(enrollment_score_record$value, "error")) conditionMessage(enrollment_score_record$value) else "",
    if (inherits(age_landmark_score_record$value, "error")) conditionMessage(age_landmark_score_record$value) else "",
    if (inherits(age_summary_failure$value, "error")) conditionMessage(age_summary_failure$value) else "",
    if (inherits(age_left_truncated_score_record$value, "error")) conditionMessage(age_left_truncated_score_record$value) else ""
  ),
  row.names = NULL
)

knitr::kable(score_condition_table)

stopifnot(
  !inherits(enrollment_score_record$value, "error"),
  !inherits(age_landmark_score_record$value, "error"),
  inherits(age_summary_failure$value, "error"),
  identical(conditionMessage(age_summary_failure$value), "Subsample size mismatch."),
  !inherits(age_left_truncated_score_record$value, "error")
)

enrollment_scores <- enrollment_score_record$value
age_landmark_scores <- age_landmark_score_record$value
age_left_truncated_scores <- age_left_truncated_score_record$value
```

The two scoring paths returned the same age-scale point estimates at every
landmark. This check does not supply the missing analytic summary variance.

```{r age-score-agreement}
ordinary_age_auc <- age_landmark_scores$AUC$score
ordinary_age_auc <- ordinary_age_auc[ordinary_age_auc$model == "Age scale", ]
ordinary_age_auc <- ordinary_age_auc[order(ordinary_age_auc$tLM), ]

ordinary_age_brier <- age_landmark_scores$Brier$score
ordinary_age_brier <- ordinary_age_brier[ordinary_age_brier$model == "Age scale", ]
ordinary_age_brier <- ordinary_age_brier[order(ordinary_age_brier$tLM), ]

left_truncated_landmark_scores <- age_left_truncated_scores$score
left_truncated_landmark_scores <- left_truncated_landmark_scores[
  order(left_truncated_landmark_scores$landmark),
]

score_agreement <- data.frame(
  metric = c("AUC", "Brier"),
  maximum_absolute_difference = c(
    max(abs(ordinary_age_auc$AUC - left_truncated_landmark_scores$AUC)),
    max(abs(ordinary_age_brier$Brier - left_truncated_landmark_scores$Brier))
  ),
  row.names = NULL
)

stopifnot(all(score_agreement$maximum_absolute_difference < 1e-12))

knitr::kable(score_agreement, digits = 16)
knitr::kable(left_truncated_landmark_scores, digits = 4)
```

```{r summary-performance}
enrollment_auc_summary <- enrollment_scores$AUC_summary$score
enrollment_auc_summary <- enrollment_auc_summary[
  enrollment_auc_summary$model == "Enrollment scale",
]
enrollment_brier_summary <- enrollment_scores$Brier_summary$score
enrollment_brier_summary <- enrollment_brier_summary[
  enrollment_brier_summary$model == "Enrollment scale",
]
age_summary <- age_left_truncated_scores$summary

summary_performance <- data.frame(
  scale = c("Enrollment time", "Attained age"),
  summary_auc = c(enrollment_auc_summary$AUC, age_summary$AUC),
  summary_auc_standard_error = c(enrollment_auc_summary$se, NA_real_),
  summary_brier = c(enrollment_brier_summary$Brier, age_summary$Brier),
  summary_brier_standard_error = c(enrollment_brier_summary$se, NA_real_),
  inference = c(
    "Original analytic summary calculation",
    age_left_truncated_scores$analytic_influence_function
  ),
  row.names = NULL
)

knitr::kable(summary_performance, digits = 4)
```

The summary values answer different prediction questions for different
landmark populations. I therefore treated this table as a descriptive
time-scale comparison, not as a test that one clock performed better.

# Calibration

The apparent calibration plots rendered at every requested landmark. I used
the same quantile grouping on both scales and retained the condition table so a
future failure will be visible rather than silently dropped.

```{r enrollment-calibration, fig.width=10, fig.height=7}
old_parameters <- par(mfrow = c(2, 3), mar = c(4, 4, 3, 1))
enrollment_calibration_arguments <- list(
  object = list("Enrollment scale" = enrollment_predictions),
  times = enrollment_landmarks,
  id_col = "id",
  cause = cause_of_interest,
  method = "quantile",
  q = 4,
  legend = FALSE,
  xlim = c(0, 1),
  ylim = c(0, 1)
)
enrollment_calibration_record <- capture_conditions(do.call(
  calplot,
  enrollment_calibration_arguments
))
par(old_parameters)

if (inherits(enrollment_calibration_record$value, "error")) {
  stop(
    "Enrollment calibration failed: ",
    conditionMessage(enrollment_calibration_record$value)
  )
}
```

```{r age-calibration, fig.width=10, fig.height=4}
old_parameters <- par(mfrow = c(1, 3), mar = c(4, 4, 3, 1))
age_calibration_arguments <- list(
  object = list("Age scale" = age_predictions),
  times = age_landmarks,
  id_col = "id",
  cause = cause_of_interest,
  method = "quantile",
  q = 4,
  legend = FALSE,
  xlim = c(0, 1),
  ylim = c(0, 1)
)
age_calibration_record <- capture_conditions(do.call(
  calplot,
  age_calibration_arguments
))
par(old_parameters)

if (inherits(age_calibration_record$value, "error")) {
  stop(
    "Age calibration failed: ",
    conditionMessage(age_calibration_record$value)
  )
}
```

```{r calibration-conditions}
calibration_conditions <- data.frame(
  scale = c("Enrollment time", "Attained age"),
  warning_count = c(
    length(enrollment_calibration_record$warnings),
    length(age_calibration_record$warnings)
  ),
  error = c(
    if (inherits(enrollment_calibration_record$value, "error")) conditionMessage(enrollment_calibration_record$value) else "",
    if (inherits(age_calibration_record$value, "error")) conditionMessage(age_calibration_record$value) else ""
  ),
  row.names = NULL
)

knitr::kable(calibration_conditions)
```

# Interpretation

The enrollment analysis used a common origin and updated attained age as a
predictor. The attained-age analysis used age as the clock, excluded age from
the right-hand side, and enforced observation by each landmark. The remaining
predictors and the two laboratory interaction structures were identical.

The age-scale risk-set pattern illustrated the practical difference: people
could enter as age advanced, so a later age landmark could contain more eligible
subjects than an earlier one. Strict entry made that membership rule explicit.

Within-window delayed entry was not part of this primary comparison. In the
separate NAFLD application, it added very little analyzable longitudinal
information because most later entrants lacked a laboratory value at their
analysis entry. That finding supports treating strict entry as the clearer
primary analysis rather than assuming delayed entry will automatically recover
substantial usable data.

# Session information

```{r session-information}
sessionInfo()
```
