---
title: "PBC landmark supermodels on enrollment and attained-age time scales"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  out.width = "95%"
)
options(width = 110)

prediction_window <- 5
cause_of_interest <- 2L
transplant_code <- 1L
censor_code <- 0L
enrollment_landmarks <- seq(0, 4, by = 1)
confidence_level <- 0.95
n_boot <- getOption("timescale.n_boot", 200L)
bootstrap_seed <- 20260922L
diagnostic_pilot_replicates <- 50L

stopifnot(
  is.numeric(n_boot), length(n_boot) == 1L,
  is.finite(n_boot), n_boot >= 2L, n_boot == as.integer(n_boot)
)
n_boot <- as.integer(n_boot)

library(here)
devtools::load_all(here::here(), quiet = TRUE)
```

# Changes in this version

The version previously reviewed omitted landmark main effects from both
supermodels. I added the centered linear and quadratic landmark terms to the
enrollment-time supermodel. I kept them out of the primary attained-age model
after checking that its prediction windows do not overlap and demonstrating
that the Cox partial likelihood cannot estimate those terms on this grid.

I also replaced the coefficient display with robust hazard ratios, confidence
intervals, and Wald tests; fitted a separate cause-specific model at every
landmark; and compared those disjoint fits with the supermodels in coefficients,
precision, and prediction performance. The performance section now separates
apparent estimates from subject-level Harrell-bootstrap optimism correction.

# Purpose

I compared the PBC landmark supermodel on two clocks. The enrollment-time
analysis asked about risk a given number of years after enrollment. The
attained-age analysis asked about risk at a given age. Both used a
`r prediction_window`-year prediction window, the same PBC records, outcome,
laboratory transformations, and non-age predictors.

These are not competing estimates of the same quantity. An enrollment landmark
conditions on survival for a specified time after enrollment; an age landmark
conditions on being observed and event-free at a specified age. The risk sets
and scientific questions therefore differ. A difference between the fitted
models is not bias in either direction.

The age scale is useful when a common time zero has no defensible clinical
meaning, as in a cohort assembled from electronic health records. It also
requires an additional eligibility rule: a person must have entered observation
by the landmark age. I used strict eligibility, `entry <= landmark < exit`, for
that reason.

# Source and API checks

I read the function definitions in `R/` and their matching help files in `man/`
before setting up the analysis. I then loaded this checkout directly and
inspected the formal arguments from the active functions.

```{r api-signatures}
api_functions <- list(
  get_pbc_long = get_pbc_long,
  stack_data = stack_data,
  add_interactions = add_interactions,
  dynamic_lm = dynamic_lm,
  dynamic_lm_LMdataframe = getS3method("dynamic_lm", "LMdataframe"),
  predict_dynamicLM = getS3method("predict", "dynamicLM"),
  score = score,
  score_left_truncated = score_left_truncated,
  calplot = calplot
)

api_signatures <- data.frame(
  function_name = names(api_functions),
  arguments = vapply(api_functions, function(fun) {
    argument_names <- names(formals(fun))
    if (length(argument_names)) paste(argument_names, collapse = ", ") else "none"
  }, character(1L)),
  row.names = NULL
)

knitr::kable(api_signatures, col.names = c("Function", "Formal arguments"))
```

The source checks changed several details from the reference snippet.

```{r source-discrepancies}
source_discrepancies <- data.frame(
  item = c(
    "Longitudinal fields",
    "Interaction call",
    "Response formula",
    "Age-scale summary score",
    "Covariate parity",
    "Calibration identifier",
    "Calibration evaluation"
  ),
  source_finding = c(
    "get_pbc_long() also constructs chol; the reference list omitted it.",
    "lm_covs is the second add_interactions() argument; omitting it interacts every tracked covariate.",
    "Hist/CSC and Surv/coxph are supported, but the response parser rejects namespace-qualified Hist or Surv calls.",
    "score(summary = TRUE) assumes the first landmark contains every eventual subject and stops for staggered age risk sets.",
    "Updated attained age belongs only in the enrollment-time model; the current age grid also cannot identify landmark main effects.",
    "calplot() defaults to ID, whereas the PBC identifier returned here is id.",
    "calplot() does not forward its caller environment when it evaluates captured arguments."
  ),
  analysis_decision = c(
    "I retained only bili and albumin as the requested updated laboratories.",
    "I named the two laboratory covariates explicitly.",
    "I used the unqualified Hist response with type = CSC for both models.",
    "I used score() for landmark values and score_left_truncated() for the age-scale arithmetic summary.",
    "I asserted equality of the substantive non-age predictors and reported both structural exceptions explicitly.",
    "I passed id_col = id explicitly.",
    "I supplied an already evaluated argument list through do.call()."
  ),
  row.names = NULL
)

knitr::kable(source_discrepancies)
```

## PBC data returned by the fork

I ran `get_pbc_long()` without modifying its complete-case restriction. The
object returned `time` and `status` for the outcome and `tstart` for the visit
time. Age was already measured in years.

```{r pbc-structure}
pbc_long <- get_pbc_long()
str(pbc_long)
```

```{r pbc-head}
knitr::kable(head(pbc_long), digits = 3)
```

```{r pbc-dimensions}
pbc_dimensions <- data.frame(
  rows = nrow(pbc_long),
  subjects = length(unique(pbc_long$id)),
  minimum_age = min(pbc_long$age),
  maximum_age = max(pbc_long$age),
  row.names = NULL
)

outcome_counts <- as.data.frame(table(pbc_long$status), stringsAsFactors = FALSE)
names(outcome_counts) <- c("status", "rows")

stopifnot(
  all(c("id", "time", "status", "tstart", "bili", "albumin") %in% names(pbc_long)),
  min(pbc_long$age) > 18,
  max(pbc_long$age) < 100
)

knitr::kable(pbc_dimensions, digits = 3)
knitr::kable(outcome_counts)
```

# Evidence that bilirubin and albumin were updated

I selected bilirubin and albumin only after checking their within-person
variation. The short extract below shows repeated measurements for subjects
with several visits.

```{r repeated-lab-examples}
visits_per_subject <- table(pbc_long$id)
example_ids <- as.integer(names(head(sort(visits_per_subject, decreasing = TRUE), 3L)))
example_visits <- do.call(rbind, lapply(example_ids, function(subject_id) {
  head(
    pbc_long[pbc_long$id == subject_id, c("id", "tstart", "bili", "albumin")],
    5L
  )
}))
rownames(example_visits) <- NULL

knitr::kable(example_visits, digits = 3)
```

I summarized both the number of visits and the within-person standard
deviations. A subject counted as changing when at least two distinct recorded
values were present.

```{r longitudinal-summaries}
visit_summary <- data.frame(
  statistic = c("Minimum", "First quartile", "Median", "Mean", "Third quartile", "Maximum"),
  visits = as.numeric(c(
    min(visits_per_subject),
    quantile(visits_per_subject, 0.25, names = FALSE, type = 8),
    median(visits_per_subject),
    mean(visits_per_subject),
    quantile(visits_per_subject, 0.75, names = FALSE, type = 8),
    max(visits_per_subject)
  )),
  row.names = NULL
)

summarize_within_person <- function(variable, label) {
  values_by_subject <- split(pbc_long[[variable]], pbc_long$id)
  within_sd <- vapply(values_by_subject, stats::sd, numeric(1L))
  data.frame(
    laboratory = label,
    subjects = length(values_by_subject),
    subjects_with_multiple_visits = sum(lengths(values_by_subject) > 1L),
    subjects_with_change = sum(vapply(
      values_by_subject,
      function(values) length(unique(values)) > 1L,
      logical(1L)
    )),
    median_within_person_sd = median(within_sd, na.rm = TRUE),
    first_quartile_within_person_sd = quantile(within_sd, 0.25, na.rm = TRUE, names = FALSE, type = 8),
    third_quartile_within_person_sd = quantile(within_sd, 0.75, na.rm = TRUE, names = FALSE, type = 8),
    row.names = NULL
  )
}

laboratory_variation <- rbind(
  summarize_within_person("bili", "Bilirubin"),
  summarize_within_person("albumin", "Albumin")
)

stopifnot(all(laboratory_variation$subjects_with_change > 0L))

knitr::kable(visit_summary, digits = 2)
knitr::kable(laboratory_variation, digits = 3)
```

# Common analysis definitions

I retained the original endpoint: status `r censor_code` was censoring, status
`r transplant_code` was transplant, and status `r cause_of_interest` was death.
I fitted a cause-specific Cox supermodel through `Hist(..., entry)` and
`type = "CSC"`, with death as the event of interest and transplant retained as
the competing event. This avoided changing the outcome when I changed the time
axis.

I log-transformed bilirubin because of its right-skewed scale and standardized
both laboratory measures once, before constructing either landmark stack. The
same transformed values therefore entered both models.

```{r analysis-definitions}
log_bili_center <- mean(log(pbc_long$bili))
log_bili_scale <- sd(log(pbc_long$bili))
albumin_center <- mean(pbc_long$albumin)
albumin_scale <- sd(pbc_long$albumin)

subject_entry_ages <- unique(pbc_long[c("id", "age")])
age_center <- mean(subject_entry_ages$age)
age_scale <- sd(subject_entry_ages$age)

pbc_analysis <- transform(
  pbc_long,
  age_at_enrollment = age,
  log_bili_z = (log(bili) - log_bili_center) / log_bili_scale,
  albumin_z = (albumin - albumin_center) / albumin_scale
)

shared_fixed_covariates <- c("male", "stage", "trt")
updated_laboratory_covariates <- c("log_bili_z", "albumin_z")
shared_model_covariates <- c(shared_fixed_covariates, updated_laboratory_covariates)
enrollment_model_covariates <- c("attained_age_z", shared_model_covariates)
age_model_covariates <- shared_model_covariates

stopifnot(
  identical(
    setdiff(enrollment_model_covariates, "attained_age_z"),
    age_model_covariates
  ),
  all(is.finite(pbc_analysis$log_bili_z)),
  all(is.finite(pbc_analysis$albumin_z))
)

make_landmark_basis <- function(center, half_range) {
  stopifnot(is.finite(center), is.finite(half_range), half_range > 0)
  force(center)
  force(half_range)
  list(
    function(time) (time - center) / half_range,
    function(time) ((time - center) / half_range)^2
  )
}

analysis_constants <- data.frame(
  quantity = c("Prediction window", "Cause of interest", "Enrollment landmarks"),
  value = c(
    paste(prediction_window, "years"),
    paste(cause_of_interest, "(death)"),
    paste(enrollment_landmarks, collapse = ", ")
  ),
  row.names = NULL
)

covariate_contract <- data.frame(
  scale = c("Enrollment time", "Attained age"),
  non_age_covariates = c(
    paste(shared_model_covariates, collapse = ", "),
    paste(shared_model_covariates, collapse = ", ")
  ),
  age_treatment = c(
    "Updated attained age entered as a predictor",
    "Attained age was the time axis and was excluded from the right-hand side"
  ),
  landmark_main_effect = c(
    "Centered linear and quadratic LM terms",
    "Omitted because the non-overlapping windows make LM constant within every risk set"
  ),
  row.names = NULL
)

knitr::kable(analysis_constants)
knitr::kable(covariate_contract)
```

# Landmark selection and risk sets

I chose the age landmarks from the empirical first, second, and third quartiles
of entry age and rounded them to one decimal place. This covered the central
span of entry ages without selecting round ages first and, as the risk-set table
shows, retained deaths and horizon controls at each landmark.

```{r age-landmark-selection}
age_landmark_probabilities <- c(0.25, 0.50, 0.75)
age_landmarks <- round(
  unname(quantile(
    subject_entry_ages$age,
    age_landmark_probabilities,
    type = 8
  )),
  1
)

age_landmark_selection <- data.frame(
  entry_age_percentile = paste0(age_landmark_probabilities * 100, "%"),
  landmark_age = age_landmarks,
  row.names = NULL
)

stopifnot(length(unique(age_landmarks)) == length(age_landmarks))
knitr::kable(age_landmark_selection, digits = 1)
```

```{r landmark-bases}
enrollment_basis <- make_landmark_basis(
  mean(enrollment_landmarks),
  diff(range(enrollment_landmarks)) / 2
)
age_basis <- make_landmark_basis(
  mean(age_landmarks),
  diff(range(age_landmarks)) / 2
)
```

For enrollment time, I carried each person's baseline age forward by the
landmark time. For attained age, I instead shifted the outcome and visit times
onto the age scale and passed entry age to the strict stacker. I did not add age
to the attained-age model's right-hand side.

```{r build-stacks}
enrollment_stack_raw <- stack_data(
  data = pbc_analysis,
  outcome = list(time = "time", status = "status"),
  lms = enrollment_landmarks,
  w = prediction_window,
  covs = list(
    fixed = c(shared_fixed_covariates, "age_at_enrollment"),
    varying = updated_laboratory_covariates
  ),
  format = "long",
  id = "id",
  rtime = "tstart"
)

enrollment_stack_raw$data$attained_age_z <- (
  enrollment_stack_raw$data$age_at_enrollment + enrollment_stack_raw$data$LM - age_center
) / age_scale

pbc_age <- transform(
  pbc_analysis,
  entry_age = age_at_enrollment,
  exit_age = age_at_enrollment + time,
  visit_age = age_at_enrollment + tstart
)

age_stack_raw <- stack_data(
  data = pbc_age,
  outcome = list(time = "exit_age", status = "status"),
  lms = age_landmarks,
  w = prediction_window,
  covs = list(
    fixed = shared_fixed_covariates,
    varying = updated_laboratory_covariates
  ),
  format = "long",
  id = "id",
  rtime = "visit_age",
  entry = "entry_age",
  entry_mode = "strict"
)

stopifnot(
  identical(age_stack_raw$entry_mode, "strict"),
  identical(age_stack_raw$entry_col, "LM"),
  all(age_stack_raw$data$entry_age <= age_stack_raw$data$LM),
  all(age_stack_raw$data$LM < age_stack_raw$data$exit_age),
  !anyDuplicated(age_stack_raw$data[c("id", "LM")]),
  !anyDuplicated(enrollment_stack_raw$data[c("id", "LM")])
)
```

```{r risk-set-table}
summarize_risk_sets <- function(stack, scale_label, time_column) {
  do.call(rbind, lapply(split(stack$data, stack$data$LM), function(landmark_data) {
    landmark <- unique(landmark_data$LM)
    stopifnot(length(landmark) == 1L)
    data.frame(
      scale = scale_label,
      landmark = landmark,
      subjects = length(unique(landmark_data$id)),
      deaths = sum(landmark_data$status == cause_of_interest),
      transplants = sum(landmark_data$status == 1L),
      horizon_controls = sum(
        landmark_data$status == 0L &
          landmark_data[[time_column]] >= landmark + prediction_window
      ),
      censorings_before_horizon = sum(
        landmark_data$status == 0L &
          landmark_data[[time_column]] < landmark + prediction_window
      ),
      row.names = NULL
    )
  }))
}

risk_sets <- rbind(
  summarize_risk_sets(enrollment_stack_raw, "Enrollment time", "time"),
  summarize_risk_sets(age_stack_raw, "Attained age", "exit_age")
)

enrollment_risk_counts <- risk_sets$subjects[risk_sets$scale == "Enrollment time"]
age_risk_counts <- risk_sets$subjects[risk_sets$scale == "Attained age"]

stopifnot(
  all(diff(enrollment_risk_counts) <= 0L),
  all(risk_sets$deaths <= risk_sets$subjects),
  all(risk_sets$transplants <= risk_sets$subjects)
)

knitr::kable(risk_sets, digits = 1)
```

The enrollment risk sets decreased monotonically. The attained-age risk sets
did not have to decrease: later landmarks lost people to events and censoring
but could also gain people who entered observation at older ages. This is why a
monotonicity check that is sensible after a shared origin is not an age-scale
eligibility check.

# Model fitting

I allowed the effects of the updated laboratory covariates to vary over a
centered quadratic landmark basis on both clocks. For enrollment time I also
used that basis as `func_lms`, producing the standard LM main effects alongside
the LM-by-laboratory interactions. I did not include LM main effects in the
primary attained-age model because its current prediction windows do not
overlap; the risk-set check below shows why those coefficients are not
identified by the Cox partial likelihood.

```{r add-landmark-interactions}
enrollment_stack <- add_interactions(
  enrollment_stack_raw,
  lm_covs = updated_laboratory_covariates,
  func_covars = enrollment_basis,
  func_lms = enrollment_basis
)
age_stack <- add_interactions(
  age_stack_raw,
  lm_covs = updated_laboratory_covariates,
  func_covars = age_basis,
  func_lms = list()
)

interaction_names <- c(
  "log_bili_z_LM1", "log_bili_z_LM2",
  "albumin_z_LM1", "albumin_z_LM2"
)
enrollment_lm_main_effects <- grep(
  "^LM[0-9]+$", names(enrollment_stack$data), value = TRUE
)

stopifnot(
  identical(enrollment_lm_main_effects, c("LM1", "LM2")),
  all(interaction_names %in% names(enrollment_stack$data)),
  all(interaction_names %in% names(age_stack$data)),
  !any(c("LM1", "LM2") %in% names(age_stack$data))
)

interaction_columns <- data.frame(
  component = c(
    updated_laboratory_covariates,
    rep("Landmark main effect", length(enrollment_lm_main_effects))
  ),
  linear_or_first_basis_column = c(
    paste0(updated_laboratory_covariates, "_LM1"),
    enrollment_lm_main_effects
  ),
  quadratic_or_second_basis_column = c(
    paste0(updated_laboratory_covariates, "_LM2"),
    rep("", length(enrollment_lm_main_effects))
  ),
  scale = c(
    rep("Both", length(updated_laboratory_covariates)),
    rep("Enrollment time only", length(enrollment_lm_main_effects))
  ),
  row.names = NULL
)

knitr::kable(interaction_columns)
```

I used the same cause-specific outcome specification on both scales. Repeated
landmark records were clustered by subject, so all inference below uses the
robust covariance returned by the cause-specific Cox fits.

```{r fit-models}
enrollment_formula <- as.formula(
  paste(
    "Hist(time, status, LM) ~",
    "attained_age_z + male + stage + trt +",
    "log_bili_z + log_bili_z_LM1 + log_bili_z_LM2 +",
    "albumin_z + albumin_z_LM1 + albumin_z_LM2 +",
    "LM1 + LM2 + cluster(id)"
  )
)

age_formula <- as.formula(
  paste(
    "Hist(exit_age, status, LM) ~",
    "male + stage + trt +",
    "log_bili_z + log_bili_z_LM1 + log_bili_z_LM2 +",
    "albumin_z + albumin_z_LM1 + albumin_z_LM2 + cluster(id)"
  )
)

capture_conditions <- function(expression) {
  warnings <- character()
  messages <- character()
  value <- withCallingHandlers(
    tryCatch(eval.parent(substitute(expression)), error = function(error) error),
    warning = function(warning) {
      warnings <<- c(warnings, conditionMessage(warning))
      invokeRestart("muffleWarning")
    },
    message = function(message) {
      messages <<- c(messages, conditionMessage(message))
      invokeRestart("muffleMessage")
    }
  )
  list(value = value, warnings = unique(warnings), messages = unique(messages))
}

enrollment_fit_record <- capture_conditions(dynamic_lm(
  enrollment_stack,
  enrollment_formula,
  type = "CSC",
  method = "breslow",
  x = TRUE
))
age_fit_record <- capture_conditions(dynamic_lm(
  age_stack,
  age_formula,
  type = "CSC",
  method = "breslow",
  x = TRUE
))

fit_condition_table <- data.frame(
  scale = c("Enrollment time", "Attained age"),
  warning_count = c(
    length(enrollment_fit_record$warnings),
    length(age_fit_record$warnings)
  ),
  error = c(
    if (inherits(enrollment_fit_record$value, "error")) conditionMessage(enrollment_fit_record$value) else "",
    if (inherits(age_fit_record$value, "error")) conditionMessage(age_fit_record$value) else ""
  ),
  row.names = NULL
)

knitr::kable(fit_condition_table)

stopifnot(
  !inherits(enrollment_fit_record$value, "error"),
  !inherits(age_fit_record$value, "error"),
  length(enrollment_fit_record$warnings) == 0L,
  length(age_fit_record$warnings) == 0L
)

enrollment_fit <- enrollment_fit_record$value
age_fit <- age_fit_record$value
```

## Why the primary age model has no LM main effect

The age windows were disjoint, whereas the enrollment windows overlapped. I
checked both the window endpoints and the actual partial-likelihood risk sets.
At each attained-age event time, every at-risk stacked row came from the same
landmark. The LM value was therefore constant within that risk set and could
not change any partial-likelihood probability. Enrollment-time event risk sets
contained rows from several landmarks, so its centered LM terms were
estimable.

```{r landmark-identifiability-check}
summarize_adjacent_windows <- function(landmarks, scale_label) {
  data.frame(
    scale = scale_label,
    earlier_landmark = head(landmarks, -1L),
    earlier_window_end = head(landmarks, -1L) + prediction_window,
    later_landmark = tail(landmarks, -1L),
    overlaps = tail(landmarks, -1L) <= head(landmarks, -1L) + prediction_window,
    row.names = NULL
  )
}

window_overlap <- rbind(
  summarize_adjacent_windows(enrollment_landmarks, "Enrollment time"),
  summarize_adjacent_windows(age_landmarks, "Attained age")
)

risk_set_landmark_counts <- function(stack, time_column, scale_label) {
  do.call(rbind, lapply(sort(unique(stack$data$status[stack$data$status > 0L])), function(cause) {
    event_times <- sort(unique(
      stack$data[[time_column]][stack$data$status == cause]
    ))
    counts <- vapply(event_times, function(event_time) {
      at_risk <- stack$data$LM < event_time & stack$data[[time_column]] >= event_time
      length(unique(stack$data$LM[at_risk]))
    }, integer(1L))
    stopifnot(length(counts) == length(event_times), all(counts >= 1L))
    data.frame(
      scale = scale_label,
      cause = cause,
      event_times = length(event_times),
      minimum_LM_values_in_risk_set = min(counts),
      maximum_LM_values_in_risk_set = max(counts),
      event_times_with_multiple_LM_values = sum(counts > 1L),
      row.names = NULL
    )
  }))
}

risk_set_lm_check <- rbind(
  risk_set_landmark_counts(enrollment_stack_raw, "time", "Enrollment time"),
  risk_set_landmark_counts(age_stack_raw, "exit_age", "Attained age")
)

stopifnot(
  any(window_overlap$overlaps[window_overlap$scale == "Enrollment time"]),
  !any(window_overlap$overlaps[window_overlap$scale == "Attained age"]),
  all(risk_set_lm_check$maximum_LM_values_in_risk_set[
    risk_set_lm_check$scale == "Attained age"
  ] == 1L),
  any(risk_set_lm_check$maximum_LM_values_in_risk_set[
    risk_set_lm_check$scale == "Enrollment time"
  ] > 1L)
)

knitr::kable(window_overlap, digits = 1)
knitr::kable(risk_set_lm_check)
```

I also fitted the deliberately over-specified age model from a fresh stack.
This was a diagnostic fit only; none of its coefficients or predictions were
used later.

```{r age-lm-main-effect-demonstration}
age_stack_lm_demo <- add_interactions(
  age_stack_raw,
  lm_covs = updated_laboratory_covariates,
  func_covars = age_basis,
  func_lms = age_basis
)
age_lm_demo_formula <- update(age_formula, . ~ . + LM1 + LM2)
age_lm_demo_record <- capture_conditions(dynamic_lm(
  age_stack_lm_demo,
  age_lm_demo_formula,
  type = "CSC",
  method = "breslow",
  x = TRUE
))

age_lm_demo_condition <- data.frame(
  warning_count = length(age_lm_demo_record$warnings),
  warnings = paste(age_lm_demo_record$warnings, collapse = " | "),
  error = if (inherits(age_lm_demo_record$value, "error")) {
    conditionMessage(age_lm_demo_record$value)
  } else "",
  row.names = NULL
)
knitr::kable(age_lm_demo_condition)

if (inherits(age_lm_demo_record$value, "error")) {
  age_lm_demo_coefficients <- data.frame(
    cause = NA_integer_, term = c("LM1", "LM2"), estimate = NA_real_,
    robust_standard_error = NA_real_, result = "fit failed"
  )
} else {
  age_lm_demo_coefficients <- do.call(rbind, lapply(
    age_lm_demo_record$value$model$causes,
    function(cause) {
      cause_fit <- age_lm_demo_record$value$model$models[[
        match(cause, age_lm_demo_record$value$model$causes)
      ]]
      terms <- c("LM1", "LM2")
      estimates <- coef(cause_fit)[terms]
      standard_errors <- sqrt(diag(vcov(cause_fit)))[terms]
      data.frame(
        cause = cause,
        term = terms,
        estimate = unname(estimates),
        robust_standard_error = unname(standard_errors),
        result = ifelse(
          is.finite(estimates) & is.finite(standard_errors) & standard_errors > 0,
          "estimable", "non-estimable"
        ),
        row.names = NULL
      )
    }
  ))
  stopifnot(all(age_lm_demo_coefficients$result == "non-estimable"))
}

knitr::kable(age_lm_demo_coefficients, digits = 4)
```

# Full supermodel coefficient comparison

The interaction bases sit on different clocks, so their coefficients are not
contrasts. I displayed every death-model term to make the two fitted
specifications transparent. The dashes identify terms that are structurally
absent from the primary attained-age model.

```{r full-coefficient-table}
extract_cause_fit <- function(supermodel, cause) {
  cause_index <- match(cause, supermodel$model$causes)
  stopifnot(!is.na(cause_index))
  supermodel$model$models[[cause_index]]
}

extract_robust_inference <- function(supermodel, cause = cause_of_interest) {
  cause_fit <- extract_cause_fit(supermodel, cause)
  estimates <- coef(cause_fit)
  robust_standard_errors <- sqrt(diag(vcov(cause_fit)))
  summary_coefficients <- summary(cause_fit)$coefficients
  stopifnot(
    "robust se" %in% colnames(summary_coefficients),
    identical(names(estimates), rownames(summary_coefficients)),
    max(abs(
      robust_standard_errors - summary_coefficients[, "robust se"]
    ), na.rm = TRUE) < 1e-12
  )
  critical_value <- qnorm(1 - (1 - confidence_level) / 2)
  data.frame(
    term = names(estimates),
    estimate = unname(estimates),
    robust_standard_error = unname(robust_standard_errors),
    lower = unname(estimates - critical_value * robust_standard_errors),
    upper = unname(estimates + critical_value * robust_standard_errors),
    hazard_ratio = exp(unname(estimates)),
    hazard_ratio_lower = exp(unname(estimates - critical_value * robust_standard_errors)),
    hazard_ratio_upper = exp(unname(estimates + critical_value * robust_standard_errors)),
    p_value = 2 * pnorm(-abs(unname(estimates / robust_standard_errors))),
    row.names = NULL
  )
}

enrollment_coefficients <- extract_robust_inference(enrollment_fit)
age_coefficients <- extract_robust_inference(age_fit)

full_term_order <- c(
  "attained_age_z", "male", "stage", "trt",
  "log_bili_z", "log_bili_z_LM1", "log_bili_z_LM2",
  "albumin_z", "albumin_z_LM1", "albumin_z_LM2", "LM1", "LM2"
)
term_labels <- c(
  attained_age_z = "Attained age (per SD)",
  male = "Male", stage = "Disease stage", trt = "Treatment",
  log_bili_z = "Log bilirubin (per SD)",
  log_bili_z_LM1 = "Log bilirubin x LM basis 1",
  log_bili_z_LM2 = "Log bilirubin x LM basis 2",
  albumin_z = "Albumin (per SD)",
  albumin_z_LM1 = "Albumin x LM basis 1",
  albumin_z_LM2 = "Albumin x LM basis 2",
  LM1 = "LM basis 1", LM2 = "LM basis 2"
)

format_hr_ci <- function(inference, term) {
  index <- match(term, inference$term)
  if (is.na(index) || !is.finite(inference$hazard_ratio[index])) return("—")
  sprintf(
    "%.3f (%.3f, %.3f)",
    inference$hazard_ratio[index],
    inference$hazard_ratio_lower[index],
    inference$hazard_ratio_upper[index]
  )
}
format_p_value <- function(inference, term) {
  index <- match(term, inference$term)
  if (is.na(index) || !is.finite(inference$p_value[index])) return("—")
  format.pval(inference$p_value[index], digits = 3, eps = 0.001)
}

coefficient_comparison <- data.frame(
  term = unname(term_labels[full_term_order]),
  enrollment_HR_95_CI = vapply(
    full_term_order, function(term) format_hr_ci(enrollment_coefficients, term),
    character(1L)
  ),
  enrollment_robust_Wald_p = vapply(
    full_term_order, function(term) format_p_value(enrollment_coefficients, term),
    character(1L)
  ),
  attained_age_HR_95_CI = vapply(
    full_term_order, function(term) format_hr_ci(age_coefficients, term),
    character(1L)
  ),
  attained_age_robust_Wald_p = vapply(
    full_term_order, function(term) format_p_value(age_coefficients, term),
    character(1L)
  ),
  row.names = NULL
)

stopifnot(
  all(c("attained_age_z", "LM1", "LM2") %in% enrollment_coefficients$term),
  !any(c("attained_age_z", "LM1", "LM2") %in% age_coefficients$term)
)

knitr::kable(coefficient_comparison, align = c("l", "c", "c", "c", "c"))
```

The landmark-specific laboratory table combines each main effect with its two
interaction coefficients. Its confidence intervals use the delta method and
the same robust covariance matrix as the full coefficient table.

```{r landmark-specific-supermodel-hazard-ratios}
dynamic_log_hazard_ratio <- function(supermodel, marker, landmarks, basis) {
  cause_fit <- extract_cause_fit(supermodel, cause_of_interest)
  marker_terms <- c(marker, paste0(marker, "_LM1"), paste0(marker, "_LM2"))
  stopifnot(all(marker_terms %in% names(coef(cause_fit))))
  design <- cbind(1, basis[[1L]](landmarks), basis[[2L]](landmarks))
  coefficients <- coef(cause_fit)[marker_terms]
  covariance <- vcov(cause_fit)[marker_terms, marker_terms, drop = FALSE]
  estimate <- as.numeric(design %*% coefficients)
  standard_error <- sqrt(rowSums((design %*% covariance) * design))
  critical_value <- qnorm(1 - (1 - confidence_level) / 2)
  data.frame(
    landmark = landmarks,
    estimate = estimate,
    standard_error = standard_error,
    lower = estimate - critical_value * standard_error,
    upper = estimate + critical_value * standard_error,
    row.names = NULL
  )
}

landmark_specific_supermodel_hr <- do.call(rbind, list(
  transform(
    dynamic_log_hazard_ratio(enrollment_fit, "log_bili_z", enrollment_landmarks, enrollment_basis),
    scale = "Enrollment time", covariate = "Log bilirubin (per SD)"
  ),
  transform(
    dynamic_log_hazard_ratio(enrollment_fit, "albumin_z", enrollment_landmarks, enrollment_basis),
    scale = "Enrollment time", covariate = "Albumin (per SD)"
  ),
  transform(
    dynamic_log_hazard_ratio(age_fit, "log_bili_z", age_landmarks, age_basis),
    scale = "Attained age", covariate = "Log bilirubin (per SD)"
  ),
  transform(
    dynamic_log_hazard_ratio(age_fit, "albumin_z", age_landmarks, age_basis),
    scale = "Attained age", covariate = "Albumin (per SD)"
  )
))
landmark_specific_supermodel_hr$HR_95_CI <- sprintf(
  "%.3f (%.3f, %.3f)",
  exp(landmark_specific_supermodel_hr$estimate),
  exp(landmark_specific_supermodel_hr$lower),
  exp(landmark_specific_supermodel_hr$upper)
)

knitr::kable(
  landmark_specific_supermodel_hr[c("scale", "landmark", "covariate", "HR_95_CI")],
  digits = 1
)
```

# Disjoint landmark models

I fitted a separate competing-risk model at each landmark using the same risk
set, endpoint, window, and substantive covariates as its supermodel. These fits
had no LM terms or interactions. The enrollment disjoint models retained
updated attained age. I did not alter a model to suppress instability.

```{r disjoint-model-fits}
subset_lmdata <- function(lmdata, landmark) {
  out <- lmdata
  out$data <- lmdata$data[lmdata$data$LM == landmark, , drop = FALSE]
  out$end_time <- landmark
  stopifnot(nrow(out$data) > 0L, all(out$data$LM == landmark))
  out
}

enrollment_disjoint_formula <- as.formula(
  paste(
    "Hist(time, status, LM) ~",
    "attained_age_z + male + stage + trt + log_bili_z + albumin_z + cluster(id)"
  )
)
age_disjoint_formula <- as.formula(
  paste(
    "Hist(exit_age, status, LM) ~",
    "male + stage + trt + log_bili_z + albumin_z + cluster(id)"
  )
)

fit_disjoint_models <- function(lmdata, landmarks, formula, time_column, scale_label) {
  records <- setNames(vector("list", length(landmarks)), as.character(landmarks))
  model_diagnostic_rows <- list()
  diagnostic_rows <- list()
  rhs <- paste(deparse(formula[[3L]]), collapse = "")

  for (landmark in landmarks) {
    one_landmark <- subset_lmdata(lmdata, landmark)
    model_record <- capture_conditions(dynamic_lm(
      one_landmark, formula, type = "CSC", method = "breslow", x = TRUE
    ))
    records[[as.character(landmark)]] <- model_record
    model_diagnostic_rows[[length(model_diagnostic_rows) + 1L]] <- data.frame(
      scale = scale_label,
      landmark = landmark,
      warning_count = length(model_record$warnings),
      warnings = paste(model_record$warnings, collapse = " | "),
      error = if (inherits(model_record$value, "error")) {
        conditionMessage(model_record$value)
      } else "",
      row.names = NULL
    )

    for (cause in c(transplant_code, cause_of_interest)) {
      direct_formula <- as.formula(paste0(
        "Surv(LM, ", time_column, ", status == ", cause, ") ~ ", rhs
      ))
      direct_record <- capture_conditions(survival::coxph(
        direct_formula, data = one_landmark$data, method = "breslow", x = TRUE
      ))
      outer_error <- if (inherits(model_record$value, "error")) {
        conditionMessage(model_record$value)
      } else ""
      direct_error <- if (inherits(direct_record$value, "error")) {
        conditionMessage(direct_record$value)
      } else ""
      nonestimable <- character()

      if (!inherits(model_record$value, "error")) {
        cause_fit <- extract_cause_fit(model_record$value, cause)
        nonestimable <- names(coef(cause_fit))[!is.finite(coef(cause_fit))]
        if (!inherits(direct_record$value, "error")) {
          common_terms <- intersect(names(coef(cause_fit)), names(coef(direct_record$value)))
          finite_terms <- common_terms[
            is.finite(coef(cause_fit)[common_terms]) &
              is.finite(coef(direct_record$value)[common_terms])
          ]
          if (length(finite_terms)) {
            stopifnot(max(abs(
              coef(cause_fit)[finite_terms] - coef(direct_record$value)[finite_terms]
            )) < 1e-8)
          }
        }
      }

      diagnostic_rows[[length(diagnostic_rows) + 1L]] <- data.frame(
        scale = scale_label,
        landmark = landmark,
        cause = cause,
        warning_count = length(direct_record$warnings),
        warnings = paste(direct_record$warnings, collapse = " | "),
        error = paste(c(outer_error, direct_error)[nzchar(c(outer_error, direct_error))], collapse = " | "),
        nonestimable_terms = paste(nonestimable, collapse = ", "),
        row.names = NULL
      )
    }
  }

  list(
    records = records,
    model_diagnostics = do.call(rbind, model_diagnostic_rows),
    diagnostics = do.call(rbind, diagnostic_rows)
  )
}

enrollment_disjoint <- fit_disjoint_models(
  enrollment_stack_raw, enrollment_landmarks, enrollment_disjoint_formula,
  "time", "Enrollment time"
)
age_disjoint <- fit_disjoint_models(
  age_stack_raw, age_landmarks, age_disjoint_formula,
  "exit_age", "Attained age"
)
disjoint_diagnostics <- rbind(
  enrollment_disjoint$diagnostics,
  age_disjoint$diagnostics
)
disjoint_model_diagnostics <- rbind(
  enrollment_disjoint$model_diagnostics,
  age_disjoint$model_diagnostics
)

knitr::kable(
  disjoint_model_diagnostics, digits = 1,
  caption = "Warnings and errors returned by each competing-risk model"
)
knitr::kable(
  disjoint_diagnostics, digits = 1,
  caption = "Cause-specific warnings, errors, and non-estimable coefficients"
)
```

## Supermodel and disjoint estimates

The long comparison table keeps sample size and both event types beside every
estimate. `fit_flag` reports any death-model warning, failure, or
non-estimability rather than silently dropping an affected estimate.

```{r disjoint-coefficient-comparison}
get_disjoint_record <- function(disjoint_result, landmark) {
  index <- which(abs(as.numeric(names(disjoint_result$records)) - landmark) < 1e-8)
  stopifnot(length(index) == 1L)
  disjoint_result$records[[index]]
}

supermodel_effect <- function(supermodel, covariate, landmark, basis) {
  if (covariate %in% updated_laboratory_covariates) {
    dynamic_log_hazard_ratio(supermodel, covariate, landmark, basis)
  } else {
    inference <- extract_robust_inference(supermodel)
    row <- inference[inference$term == covariate, ]
    stopifnot(nrow(row) == 1L)
    data.frame(
      landmark = landmark, estimate = row$estimate,
      standard_error = row$robust_standard_error,
      lower = row$lower, upper = row$upper
    )
  }
}

disjoint_effect <- function(disjoint_result, landmark, covariate) {
  record <- get_disjoint_record(disjoint_result, landmark)
  if (inherits(record$value, "error")) {
    return(data.frame(
      estimate = NA_real_, standard_error = NA_real_,
      lower = NA_real_, upper = NA_real_
    ))
  }
  inference <- extract_robust_inference(record$value)
  row <- inference[inference$term == covariate, ]
  if (!nrow(row)) {
    return(data.frame(
      estimate = NA_real_, standard_error = NA_real_,
      lower = NA_real_, upper = NA_real_
    ))
  }
  out <- row[c("estimate", "robust_standard_error", "lower", "upper")]
  names(out)[names(out) == "robust_standard_error"] <- "standard_error"
  out
}

comparison_specifications <- list(
  list(
    scale = "Enrollment time", landmarks = enrollment_landmarks,
    supermodel = enrollment_fit, disjoint = enrollment_disjoint,
    basis = enrollment_basis,
    covariates = c("attained_age_z", shared_model_covariates)
  ),
  list(
    scale = "Attained age", landmarks = age_landmarks,
    supermodel = age_fit, disjoint = age_disjoint,
    basis = age_basis, covariates = shared_model_covariates
  )
)

comparison_rows <- list()
for (specification in comparison_specifications) {
  for (landmark in specification$landmarks) {
    counts <- risk_sets[
      risk_sets$scale == specification$scale & risk_sets$landmark == landmark,
    ]
    stopifnot(nrow(counts) == 1L)
    diagnostics <- specification$disjoint$diagnostics[
      specification$disjoint$diagnostics$landmark == landmark &
        specification$disjoint$diagnostics$cause == cause_of_interest,
    ]
    stopifnot(nrow(diagnostics) == 1L)

    for (covariate in specification$covariates) {
      supermodel_row <- supermodel_effect(
        specification$supermodel, covariate, landmark, specification$basis
      )
      disjoint_row <- disjoint_effect(
        specification$disjoint, landmark, covariate
      )
      flags <- c(
        if (nzchar(diagnostics$error)) "fit failure" else character(),
        if (diagnostics$warning_count > 0L) "warning" else character(),
        if (!is.finite(disjoint_row$estimate) ||
            !is.finite(disjoint_row$standard_error)) "non-estimable" else character()
      )
      comparison_rows[[length(comparison_rows) + 1L]] <- data.frame(
        scale = specification$scale,
        landmark = landmark,
        subjects = counts$subjects,
        deaths = counts$deaths,
        transplants = counts$transplants,
        covariate = covariate,
        supermodel_estimate = supermodel_row$estimate,
        supermodel_standard_error = supermodel_row$standard_error,
        supermodel_lower = supermodel_row$lower,
        supermodel_upper = supermodel_row$upper,
        disjoint_estimate = disjoint_row$estimate,
        disjoint_standard_error = disjoint_row$standard_error,
        disjoint_lower = disjoint_row$lower,
        disjoint_upper = disjoint_row$upper,
        fit_flag = if (length(flags)) paste(unique(flags), collapse = "; ") else "ok",
        row.names = NULL
      )
    }
  }
}
disjoint_comparison <- do.call(rbind, comparison_rows)
disjoint_comparison$supermodel_HR_95_CI <- sprintf(
  "%.3f (%.3f, %.3f)",
  exp(disjoint_comparison$supermodel_estimate),
  exp(disjoint_comparison$supermodel_lower),
  exp(disjoint_comparison$supermodel_upper)
)
disjoint_comparison$disjoint_HR_95_CI <- ifelse(
  is.finite(disjoint_comparison$disjoint_estimate),
  sprintf(
    "%.3f (%.3f, %.3f)",
    exp(disjoint_comparison$disjoint_estimate),
    exp(disjoint_comparison$disjoint_lower),
    exp(disjoint_comparison$disjoint_upper)
  ),
  "—"
)
disjoint_comparison$CI_width_ratio_log_scale <- (
  disjoint_comparison$disjoint_upper - disjoint_comparison$disjoint_lower
) / (
  disjoint_comparison$supermodel_upper - disjoint_comparison$supermodel_lower
)

covariate_labels <- c(term_labels, log_bili_z = "Log bilirubin (per SD)",
                      albumin_z = "Albumin (per SD)")
comparison_display <- transform(
  disjoint_comparison,
  covariate = unname(covariate_labels[covariate])
)

knitr::kable(
  comparison_display[c(
    "scale", "landmark", "subjects", "deaths", "transplants", "covariate",
    "supermodel_HR_95_CI", "disjoint_HR_95_CI",
    "CI_width_ratio_log_scale", "fit_flag"
  )],
  digits = 2
)
```

I measured efficiency on the log-HR scale, where confidence-interval width is
proportional to the robust standard error. The spread table describes how far
the separate landmark estimates moved across each scale.

```{r disjoint-efficiency-summary}
efficiency_summary <- do.call(rbind, lapply(
  split(disjoint_comparison, list(
    disjoint_comparison$scale, disjoint_comparison$covariate
  ), drop = TRUE),
  function(rows) {
    finite <- is.finite(rows$disjoint_estimate)
    data.frame(
      scale = rows$scale[1L],
      covariate = unname(covariate_labels[rows$covariate[1L]]),
      estimable_landmarks = sum(finite),
      disjoint_log_HR_range = if (any(finite)) {
        diff(range(rows$disjoint_estimate[finite]))
      } else NA_real_,
      disjoint_log_HR_SD = if (sum(finite) > 1L) {
        sd(rows$disjoint_estimate[finite])
      } else NA_real_,
      median_CI_width_ratio = median(
        rows$CI_width_ratio_log_scale[finite], na.rm = TRUE
      ),
      maximum_CI_width_ratio = max(
        rows$CI_width_ratio_log_scale[finite], na.rm = TRUE
      ),
      row.names = NULL
    )
  }
))
rownames(efficiency_summary) <- NULL
knitr::kable(efficiency_summary, digits = 3)
```

## Supermodel curves and disjoint estimates

The points and error bars are the separate landmark fits. The line and grey
band are the supermodel estimate and robust confidence interval. I used the same
vertical scale within each panel, so instability is visible rather than
rescaled away.

```{r dynamic-log-hazard-ratios, fig.width=10, fig.height=8}
plot_effect_comparison <- function(curve, disjoint_rows, title, x_label,
                                   y_label = "Log hazard ratio") {
  plot_range <- range(
    curve$lower, curve$upper,
    disjoint_rows$disjoint_lower, disjoint_rows$disjoint_upper,
    finite = TRUE
  )
  plot(
    curve$landmark, curve$estimate, type = "n",
    xlab = x_label, ylab = y_label, main = title, ylim = plot_range
  )
  polygon(
    c(curve$landmark, rev(curve$landmark)),
    c(curve$lower, rev(curve$upper)),
    border = NA, col = "grey85"
  )
  abline(h = 0, lty = 2, col = "grey45")
  lines(curve$landmark, curve$estimate, lwd = 2)
  finite <- is.finite(disjoint_rows$disjoint_estimate) &
    is.finite(disjoint_rows$disjoint_lower) &
    is.finite(disjoint_rows$disjoint_upper)
  arrows(
    disjoint_rows$landmark[finite], disjoint_rows$disjoint_lower[finite],
    disjoint_rows$landmark[finite], disjoint_rows$disjoint_upper[finite],
    angle = 90, code = 3, length = 0.04
  )
  points(
    disjoint_rows$landmark[finite],
    disjoint_rows$disjoint_estimate[finite], pch = 19
  )
}

enrollment_grid <- seq(
  min(enrollment_landmarks), max(enrollment_landmarks), length.out = 200L
)
age_grid <- seq(min(age_landmarks), max(age_landmarks), length.out = 200L)
dynamic_curves <- list(
  enrollment_bili = dynamic_log_hazard_ratio(
    enrollment_fit, "log_bili_z", enrollment_grid, enrollment_basis
  ),
  age_bili = dynamic_log_hazard_ratio(
    age_fit, "log_bili_z", age_grid, age_basis
  ),
  enrollment_albumin = dynamic_log_hazard_ratio(
    enrollment_fit, "albumin_z", enrollment_grid, enrollment_basis
  ),
  age_albumin = dynamic_log_hazard_ratio(
    age_fit, "albumin_z", age_grid, age_basis
  )
)

old_parameters <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))
plot_effect_comparison(
  dynamic_curves$enrollment_bili,
  disjoint_comparison[
    disjoint_comparison$scale == "Enrollment time" &
      disjoint_comparison$covariate == "log_bili_z", ],
  "Bilirubin: enrollment time", "Years after enrollment"
)
plot_effect_comparison(
  dynamic_curves$age_bili,
  disjoint_comparison[
    disjoint_comparison$scale == "Attained age" &
      disjoint_comparison$covariate == "log_bili_z", ],
  "Bilirubin: attained age", "Attained age"
)
plot_effect_comparison(
  dynamic_curves$enrollment_albumin,
  disjoint_comparison[
    disjoint_comparison$scale == "Enrollment time" &
      disjoint_comparison$covariate == "albumin_z", ],
  "Albumin: enrollment time", "Years after enrollment"
)
plot_effect_comparison(
  dynamic_curves$age_albumin,
  disjoint_comparison[
    disjoint_comparison$scale == "Attained age" &
      disjoint_comparison$covariate == "albumin_z", ],
  "Albumin: attained age", "Attained age"
)
par(old_parameters)
```

The fixed-effect panels make the pooling distinction explicit. The
enrollment and age supermodels each use one effect for these covariates, while
the disjoint estimates spend a separate coefficient at every landmark.

```{r fixed-effect-supermodel-disjoint, fig.width=11, fig.height=7}
constant_effect_curve <- function(supermodel, covariate, grid) {
  row <- extract_robust_inference(supermodel)
  row <- row[row$term == covariate, ]
  stopifnot(nrow(row) == 1L)
  data.frame(
    landmark = grid,
    estimate = row$estimate,
    lower = row$lower,
    upper = row$upper
  )
}

old_parameters <- par(mfrow = c(2, 3), mar = c(4, 4, 3, 1))
for (scale_label in c("Enrollment time", "Attained age")) {
  supermodel <- if (scale_label == "Enrollment time") enrollment_fit else age_fit
  grid <- if (scale_label == "Enrollment time") enrollment_grid else age_grid
  x_label <- if (scale_label == "Enrollment time") "Years after enrollment" else "Attained age"
  for (covariate in shared_fixed_covariates) {
    plot_effect_comparison(
      constant_effect_curve(supermodel, covariate, grid),
      disjoint_comparison[
        disjoint_comparison$scale == scale_label &
          disjoint_comparison$covariate == covariate, ],
      paste(unname(term_labels[covariate]), scale_label, sep = ": "),
      x_label
    )
  }
}
par(old_parameters)
```

# Predictive performance

I first computed apparent performance on the exact rows used for model fitting.
I then estimated Harrell's bootstrap optimism by resampling subjects, rebuilding
every stack, refitting both supermodels and every disjoint model, and evaluating
each bootstrap fit in both its resample and the original cohort. Duplicated
subjects received new identifiers, so clustering never treated two bootstrap
copies as the same person.

## Apparent landmark-specific performance

```{r predictions-and-scores}
prediction_key <- function(data) {
  paste(data$id, sprintf("%.12g", data$LM), sep = "::")
}

align_prediction <- function(prediction, target_stack) {
  target_key <- prediction_key(target_stack$data)
  source_key <- prediction_key(prediction$data)
  stopifnot(!anyDuplicated(target_key), !anyDuplicated(source_key))
  index <- match(target_key, source_key)
  stopifnot(!anyNA(index))
  prediction$data <- prediction$data[index, , drop = FALSE]
  prediction$preds <- prediction$preds[index, , drop = FALSE]
  stopifnot(identical(prediction_key(prediction$data), target_key))
  prediction
}

subset_prediction <- function(prediction, landmark) {
  index <- prediction$data$LM == landmark
  out <- prediction
  out$data <- prediction$data[index, , drop = FALSE]
  out$preds <- prediction$preds[index, , drop = FALSE]
  stopifnot(nrow(out$data) == nrow(out$preds), nrow(out$data) > 0L)
  out
}

predict_disjoint_by_landmark <- function(disjoint_result, target_stack,
                                         landmarks) {
  setNames(lapply(landmarks, function(landmark) {
    record <- get_disjoint_record(disjoint_result, landmark)
    if (inherits(record$value, "error")) return(record$value)
    target <- subset_lmdata(target_stack, landmark)
    prediction_record <- capture_conditions(predict(
      record$value, newdata = target, cause = cause_of_interest
    ))
    prediction_record$value
  }), as.character(landmarks))
}

enrollment_prediction_record <- capture_conditions(predict(
  enrollment_fit, newdata = enrollment_stack, cause = cause_of_interest
))
age_prediction_record <- capture_conditions(predict(
  age_fit, newdata = age_stack, cause = cause_of_interest
))
enrollment_disjoint_prediction_records <- predict_disjoint_by_landmark(
  enrollment_disjoint, enrollment_stack_raw, enrollment_landmarks
)
age_disjoint_prediction_records <- predict_disjoint_by_landmark(
  age_disjoint, age_stack_raw, age_landmarks
)

prediction_records <- list(
  enrollment_prediction_record,
  age_prediction_record
)
stopifnot(
  !any(vapply(prediction_records, function(record) {
    inherits(record$value, "error")
  }, logical(1L))),
  all(vapply(prediction_records, function(record) {
    length(record$warnings) == 0L
  }, logical(1L)))
)

enrollment_predictions <- align_prediction(
  enrollment_prediction_record$value, enrollment_stack
)
age_predictions <- align_prediction(age_prediction_record$value, age_stack)

score_predictions_by_landmark <- function(super_prediction,
                                          disjoint_predictions,
                                          landmarks, scale_label) {
  rows <- list()
  failures <- list()
  for (landmark in landmarks) {
    super_one <- subset_prediction(super_prediction, landmark)
    model_list <- list(Supermodel = super_one)
    disjoint_one <- disjoint_predictions[[as.character(landmark)]]
    if (inherits(disjoint_one, "error")) {
      failures[[length(failures) + 1L]] <- data.frame(
        scale = scale_label, landmark = landmark, method = "Disjoint",
        stage = "prediction", error = conditionMessage(disjoint_one)
      )
    } else {
      stopifnot(identical(
        prediction_key(super_one$data), prediction_key(disjoint_one$data)
      ))
      disjoint_one$data <- super_one$data
      model_list$Disjoint <- disjoint_one
    }
    score_record <- capture_conditions(score(
      model_list, times = landmark, metrics = c("auc", "brier"),
      id_col = "id", cause = cause_of_interest,
      summary = FALSE, contrasts = FALSE
    ))
    if (inherits(score_record$value, "error")) {
      failures[[length(failures) + 1L]] <- data.frame(
        scale = scale_label, landmark = landmark, method = "All available",
        stage = "score", error = conditionMessage(score_record$value)
      )
      next
    }
    auc <- as.data.frame(score_record$value$AUC$score)
    brier <- as.data.frame(score_record$value$Brier$score)
    auc$model <- as.character(auc$model)
    brier$model <- as.character(brier$model)
    available <- names(model_list)
    auc <- auc[auc$model %in% available, c("model", "AUC")]
    brier <- brier[brier$model %in% available, c("model", "Brier")]
    metric <- merge(auc, brier, by = "model", sort = FALSE)
    rows[[length(rows) + 1L]] <- transform(
      metric, scale = scale_label, landmark = landmark
    )[c("scale", "landmark", "model", "AUC", "Brier")]
  }
  list(
    score = if (length(rows)) do.call(rbind, rows) else NULL,
    failures = if (length(failures)) do.call(rbind, failures) else data.frame()
  )
}

enrollment_apparent <- score_predictions_by_landmark(
  enrollment_predictions, enrollment_disjoint_prediction_records,
  enrollment_landmarks, "Enrollment time"
)
age_apparent <- score_predictions_by_landmark(
  age_predictions, age_disjoint_prediction_records,
  age_landmarks, "Attained age"
)
ordinary_apparent_scores <- rbind(
  enrollment_apparent$score, age_apparent$score
)
apparent_prediction_failures <- rbind(
  enrollment_apparent$failures, age_apparent$failures
)

age_landmark_score_record <- capture_conditions(score(
  list("Age scale" = age_predictions),
  times = age_landmarks, metrics = c("auc", "brier"),
  id_col = "id", cause = cause_of_interest,
  summary = FALSE, contrasts = FALSE
))
stopifnot(!inherits(age_landmark_score_record$value, "error"))
```

Before selecting a common scoring path, I checked `score_left_truncated()`
against `score()` for both methods on both scales.

```{r scoring-function-agreement}
score_lt_by_landmark <- function(super_prediction, disjoint_predictions,
                                 landmarks, scale_label, time_column) {
  rows <- list()
  for (landmark in landmarks) {
    predictions <- list(
      Supermodel = subset_prediction(super_prediction, landmark),
      Disjoint = disjoint_predictions[[as.character(landmark)]]
    )
    predictions <- predictions[!vapply(predictions, inherits, logical(1L), what = "error")]
    for (method in names(predictions)) {
      prediction <- predictions[[method]]
      record <- capture_conditions(score_left_truncated(
        risk = setNames(data.frame(prediction$preds$risk), method),
        data = prediction$data,
        time_col = time_column, status_col = "status", entry_col = "LM",
        landmark_col = "LM", id_col = "id", cause = cause_of_interest,
        w = prediction_window
      ))
      if (!inherits(record$value, "error")) {
        row <- record$value$score
        row$scale <- scale_label
        rows[[length(rows) + 1L]] <- row[c(
          "scale", "landmark", "model", "AUC", "Brier"
        )]
      }
    }
  }
  do.call(rbind, rows)
}

left_truncated_apparent_scores <- rbind(
  score_lt_by_landmark(
    enrollment_predictions, enrollment_disjoint_prediction_records,
    enrollment_landmarks, "Enrollment time", "time"
  ),
  score_lt_by_landmark(
    age_predictions, age_disjoint_prediction_records,
    age_landmarks, "Attained age", "exit_age"
  )
)

score_comparison <- merge(
  ordinary_apparent_scores,
  left_truncated_apparent_scores[c("scale", "landmark", "model", "AUC", "Brier")],
  by = c("scale", "landmark", "model"),
  suffixes = c("_score", "_score_left_truncated"),
  sort = FALSE
)
score_agreement <- do.call(rbind, lapply(
  split(score_comparison, list(score_comparison$scale, score_comparison$model), drop = TRUE),
  function(rows) {
    data.frame(
      scale = rows$scale[1L], model = rows$model[1L],
      maximum_AUC_difference = max(abs(
        rows$AUC_score - rows$AUC_score_left_truncated
      )),
      maximum_Brier_difference = max(abs(
        rows$Brier_score - rows$Brier_score_left_truncated
      )),
      row.names = NULL
    )
  }
))
rownames(score_agreement) <- NULL

agreement_tolerance <- 1e-12
stopifnot(all(
  score_agreement$maximum_AUC_difference[
    score_agreement$scale == "Attained age"
  ] < agreement_tolerance &
    score_agreement$maximum_Brier_difference[
      score_agreement$scale == "Attained age"
    ] < agreement_tolerance
))
all_paths_agree <- all(
  score_agreement$maximum_AUC_difference < agreement_tolerance &
    score_agreement$maximum_Brier_difference < agreement_tolerance
)
scoring_decision <- if (all_paths_agree) {
  "The two functions agreed on both scales, so I used score_left_truncated() for the comparison."
} else {
  paste(
    "The functions did not agree on both scales. I therefore used score()",
    "with summary = FALSE for every landmark, method, and bootstrap sample,",
    "which kept the estimator identical across the comparison."
  )
}

knitr::kable(score_agreement, digits = 16)
cat(scoring_decision)
```

The apparent results below use `score()` on identical rows for the two
approaches.

```{r apparent-performance-table}
apparent_grid <- rbind(
  expand.grid(
    scale = "Enrollment time", landmark = enrollment_landmarks,
    model = c("Supermodel", "Disjoint"), stringsAsFactors = FALSE
  ),
  expand.grid(
    scale = "Attained age", landmark = age_landmarks,
    model = c("Supermodel", "Disjoint"), stringsAsFactors = FALSE
  )
)
apparent_performance <- merge(
  apparent_grid, ordinary_apparent_scores,
  by = c("scale", "landmark", "model"), all.x = TRUE, sort = FALSE
)
apparent_performance$performance_flag <- ifelse(
  is.finite(apparent_performance$AUC) & is.finite(apparent_performance$Brier),
  "ok", "prediction or score unavailable"
)
names(apparent_performance)[names(apparent_performance) == "model"] <- "method"
knitr::kable(apparent_performance, digits = 4)

if (nrow(apparent_prediction_failures)) {
  knitr::kable(apparent_prediction_failures)
}
```

## Existing summary-metric behavior

The current summary implementation assumes that the first landmark contains
every subject who appears later. That assumption holds after a shared
enrollment origin but fails under staggered age entry. I retained the exact
age-scale failure and the pending analytic variance statement.

```{r summary-performance}
enrollment_score_record <- capture_conditions(score(
  list("Enrollment scale" = enrollment_predictions),
  times = enrollment_landmarks,
  metrics = c("auc", "brier"), id_col = "id", cause = cause_of_interest,
  summary = TRUE, contrasts = FALSE
))
age_summary_failure <- capture_conditions(score(
  list("Age scale" = age_predictions),
  times = age_landmarks,
  metrics = c("auc", "brier"), id_col = "id", cause = cause_of_interest,
  summary = TRUE, contrasts = FALSE
))
age_left_truncated_score_record <- capture_conditions(score_left_truncated(
  risk = setNames(data.frame(age_predictions$preds$risk), "Age scale"),
  data = age_predictions$data,
  time_col = "exit_age", status_col = "status", entry_col = "LM",
  landmark_col = "LM", id_col = "id", cause = cause_of_interest,
  w = prediction_window
))

score_condition_table <- data.frame(
  operation = c(
    "Enrollment score with summary", "Age score without summary",
    "Age score with summary", "Age left-truncated point summary"
  ),
  warning_count = c(
    length(enrollment_score_record$warnings),
    length(age_landmark_score_record$warnings),
    length(age_summary_failure$warnings),
    length(age_left_truncated_score_record$warnings)
  ),
  error = c(
    if (inherits(enrollment_score_record$value, "error")) conditionMessage(enrollment_score_record$value) else "",
    if (inherits(age_landmark_score_record$value, "error")) conditionMessage(age_landmark_score_record$value) else "",
    if (inherits(age_summary_failure$value, "error")) conditionMessage(age_summary_failure$value) else "",
    if (inherits(age_left_truncated_score_record$value, "error")) conditionMessage(age_left_truncated_score_record$value) else ""
  ),
  row.names = NULL
)
knitr::kable(score_condition_table)

stopifnot(
  !inherits(enrollment_score_record$value, "error"),
  inherits(age_summary_failure$value, "error"),
  identical(conditionMessage(age_summary_failure$value), "Subsample size mismatch."),
  !inherits(age_left_truncated_score_record$value, "error")
)

enrollment_scores <- enrollment_score_record$value
age_left_truncated_scores <- age_left_truncated_score_record$value
enrollment_auc_summary <- enrollment_scores$AUC_summary$score
enrollment_auc_summary <- enrollment_auc_summary[
  enrollment_auc_summary$model == "Enrollment scale",
]
enrollment_brier_summary <- enrollment_scores$Brier_summary$score
enrollment_brier_summary <- enrollment_brier_summary[
  enrollment_brier_summary$model == "Enrollment scale",
]
age_summary <- age_left_truncated_scores$summary

summary_performance <- data.frame(
  scale = c("Enrollment time", "Attained age"),
  summary_auc = c(enrollment_auc_summary$AUC, age_summary$AUC),
  summary_auc_standard_error = c(enrollment_auc_summary$se, NA_real_),
  summary_brier = c(enrollment_brier_summary$Brier, age_summary$Brier),
  summary_brier_standard_error = c(enrollment_brier_summary$se, NA_real_),
  inference = c(
    "Original analytic summary calculation",
    age_left_truncated_scores$analytic_influence_function
  ),
  row.names = NULL
)
knitr::kable(summary_performance, digits = 4)
```

These summary values answer different prediction questions in different
landmark populations. I treated them as descriptive rather than as evidence
that one clock was better.

## Harrell-bootstrap optimism correction

The bootstrap used `r n_boot` subject resamples and seed `r bootstrap_seed`.
The transformations, landmarks, and prediction window stayed fixed at their
original values.

```{r harrell-bootstrap, cache=FALSE}
build_analysis_stacks <- function(longitudinal_data) {
  enrollment_raw <- stack_data(
    data = longitudinal_data,
    outcome = list(time = "time", status = "status"),
    lms = enrollment_landmarks, w = prediction_window,
    covs = list(
      fixed = c(shared_fixed_covariates, "age_at_enrollment"),
      varying = updated_laboratory_covariates
    ),
    format = "long", id = "id", rtime = "tstart"
  )
  enrollment_raw$data$attained_age_z <- (
    enrollment_raw$data$age_at_enrollment + enrollment_raw$data$LM - age_center
  ) / age_scale

  age_long <- transform(
    longitudinal_data,
    entry_age = age_at_enrollment,
    exit_age = age_at_enrollment + time,
    visit_age = age_at_enrollment + tstart
  )
  age_raw <- stack_data(
    data = age_long,
    outcome = list(time = "exit_age", status = "status"),
    lms = age_landmarks, w = prediction_window,
    covs = list(
      fixed = shared_fixed_covariates,
      varying = updated_laboratory_covariates
    ),
    format = "long", id = "id", rtime = "visit_age",
    entry = "entry_age", entry_mode = "strict"
  )

  enrollment_model <- add_interactions(
    enrollment_raw,
    lm_covs = updated_laboratory_covariates,
    func_covars = enrollment_basis,
    func_lms = enrollment_basis
  )
  age_model <- add_interactions(
    age_raw,
    lm_covs = updated_laboratory_covariates,
    func_covars = age_basis,
    func_lms = list()
  )
  stopifnot(
    identical(
      prediction_key(enrollment_raw$data),
      prediction_key(enrollment_model$data)
    ),
    identical(prediction_key(age_raw$data), prediction_key(age_model$data)),
    anyDuplicated(enrollment_raw$data[c("id", "LM")]) == 0L,
    anyDuplicated(enrollment_model$data[c("id", "LM")]) == 0L,
    anyDuplicated(age_raw$data[c("id", "LM")]) == 0L,
    anyDuplicated(age_model$data[c("id", "LM")]) == 0L
  )
  list(
    enrollment_raw = enrollment_raw, enrollment_model = enrollment_model,
    age_raw = age_raw, age_model = age_model
  )
}

resample_subjects <- function(longitudinal_data, sampled_ids) {
  pieces <- lapply(seq_along(sampled_ids), function(index) {
    subject_rows <- longitudinal_data[
      longitudinal_data$id == sampled_ids[index], , drop = FALSE
    ]
    stopifnot(nrow(subject_rows) > 0L)
    subject_rows$.bootstrap_source_id <- subject_rows$id
    subject_rows$id <- index
    subject_rows
  })
  out <- do.call(rbind, pieces)
  rownames(out) <- NULL
  stopifnot(
    length(unique(out$id)) == length(sampled_ids),
    all(vapply(
      split(out$.bootstrap_source_id, out$id),
      function(source_ids) length(unique(source_ids)) == 1L,
      logical(1L)
    ))
  )
  out$.bootstrap_source_id <- NULL
  out
}

fit_bootstrap_disjoint <- function(lmdata, landmarks, formula) {
  setNames(lapply(landmarks, function(landmark) {
    capture_conditions(dynamic_lm(
      subset_lmdata(lmdata, landmark), formula,
      type = "CSC", method = "breslow", x = TRUE
    ))
  }), as.character(landmarks))
}

bind_non_null <- function(objects) {
  objects <- objects[!vapply(objects, is.null, logical(1L))]
  if (length(objects)) do.call(rbind, objects) else NULL
}

score_prediction_diagnostic <- function(prediction, landmark) {
  record <- capture_conditions(score(
    list(Model = prediction), times = landmark,
    metrics = c("auc", "brier"), id_col = "id",
    cause = cause_of_interest, summary = FALSE, contrasts = FALSE
  ))
  if (inherits(record$value, "error")) {
    return(list(ok = FALSE, error = conditionMessage(record$value)))
  }
  auc <- as.data.frame(record$value$AUC$score)
  brier <- as.data.frame(record$value$Brier$score)
  auc <- auc[as.character(auc$model) == "Model", , drop = FALSE]
  brier <- brier[as.character(brier$model) == "Model", , drop = FALSE]
  ok <- nrow(auc) == 1L && nrow(brier) == 1L &&
    is.finite(auc$AUC) && is.finite(brier$Brier)
  list(ok = ok, error = if (ok) "" else "non-finite or missing score")
}

direct_age_cause_fit <- function(lmdata, formula, cause) {
  rhs <- paste(deparse(formula[[3L]]), collapse = "")
  direct_formula <- as.formula(paste0(
    "Surv(LM, exit_age, status == ", cause, ") ~ ", rhs
  ))
  capture_conditions(survival::coxph(
    direct_formula, data = lmdata$data, method = "breslow", x = TRUE
  ))
}

diagnose_age_fit <- function(fit_record, training_stack, formula, landmarks,
                             replicate, method) {
  rows <- list()
  model_error <- if (inherits(fit_record$value, "error")) {
    conditionMessage(fit_record$value)
  } else ""

  for (landmark in landmarks) {
    one_landmark <- subset_lmdata(training_stack, landmark)
    for (cause in c(transplant_code, cause_of_interest)) {
      event_count <- sum(one_landmark$data$status == cause)
      cause_fit <- NULL
      coefficient_names <- character()
      if (!inherits(fit_record$value, "error")) {
        cause_fit <- extract_cause_fit(fit_record$value, cause)
        coefficient_names <- names(coef(cause_fit))[!is.finite(coef(cause_fit))]
      }
      needs_direct_check <- inherits(fit_record$value, "error") ||
        length(fit_record$warnings) > 0L || length(coefficient_names) > 0L
      direct_record <- if (needs_direct_check) {
        direct_age_cause_fit(training_stack, formula, cause)
      } else NULL
      cause_error <- if (!is.null(direct_record) &&
                         inherits(direct_record$value, "error")) {
        conditionMessage(direct_record$value)
      } else ""
      cause_warning <- if (!is.null(direct_record)) {
        paste(direct_record$warnings, collapse = " | ")
      } else ""
      if (!is.null(direct_record) && !inherits(direct_record$value, "error")) {
        direct_coefficients <- coef(direct_record$value)
        coefficient_names <- union(
          coefficient_names,
          names(direct_coefficients)[!is.finite(direct_coefficients)]
        )
      }
      cause_event_times <- training_stack$data$exit_age[
        training_stack$data$status == cause
      ]
      last_event_time <- if (length(cause_event_times)) {
        max(cause_event_times)
      } else NA_real_
      horizon <- landmark + prediction_window
      rows[[length(rows) + 1L]] <- data.frame(
        replicate = replicate,
        method = method,
        landmark = landmark,
        cause = cause,
        events = event_count,
        model_fit_error = model_error,
        cause_fit_error = cause_error,
        cause_fit_warning = cause_warning,
        nonfinite_coefficients = paste(coefficient_names, collapse = ", "),
        horizon = horizon,
        last_event_time = last_event_time,
        horizon_past_last_event = is.finite(last_event_time) &&
          horizon > last_event_time,
        row.names = NULL
      )
    }
  }
  do.call(rbind, rows)
}

diagnose_age_predictions <- function(fit_record, training_stack, test_stack,
                                     landmarks, replicate, method) {
  rows <- list()
  model_error <- inherits(fit_record$value, "error")
  predictions <- list()
  for (sample_name in c("bootstrap_train", "original_test")) {
    target <- if (sample_name == "bootstrap_train") training_stack else test_stack
    predictions[[sample_name]] <- if (model_error) {
      list(value = fit_record$value, warnings = character(), messages = character())
    } else {
      capture_conditions(predict(
        fit_record$value, newdata = target, cause = cause_of_interest,
        complete = FALSE
      ))
    }
  }

  for (landmark in landmarks) {
    row <- data.frame(
      replicate = replicate,
      method = method,
      landmark = landmark,
      train_prediction_error = "",
      test_prediction_error = "",
      train_na_predictions = NA_integer_,
      test_na_predictions = NA_integer_,
      train_scoring_error = "",
      test_scoring_error = "",
      row.names = NULL
    )
    for (sample_name in names(predictions)) {
      prefix <- if (sample_name == "bootstrap_train") "train" else "test"
      prediction_record <- predictions[[sample_name]]
      if (inherits(prediction_record$value, "error")) {
        row[[paste0(prefix, "_prediction_error")]] <-
          conditionMessage(prediction_record$value)
        next
      }
      prediction <- subset_prediction(prediction_record$value, landmark)
      na_predictions <- sum(is.na(prediction$preds$risk))
      row[[paste0(prefix, "_na_predictions")]] <- na_predictions
      score_record <- if (na_predictions > 0L) {
        list(ok = FALSE, error = "NA predictions")
      } else {
        score_prediction_diagnostic(prediction, landmark)
      }
      row[[paste0(prefix, "_scoring_error")]] <- score_record$error
    }
    rows[[length(rows) + 1L]] <- row
  }
  do.call(rbind, rows)
}

diagnose_age_model <- function(fit_record, training_stack, test_stack, formula,
                               landmarks, replicate, method) {
  list(
    fit = diagnose_age_fit(
      fit_record, training_stack, formula, landmarks, replicate, method
    ),
    prediction = diagnose_age_predictions(
      fit_record, training_stack, test_stack, landmarks, replicate, method
    )
  )
}

score_one_prediction <- function(prediction, landmark) {
  record <- capture_conditions(score(
    list(Model = prediction), times = landmark,
    metrics = c("auc", "brier"), id_col = "id",
    cause = cause_of_interest, summary = FALSE, contrasts = FALSE
  ))
  if (inherits(record$value, "error")) return(NULL)
  auc <- as.data.frame(record$value$AUC$score)
  brier <- as.data.frame(record$value$Brier$score)
  auc$model <- as.character(auc$model)
  brier$model <- as.character(brier$model)
  auc <- auc[auc$model == "Model", ]
  brier <- brier[brier$model == "Model", ]
  if (nrow(auc) != 1L || nrow(brier) != 1L ||
      !is.finite(auc$AUC) || !is.finite(brier$Brier)) return(NULL)
  data.frame(AUC = auc$AUC, Brier = brier$Brier)
}

evaluate_supermodel_bootstrap <- function(fit_record, training_stack,
                                          test_stack, landmarks,
                                          replicate, scale_label) {
  if (inherits(fit_record$value, "error")) return(NULL)
  training_prediction <- capture_conditions(predict(
    fit_record$value, newdata = training_stack, cause = cause_of_interest
  ))
  test_prediction <- capture_conditions(predict(
    fit_record$value, newdata = test_stack, cause = cause_of_interest
  ))
  if (inherits(training_prediction$value, "error") ||
      inherits(test_prediction$value, "error")) return(NULL)

  rows <- list()
  for (landmark in landmarks) {
    for (sample_name in c("bootstrap_train", "original_test")) {
      prediction <- if (sample_name == "bootstrap_train") {
        training_prediction$value
      } else test_prediction$value
      metric <- score_one_prediction(
        subset_prediction(prediction, landmark), landmark
      )
      if (!is.null(metric)) {
        rows[[length(rows) + 1L]] <- data.frame(
          replicate = replicate, scale = scale_label,
          method = "Supermodel", landmark = landmark,
          sample = sample_name, AUC = metric$AUC, Brier = metric$Brier
        )
      }
    }
  }
  if (length(rows)) do.call(rbind, rows) else NULL
}

evaluate_disjoint_bootstrap <- function(fit_records, training_stack,
                                        test_stack, landmarks,
                                        replicate, scale_label) {
  rows <- list()
  for (landmark in landmarks) {
    record <- fit_records[[as.character(landmark)]]
    if (inherits(record$value, "error")) next
    training_prediction <- capture_conditions(predict(
      record$value, newdata = subset_lmdata(training_stack, landmark),
      cause = cause_of_interest
    ))
    test_prediction <- capture_conditions(predict(
      record$value, newdata = subset_lmdata(test_stack, landmark),
      cause = cause_of_interest
    ))
    predictions <- list(
      bootstrap_train = training_prediction,
      original_test = test_prediction
    )
    for (sample_name in names(predictions)) {
      prediction_record <- predictions[[sample_name]]
      if (inherits(prediction_record$value, "error")) next
      metric <- score_one_prediction(prediction_record$value, landmark)
      if (!is.null(metric)) {
        rows[[length(rows) + 1L]] <- data.frame(
          replicate = replicate, scale = scale_label,
          method = "Disjoint", landmark = landmark,
          sample = sample_name, AUC = metric$AUC, Brier = metric$Brier
        )
      }
    }
  }
  if (length(rows)) do.call(rbind, rows) else NULL
}

original_stacks <- list(
  enrollment_raw = enrollment_stack_raw,
  enrollment_model = enrollment_stack,
  age_raw = age_stack_raw,
  age_model = age_stack
)

set.seed(bootstrap_seed)
original_ids <- unique(pbc_analysis$id)
bootstrap_samples <- lapply(seq_len(n_boot), function(replicate) {
  sample(original_ids, length(original_ids), replace = TRUE)
})

run_bootstrap_replicate <- function(replicate) {
  bootstrap_long_record <- capture_conditions(resample_subjects(
    pbc_analysis, bootstrap_samples[[replicate]]
  ))
  if (inherits(bootstrap_long_record$value, "error")) {
    return(list(
      scores = NULL,
      stack = data.frame(
        replicate = replicate,
        resample_error = conditionMessage(bootstrap_long_record$value),
        stack_build_error = "",
        age_raw_duplicate_id_lm = NA_integer_,
        age_model_duplicate_id_lm = NA_integer_
      ),
      fit = NULL, prediction = NULL, unexpected = NULL
    ))
  }
  stacks_record <- capture_conditions(build_analysis_stacks(
    bootstrap_long_record$value
  ))
  if (inherits(stacks_record$value, "error")) {
    return(list(
      scores = NULL,
      stack = data.frame(
        replicate = replicate,
        resample_error = "",
        stack_build_error = conditionMessage(stacks_record$value),
        age_raw_duplicate_id_lm = NA_integer_,
        age_model_duplicate_id_lm = NA_integer_
      ),
      fit = NULL, prediction = NULL, unexpected = NULL
    ))
  }
  stacks <- stacks_record$value
  stack_diagnostic <- data.frame(
    replicate = replicate,
    resample_error = "",
    stack_build_error = "",
    age_raw_duplicate_id_lm = anyDuplicated(
      stacks$age_raw$data[c("id", "LM")]
    ),
    age_model_duplicate_id_lm = anyDuplicated(
      stacks$age_model$data[c("id", "LM")]
    )
  )

  enrollment_super_fit <- capture_conditions(dynamic_lm(
    stacks$enrollment_model, enrollment_formula,
    type = "CSC", method = "breslow", x = TRUE
  ))
  age_super_fit <- capture_conditions(dynamic_lm(
    stacks$age_model, age_formula,
    type = "CSC", method = "breslow", x = TRUE
  ))
  enrollment_disjoint_fits <- fit_bootstrap_disjoint(
    stacks$enrollment_raw, enrollment_landmarks,
    enrollment_disjoint_formula
  )
  age_disjoint_fits <- fit_bootstrap_disjoint(
    stacks$age_raw, age_landmarks, age_disjoint_formula
  )

  score_rows <- list(
    evaluate_supermodel_bootstrap(
      enrollment_super_fit, stacks$enrollment_model,
      original_stacks$enrollment_model, enrollment_landmarks,
      replicate, "Enrollment time"
    ),
    evaluate_disjoint_bootstrap(
      enrollment_disjoint_fits, stacks$enrollment_raw,
      original_stacks$enrollment_raw, enrollment_landmarks,
      replicate, "Enrollment time"
    ),
    evaluate_supermodel_bootstrap(
      age_super_fit, stacks$age_model, original_stacks$age_model,
      age_landmarks, replicate, "Attained age"
    ),
    evaluate_disjoint_bootstrap(
      age_disjoint_fits, stacks$age_raw, original_stacks$age_raw,
      age_landmarks, replicate, "Attained age"
    )
  )
  super_diagnostic <- diagnose_age_model(
    age_super_fit, stacks$age_model, original_stacks$age_model,
    age_formula, age_landmarks, replicate, "Supermodel"
  )
  disjoint_diagnostics <- lapply(age_landmarks, function(landmark) {
    diagnose_age_model(
      age_disjoint_fits[[as.character(landmark)]],
      subset_lmdata(stacks$age_raw, landmark),
      subset_lmdata(original_stacks$age_raw, landmark),
      age_disjoint_formula, landmark, replicate, "Disjoint"
    )
  })

  list(
    scores = bind_non_null(score_rows),
    stack = stack_diagnostic,
    fit = bind_non_null(c(
      list(super_diagnostic$fit),
      lapply(disjoint_diagnostics, `[[`, "fit")
    )),
    prediction = bind_non_null(c(
      list(super_diagnostic$prediction),
      lapply(disjoint_diagnostics, `[[`, "prediction")
    )),
    unexpected = NULL
  )
}

bootstrap_cores <- getOption(
  "timescale.bootstrap.cores",
  if (.Platform$OS.type == "windows") 1L else min(4L, parallel::detectCores(logical = FALSE))
)
safe_bootstrap_replicate <- function(replicate) {
  tryCatch(
    run_bootstrap_replicate(replicate),
    error = function(error) list(
      scores = NULL, stack = NULL, fit = NULL, prediction = NULL,
      unexpected = data.frame(
        replicate = replicate,
        error = conditionMessage(error),
        row.names = NULL
      )
    )
  )
}
if (.Platform$OS.type == "windows" || bootstrap_cores <= 1L) {
  bootstrap_results <- lapply(seq_len(n_boot), safe_bootstrap_replicate)
} else {
  bootstrap_results <- parallel::mclapply(
    seq_len(n_boot), safe_bootstrap_replicate,
    mc.cores = bootstrap_cores, mc.preschedule = FALSE
  )
}
bootstrap_scores <- bind_non_null(lapply(bootstrap_results, `[[`, "scores"))
age_stack_diagnostics <- bind_non_null(lapply(bootstrap_results, `[[`, "stack"))
age_fit_diagnostics <- bind_non_null(lapply(bootstrap_results, `[[`, "fit"))
age_prediction_diagnostics <- bind_non_null(lapply(
  bootstrap_results, `[[`, "prediction"
))
unexpected_bootstrap_errors <- bind_non_null(lapply(
  bootstrap_results, `[[`, "unexpected"
))
stopifnot(!is.null(bootstrap_scores))
rownames(bootstrap_scores) <- NULL

bootstrap_train <- bootstrap_scores[
  bootstrap_scores$sample == "bootstrap_train",
  c("replicate", "scale", "method", "landmark", "AUC", "Brier")
]
bootstrap_test <- bootstrap_scores[
  bootstrap_scores$sample == "original_test",
  c("replicate", "scale", "method", "landmark", "AUC", "Brier")
]
paired_bootstrap <- merge(
  bootstrap_train, bootstrap_test,
  by = c("replicate", "scale", "method", "landmark"),
  suffixes = c("_train", "_test")
)

bootstrap_optimism <- do.call(rbind, lapply(
  split(paired_bootstrap, list(
    paired_bootstrap$scale, paired_bootstrap$method,
    paired_bootstrap$landmark
  ), drop = TRUE),
  function(rows) {
    auc_optimism <- rows$AUC_train - rows$AUC_test
    brier_optimism <- rows$Brier_train - rows$Brier_test
    data.frame(
      scale = rows$scale[1L], method = rows$method[1L],
      landmark = rows$landmark[1L],
      successful_replicates = length(unique(rows$replicate)),
      failed_replicates = n_boot - length(unique(rows$replicate)),
      mean_AUC_optimism = mean(auc_optimism),
      Monte_Carlo_SE_AUC_optimism = sd(auc_optimism) / sqrt(length(auc_optimism)),
      mean_Brier_optimism = mean(brier_optimism),
      Monte_Carlo_SE_Brier_optimism = sd(brier_optimism) / sqrt(length(brier_optimism)),
      row.names = NULL
    )
  }
))
rownames(bootstrap_optimism) <- NULL

optimism_corrected_performance <- merge(
  apparent_performance, bootstrap_optimism,
  by = c("scale", "method", "landmark"), all.x = TRUE, sort = FALSE
)
optimism_corrected_performance$successful_replicates[
  is.na(optimism_corrected_performance$successful_replicates)
] <- 0L
optimism_corrected_performance$failed_replicates[
  is.na(optimism_corrected_performance$failed_replicates)
] <- n_boot
optimism_corrected_performance$optimism_corrected_AUC <- with(
  optimism_corrected_performance, AUC - mean_AUC_optimism
)
optimism_corrected_performance$optimism_corrected_Brier <- with(
  optimism_corrected_performance, Brier - mean_Brier_optimism
)

stopifnot(
  all(optimism_corrected_performance$successful_replicates >= 0L),
  all(optimism_corrected_performance$failed_replicates >= 0L),
  all(optimism_corrected_performance$failed_replicates <= n_boot),
  all(
    optimism_corrected_performance$successful_replicates +
      optimism_corrected_performance$failed_replicates == n_boot
  )
)

age_stack_summary <- data.frame(
  bootstrap_replicates = n_boot,
  resample_errors = sum(nzchar(age_stack_diagnostics$resample_error)),
  stack_build_errors = sum(nzchar(age_stack_diagnostics$stack_build_error)),
  maximum_age_raw_duplicate_id_lm = max(
    age_stack_diagnostics$age_raw_duplicate_id_lm, na.rm = TRUE
  ),
  maximum_age_model_duplicate_id_lm = max(
    age_stack_diagnostics$age_model_duplicate_id_lm, na.rm = TRUE
  ),
  unexpected_replicate_errors = if (is.null(unexpected_bootstrap_errors)) {
    0L
  } else nrow(unexpected_bootstrap_errors),
  row.names = NULL
)

age_score_pair_counts <- aggregate(
  replicate ~ method + landmark,
  paired_bootstrap[paired_bootstrap$scale == "Attained age", ],
  function(replicates) length(unique(replicates))
)
names(age_score_pair_counts)[names(age_score_pair_counts) == "replicate"] <-
  "complete_score_pairs"
age_score_pair_grid <- expand.grid(
  method = c("Supermodel", "Disjoint"),
  landmark = age_landmarks,
  stringsAsFactors = FALSE
)
age_score_pair_counts <- merge(
  age_score_pair_grid, age_score_pair_counts,
  by = c("method", "landmark"), all.x = TRUE, sort = FALSE
)
age_score_pair_counts$complete_score_pairs[
  is.na(age_score_pair_counts$complete_score_pairs)
] <- 0L
age_score_pair_counts$failed_score_pairs <-
  n_boot - age_score_pair_counts$complete_score_pairs

age_fit_failure_tally <- do.call(rbind, lapply(
  split(age_fit_diagnostics, list(
    age_fit_diagnostics$method,
    age_fit_diagnostics$landmark,
    age_fit_diagnostics$cause
  ), drop = TRUE),
  function(rows) {
    data.frame(
      method = rows$method[1L],
      landmark = rows$landmark[1L],
      cause = rows$cause[1L],
      zero_event_replicates = sum(rows$events == 0L),
      model_fit_error_replicates = sum(nzchar(rows$model_fit_error)),
      cause_fit_error_replicates = sum(nzchar(rows$cause_fit_error)),
      warning_replicates = sum(nzchar(rows$cause_fit_warning)),
      nonfinite_coefficient_replicates = sum(
        nzchar(rows$nonfinite_coefficients)
      ),
      horizon_past_last_event_replicates = sum(
        rows$horizon_past_last_event
      ),
      row.names = NULL
    )
  }
))
age_fit_failure_tally <- merge(
  age_fit_failure_tally, age_score_pair_counts,
  by = c("method", "landmark"), all.x = TRUE, sort = FALSE
)
rownames(age_fit_failure_tally) <- NULL

age_prediction_failure_tally <- do.call(rbind, lapply(
  split(age_prediction_diagnostics, list(
    age_prediction_diagnostics$method,
    age_prediction_diagnostics$landmark
  ), drop = TRUE),
  function(rows) {
    data.frame(
      method = rows$method[1L],
      landmark = rows$landmark[1L],
      train_prediction_error_replicates = sum(nzchar(
        rows$train_prediction_error
      )),
      test_prediction_error_replicates = sum(nzchar(
        rows$test_prediction_error
      )),
      train_NA_prediction_replicates = sum(
        rows$train_na_predictions > 0L, na.rm = TRUE
      ),
      test_NA_prediction_replicates = sum(
        rows$test_na_predictions > 0L, na.rm = TRUE
      ),
      train_scoring_error_replicates = sum(nzchar(
        rows$train_scoring_error
      )),
      test_scoring_error_replicates = sum(nzchar(
        rows$test_scoring_error
      )),
      row.names = NULL
    )
  }
))
age_prediction_failure_tally <- merge(
  age_prediction_failure_tally, age_score_pair_counts,
  by = c("method", "landmark"), all.x = TRUE, sort = FALSE
)
rownames(age_prediction_failure_tally) <- NULL

failure_reason_rows <- list()
normalize_failure_reason <- function(reason) {
  if (grepl("^arguments imply differing number of rows:", reason)) {
    return("zero-event model returned no coefficient rows")
  }
  if (grepl("infinite argument in coxph.wtest", reason, fixed = TRUE)) {
    return("Cox Wald calculation had an infinite argument")
  }
  if (grepl("Ran out of iterations", reason, fixed = TRUE)) {
    return("Cox fit did not converge")
  }
  if (grepl("beta may be infinite", reason, fixed = TRUE)) {
    return("one or more Cox coefficients may be infinite")
  }
  reason
}
append_failure_reason <- function(method, landmark, cause, stage, reason) {
  failure_reason_rows[[length(failure_reason_rows) + 1L]] <<- data.frame(
    method = method,
    landmark = landmark,
    cause = as.character(cause),
    stage = stage,
    reason = normalize_failure_reason(reason),
    stringsAsFactors = FALSE
  )
}
for (row_index in seq_len(nrow(age_fit_diagnostics))) {
  row <- age_fit_diagnostics[row_index, ]
  if (row$events == 0L) {
    append_failure_reason(
      row$method, row$landmark, row$cause,
      "event count", "zero events inside the prediction window"
    )
  }
  if (nzchar(row$model_fit_error)) {
    append_failure_reason(
      row$method, row$landmark, row$cause,
      "competing-risk fit", row$model_fit_error
    )
  }
  if (nzchar(row$cause_fit_error)) {
    append_failure_reason(
      row$method, row$landmark, row$cause,
      "cause-specific fit", row$cause_fit_error
    )
  }
  if (nzchar(row$cause_fit_warning)) {
    append_failure_reason(
      row$method, row$landmark, row$cause,
      "cause-specific fit", row$cause_fit_warning
    )
  }
  if (nzchar(row$nonfinite_coefficients)) {
    append_failure_reason(
      row$method, row$landmark, row$cause,
      "cause-specific fit",
      "one or more non-finite coefficients"
    )
  }
}
for (row_index in seq_len(nrow(age_prediction_diagnostics))) {
  row <- age_prediction_diagnostics[row_index, ]
  prediction_fields <- c(
    train_prediction_error = "train prediction",
    test_prediction_error = "test prediction",
    train_scoring_error = "train scoring",
    test_scoring_error = "test scoring"
  )
  for (field in names(prediction_fields)) {
    if (nzchar(row[[field]])) {
      append_failure_reason(
        row$method, row$landmark, "joint CIF",
        unname(prediction_fields[field]), row[[field]]
      )
    }
  }
  if (is.finite(row$train_na_predictions) && row$train_na_predictions > 0L) {
    append_failure_reason(
      row$method, row$landmark, "joint CIF",
      "train prediction", "one or more NA risks"
    )
  }
  if (is.finite(row$test_na_predictions) && row$test_na_predictions > 0L) {
    append_failure_reason(
      row$method, row$landmark, "joint CIF",
      "test prediction", "one or more NA risks"
    )
  }
}
age_failure_reason_tally <- aggregate(
  list(replicates = rep(1L, length(failure_reason_rows))),
  by = do.call(rbind, failure_reason_rows),
  FUN = sum
)
age_failure_reason_tally <- age_failure_reason_tally[
  order(
    age_failure_reason_tally$method,
    age_failure_reason_tally$landmark,
    age_failure_reason_tally$cause,
    age_failure_reason_tally$stage,
    age_failure_reason_tally$reason
  ),
]
rownames(age_failure_reason_tally) <- NULL

super_transplant_zero_by_replicate <- aggregate(
  events ~ replicate,
  age_fit_diagnostics[
    age_fit_diagnostics$method == "Supermodel" &
      age_fit_diagnostics$cause == transplant_code,
  ],
  function(events) any(events == 0L)
)
names(super_transplant_zero_by_replicate)[2L] <-
  "zero_transplants_in_any_age_window"
age_super_pair_grid <- expand.grid(
  replicate = seq_len(n_boot),
  landmark = age_landmarks,
  stringsAsFactors = FALSE
)
age_super_complete_pairs <- paired_bootstrap[
  paired_bootstrap$scale == "Attained age" &
    paired_bootstrap$method == "Supermodel",
  c("replicate", "landmark")
]
age_super_complete_pairs$complete_score_pair <- TRUE
age_super_pair_grid <- merge(
  age_super_pair_grid, age_super_complete_pairs,
  by = c("replicate", "landmark"), all.x = TRUE, sort = FALSE
)
age_super_pair_grid$complete_score_pair[
  is.na(age_super_pair_grid$complete_score_pair)
] <- FALSE
age_super_pair_grid <- merge(
  age_super_pair_grid, super_transplant_zero_by_replicate,
  by = "replicate", all.x = TRUE, sort = FALSE
)
age_super_zero_event_link <- do.call(rbind, lapply(
  split(age_super_pair_grid, list(
    age_super_pair_grid$landmark,
    age_super_pair_grid$zero_transplants_in_any_age_window
  ), drop = TRUE),
  function(rows) {
    data.frame(
      landmark = rows$landmark[1L],
      zero_transplants_in_any_age_window =
        rows$zero_transplants_in_any_age_window[1L],
      replicates = nrow(rows),
      complete_score_pairs = sum(rows$complete_score_pair),
      failed_score_pairs = sum(!rows$complete_score_pair),
      row.names = NULL
    )
  }
))
rownames(age_super_zero_event_link) <- NULL

age_pair_status <- expand.grid(
  replicate = seq_len(n_boot),
  method = c("Supermodel", "Disjoint"),
  landmark = age_landmarks,
  stringsAsFactors = FALSE
)
complete_age_pairs <- paired_bootstrap[
  paired_bootstrap$scale == "Attained age",
  c("replicate", "method", "landmark")
]
complete_age_pairs$complete_score_pair <- TRUE
age_pair_status <- merge(
  age_pair_status, complete_age_pairs,
  by = c("replicate", "method", "landmark"), all.x = TRUE, sort = FALSE
)
age_pair_status$complete_score_pair[
  is.na(age_pair_status$complete_score_pair)
] <- FALSE
age_horizon_link <- merge(
  age_fit_diagnostics, age_pair_status,
  by = c("replicate", "method", "landmark"), all.x = TRUE, sort = FALSE
)
age_horizon_summary <- do.call(rbind, lapply(
  split(age_horizon_link, list(
    age_horizon_link$method,
    age_horizon_link$landmark,
    age_horizon_link$cause
  ), drop = TRUE),
  function(rows) {
    data.frame(
      method = rows$method[1L],
      landmark = rows$landmark[1L],
      cause = rows$cause[1L],
      horizon_past_last_event_replicates = sum(
        rows$horizon_past_last_event
      ),
      complete_pairs_despite_past_horizon = sum(
        rows$horizon_past_last_event & rows$complete_score_pair
      ),
      row.names = NULL
    )
  }
))
rownames(age_horizon_summary) <- NULL

enrollment_super_replicates_retained <- length(unique(
  paired_bootstrap$replicate[
    paired_bootstrap$scale == "Enrollment time" &
      paired_bootstrap$method == "Supermodel" &
      paired_bootstrap$landmark == enrollment_landmarks[1L]
  ]
))

age_performance_pair_counts <- optimism_corrected_performance[
  optimism_corrected_performance$scale == "Attained age",
  c("method", "landmark", "successful_replicates")
]
age_pair_count_validation <- merge(
  age_score_pair_counts, age_performance_pair_counts,
  by = c("method", "landmark"), all = TRUE, sort = FALSE
)

stopifnot(
  nrow(age_stack_diagnostics) == n_boot,
  age_stack_summary$maximum_age_raw_duplicate_id_lm == 0L,
  age_stack_summary$maximum_age_model_duplicate_id_lm == 0L,
  nrow(age_pair_count_validation) == nrow(age_score_pair_counts),
  all(age_pair_count_validation$complete_score_pairs ==
        age_pair_count_validation$successful_replicates),
  enrollment_super_replicates_retained == n_boot
)
```

```{r optimism-corrected-performance-table}
knitr::kable(
  optimism_corrected_performance[c(
    "scale", "landmark", "method", "AUC", "mean_AUC_optimism",
    "optimism_corrected_AUC", "Brier", "mean_Brier_optimism",
    "optimism_corrected_Brier", "successful_replicates",
    "failed_replicates"
  )],
  digits = 4
)
```

The failure table counts a replicate as successful only when both the
bootstrap-resample and original-cohort scores were available for that model and
landmark. This makes sparse disjoint-model failures visible in the denominator.

```{r bootstrap-failure-table}
bootstrap_failure_table <- optimism_corrected_performance[c(
  "scale", "landmark", "method", "successful_replicates", "failed_replicates"
)]
knitr::kable(bootstrap_failure_table)
```

### Age-scale bootstrap failure diagnosis

I first instrumented `r diagnostic_pilot_replicates` resamples while diagnosing
the failure path, then retained the same instrumentation for the complete
`r n_boot`-resample run reported here. The fit table is cause-specific. The
prediction table is model-specific because a death cumulative-incidence
prediction uses both the death and transplant cause-specific fits.

```{r age-bootstrap-diagnostic-tables}
zero_transplant_super_replicates <- sum(
  super_transplant_zero_by_replicate$zero_transplants_in_any_age_window
)
zero_transplant_pairs <- age_super_zero_event_link[
  age_super_zero_event_link$zero_transplants_in_any_age_window,
]
zero_transplant_pairs_all_failed <- all(
  zero_transplant_pairs$complete_score_pairs == 0L
)
zero_transplant_resample_word <- if (
  zero_transplant_super_replicates == 1L
) "resample" else "resamples"
whole_replicates_discarded <- n_boot - enrollment_super_replicates_retained
horizon_flag_has_successes <- any(
  age_horizon_summary$complete_pairs_despite_past_horizon > 0L
)
horizon_success_phrase <- if (horizon_flag_has_successes) {
  "did occur"
} else "did not occur"
age_failure_classification <- if (
  zero_transplant_super_replicates > 0L &&
    zero_transplant_pairs_all_failed &&
    whole_replicates_discarded == 0L
) {
  "genuine sparse competing-event fits, not an ID or whole-replicate control-flow bug"
} else {
  "not resolved by the recorded diagnostics"
}

knitr::kable(age_stack_summary)
knitr::kable(age_fit_failure_tally)
knitr::kable(age_prediction_failure_tally)
knitr::kable(age_failure_reason_tally)
knitr::kable(age_super_zero_event_link)
knitr::kable(age_horizon_summary)
```

All resampled subject copies had distinct analysis identifiers, and the maximum
duplicate count for `(id, LM)` was
`r age_stack_summary$maximum_age_model_duplicate_id_lm`. There were
`r age_stack_summary$stack_build_errors` stack-build errors and
`r age_stack_summary$unexpected_replicate_errors` unexpected replicate errors.
The enrollment supermodel retained complete pairs from all `r n_boot`
resamples, so the safety wrapper discarded `r whole_replicates_discarded` whole
replicates when an age component failed.

The dominant mechanism was sparse transplantation. In
`r zero_transplant_super_replicates` `r zero_transplant_resample_word`, at least
one age window had no transplants. Every such resample lacked a complete
age-supermodel score pair at
every age landmark. The cause-specific table shows the associated convergence
warnings and non-finite transplant-model coefficients. Because the age
supermodel uses one competing-risk fit across all landmarks, an undefined
transplant linear predictor prevents the death cumulative-incidence prediction
at every landmark; this is why the supermodel success counts were nearly
uniform. Scoring remained landmark-specific and did not discard the other age
landmarks when only a score calculation failed.

The horizon check did not explain the failures. Complete score pairs
`r horizon_success_phrase` in rows whose horizon passed the last observed event,
consistent with the step-function baseline hazard carrying forward after that
event. I therefore classified the failure as `r age_failure_classification`.

I did not alter the age model or replace failed replicates. The optimism-corrected
age values are conditional on the subset of resamples with finite competing-risk
fits and complete train/test scores, so I would not use them as primary
performance estimates. A future analysis could evaluate a denser age grid, a
simpler transplant model, or a prespecified penalized competing-risk fit, but
each would change the model and requires its own validation.

On the current attained-age grid, each landmark already has its own baseline
because the windows do not overlap. With `r length(age_landmarks)` landmarks
and a quadratic laboratory basis containing `r length(age_basis) + 1L`
coefficients per laboratory, the laboratory effects are also effectively
saturated across those landmarks. The primary attained-age supermodel therefore
pools only male sex, disease stage, and treatment across landmarks; it cannot
show the smoothing advantage that would require a denser, overlapping grid.

# Calibration

The apparent calibration plots rendered at every requested landmark. I used
the same quantile grouping on both scales and retained the condition table so a
future failure will be visible rather than silently dropped.

```{r enrollment-calibration, fig.width=10, fig.height=7}
old_parameters <- par(mfrow = c(2, 3), mar = c(4, 4, 3, 1))
enrollment_calibration_arguments <- list(
  object = list("Enrollment scale" = enrollment_predictions),
  times = enrollment_landmarks,
  id_col = "id",
  cause = cause_of_interest,
  method = "quantile",
  q = 4,
  legend = FALSE,
  xlim = c(0, 1),
  ylim = c(0, 1)
)
enrollment_calibration_record <- capture_conditions(do.call(
  calplot,
  enrollment_calibration_arguments
))
par(old_parameters)

if (inherits(enrollment_calibration_record$value, "error")) {
  stop(
    "Enrollment calibration failed: ",
    conditionMessage(enrollment_calibration_record$value)
  )
}
```

```{r age-calibration, fig.width=10, fig.height=4}
old_parameters <- par(mfrow = c(1, 3), mar = c(4, 4, 3, 1))
age_calibration_arguments <- list(
  object = list("Age scale" = age_predictions),
  times = age_landmarks,
  id_col = "id",
  cause = cause_of_interest,
  method = "quantile",
  q = 4,
  legend = FALSE,
  xlim = c(0, 1),
  ylim = c(0, 1)
)
age_calibration_record <- capture_conditions(do.call(
  calplot,
  age_calibration_arguments
))
par(old_parameters)

if (inherits(age_calibration_record$value, "error")) {
  stop(
    "Age calibration failed: ",
    conditionMessage(age_calibration_record$value)
  )
}
```

```{r calibration-conditions}
calibration_conditions <- data.frame(
  scale = c("Enrollment time", "Attained age"),
  warning_count = c(
    length(enrollment_calibration_record$warnings),
    length(age_calibration_record$warnings)
  ),
  error = c(
    if (inherits(enrollment_calibration_record$value, "error")) conditionMessage(enrollment_calibration_record$value) else "",
    if (inherits(age_calibration_record$value, "error")) conditionMessage(age_calibration_record$value) else ""
  ),
  row.names = NULL
)

knitr::kable(calibration_conditions)
```

# Interpretation

The enrollment analysis used a common origin and updated attained age as a
predictor. The attained-age analysis used age as the clock, excluded age from
the right-hand side, and enforced observation by each landmark. The substantive
non-age covariates and the two laboratory interaction structures were identical.
The enrollment supermodel additionally included estimable centered LM main
effects. The primary attained-age supermodel did not: its non-overlapping
windows made LM constant within every event risk set, and the demonstration fit
returned non-estimable LM coefficients.

The age-scale risk-set pattern illustrated the practical difference: people
could enter as age advanced, so a later age landmark could contain more eligible
subjects than an earlier one. Strict entry made that membership rule explicit.

The disjoint fits showed the cost of spending a full coefficient vector at each
landmark. I retained every warning and non-estimable coefficient in the output,
and I left unavailable predictions missing rather than changing a sparse model.
The bootstrap likewise reports the number of complete resample-and-test pairs
for every model and landmark, so the corrected performance estimates can be
read together with their failure counts.

The bootstrap diagnosis traced the low attained-age completion rate to sparse
transplant fits, not duplicated bootstrap identifiers or whole-replicate loss.
Because the death cumulative incidence depends on the transplant hazard, a
non-finite transplant fit invalidated predictions from the shared age
supermodel. I therefore treated the corrected attained-age performance as
conditional and exploratory rather than as a primary estimate.

This attained-age comparison cannot demonstrate smoothing of the laboratory
effects. Its disjoint windows give each landmark a separate baseline, and its
quadratic laboratory basis is effectively saturated over the observed age
landmarks. On this grid, the attained-age supermodel pools male sex, disease
stage, and treatment; a denser overlapping grid would be needed to study the
additional pooling available from an LM main effect and smoother laboratory
trajectories.

Within-window delayed entry was not part of this primary comparison. In the
separate NAFLD application, it added very little analyzable longitudinal
information because most later entrants lacked a laboratory value at their
analysis entry. That finding supports treating strict entry as the clearer
primary analysis rather than assuming delayed entry will automatically recover
substantial usable data.

# Session information

```{r session-information}
sessionInfo()
```
